### 2017-08-29 README3.Rmd
updating Rmd page

---
title: "MLH1_md"
author: "April"
date: "August 29, 2017"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
library(plyr) 
library(gplots)
library(ggplot2)
library(knitr) #for md tables 
#read in working environment that was already set
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

##FIX ORDER HERE
MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
                                     "WSB", "PWD","MSM","CAST"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
#set the order of another column, based on another variable (so that when)
library(dplyr)
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(file.name = factor(file.name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

#save a figure of boxplots for each mouse
ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )

ggsave("boxplots_mice.png")

MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]

ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
ggsave("Female_boxplots.png")

ggplot(MLH1_by_M_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
ggsave("Male_boxplots.png")
```

This is where the plot is discribed.

```{r setup, include=FALSE, echo=FALSE}                 
#Quality / number of passing mice
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table

passed_mice <- as.data.frame(q_cutoff_table[q_cutoff_table$q_l3 >= 15,]$mouse, 
                             col.names= "mouse")
colnames(passed_mice) <- c("mouse")

passed_mice$strain <- ifelse(grepl("WSB", passed_mice$mouse), "WSB", 
                   ifelse(grepl("G", passed_mice$mouse), "G",
                    ifelse(grepl("CAST", passed_mice$mouse), "CAST",
                   ifelse(grepl("MSM", passed_mice$mouse), "MSM",
                  ifelse(grepl("LEW", passed_mice$mouse), "LEWES", 
                   ifelse(grepl("LEWES", passed_mice$mouse), "LEWES",                                          
                  ifelse(grepl("PWD", passed_mice$mouse), "PWD", "other")))))))

passed_mice$sex <- ifelse(grepl("_f", passed_mice$mouse), "female",                                         
                          ifelse(grepl("_m", passed_mice$mouse), "male", "other"))

#make a table of the number of mice passed for each category.
Pass_mice_category <- ddply(passed_mice, .(strain, sex), summarise,
                    "Passing Mice" =  length(mouse)
)

not_passed_mice <- setdiff(q_cutoff_table$mouse, passed_mice$mouse)
not_passed_mice <- data.frame(not_passed_mice)
colnames(not_passed_mice) <- c("mouse")

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")

#convert tables into md useing Kable
Pass_mice_category_kable <- kable(Pass_mice_category, format = "markdown")
passed_mice_kable <- kable(passed_mice, format = "markdown")
not_passed_mice_kable <- kable(not_passed_mice, format = "markdown")

```

#### Tables of Passing mice
Pass_mice_category_kable

another table

passed_mice_kable
```{r setup, include=FALSE, echo=FALSE}

passed_mice_kable

```