### 2017-08-29 rmd_tutorial.rmd
tryiung the tutorial

---
title: "kableExtraTutorial"
author: "April"
date: "August 29, 2017"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
#install.packages("kableExtra")

# For dev version
# install.packages("devtools")
#devtools::install_github("haozhu233/kableExtra")
```

```{r}
library(knitr)
library(kableExtra)
dt <- mtcars[1:5, 1:6]

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

```

options(knitr.table.format = "html") 
## If you don't define format here, you'll need put `format = "html"` in every kable function.

```{r}
kable(dt)

kable(head(MLH1_data) )
```
```{r}
##FIX ORDER HERE
MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
                                     "WSB", "PWD","MSM","CAST"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
#set the order of another column, based on another variable (so that when)
library(dplyr)
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(file.name = factor(file.name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
```

```{r}
#save a figure of boxplots for each mouse
ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )

```

```{r}
#Quality / number of passing mice
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table

passed_mice <- as.data.frame(q_cutoff_table[q_cutoff_table$q_l3 >= 15,]$mouse, 
                             col.names= "mouse")
colnames(passed_mice) <- c("mouse")

passed_mice$strain <- ifelse(grepl("WSB", passed_mice$mouse), "WSB", 
                   ifelse(grepl("G", passed_mice$mouse), "G",
                    ifelse(grepl("CAST", passed_mice$mouse), "CAST",
                   ifelse(grepl("MSM", passed_mice$mouse), "MSM",
                  ifelse(grepl("LEW", passed_mice$mouse), "LEWES", 
                   ifelse(grepl("LEWES", passed_mice$mouse), "LEWES",                                          
                  ifelse(grepl("PWD", passed_mice$mouse), "PWD", "other")))))))

passed_mice$sex <- ifelse(grepl("_f", passed_mice$mouse), "female",                                         
                          ifelse(grepl("_m", passed_mice$mouse), "male", "other"))

#make a table of the number of mice passed for each category.
Pass_mice_category <- ddply(passed_mice, .(strain, sex), summarise,
                    "Passing Mice" =  length(mouse)
)

not_passed_mice <- setdiff(q_cutoff_table$mouse, passed_mice$mouse)
not_passed_mice <- data.frame(not_passed_mice)
colnames(not_passed_mice) <- c("mouse")

```

```{r}
kable(passed_mice)
kable(not_passed_mice)
kable(Pass_mice_category)
```

```{r}
kable(dt) %>%
  kable_styling()
```