### 2017-09-17 rmd_tutorial.rmd
added notes for improving Rmd report

---
title: "kableExtraTutorial"
author: "April Peterson"
date: "August 29, 2017"
output: github_document
---

include is for blocking output messages. echo false is for silencing the code commands.
```{r setup, echo=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)

#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

```

This Rmd file displays the boxplots of MLH1 distributions.

```{r, echo=FALSE}
##FIX ORDER HERE
MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
                                     "WSB", "PWD","MSM","CAST"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
#set the order of another column, based on another variable (so that when)

MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(file.name = factor(file.name)) #another category that you want the order to match
# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
```

Below is a figure of boxplots all mice from current data batches.

```{r, echo=FALSE, fig.width=3.5, fig.height = 2}
#save a figure of boxplots for each mouse
bb <- ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )

bb
```

Below are the boxplots for female MLH1 distributions.

```{r, echo=FALSE}
ff <- ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
ff

```

Below are the boxplots for male MLH1 distributions.

```{r, echo=FALSE}
mm <- ggplot(MLH1_by_M_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
mm
```

Creating Tables with ddply is currently broken Rmd.

```{r}

#q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
 #                       total =  length(nMLH1.foci),
  #                      q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
   #                     q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE)
#)
#q_cutoff_table
```

 figure will be generated below. this is the text above that prefaces it, how and why the figure(s) are set up.
 The following figure will display the boxplots of MLH1 distributions seperated by sexes. 
```{r, MLH1 box plots, fig.width=2, fig.height=3, fig.cap="this is the figure caption"}

```
This is the text that discusses the inferences of the figures.

This is the text above the table, this will show category wide stats of MLH1 counts.
```{r}
#use kable to make tables. warning this might not work for github Rmd.
kable(my table)
```

text before the table for mouse specific tables. (make them smaller and more compact?)
```{r, mouse table}

```

```{r, MxF hetC plot}

```