### 2017-12-07 test4pvalues.Rmd
catching up on updates

---
title: "test_for_p.values"
author: "April Peterson"
date: "December 6, 2017"
output: html_document
---

first the larger data

I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.

```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
#what is the purpose of this block
pMLH1_data <- MLH1_data[MLH1_data$category == "PWD male", ]

#4jan17_LEW_f3, 4jan17_LEW_f6
two_mouse_data <- MLH1_data[ (MLH1_data$mouse == "4jan17_LEW_f3" | MLH1_data$mouse == "4jan17_LEW_f6"), ]

cat_list <- split(pMLH1_data, (pMLH1_data$mouse))
#list of dfs, that are split by mouse
#cat_list$'4jan17_LEW_f3'$quality  AND  cat_list[[1]]$quality
#cat_list[1]$quality

#remove sections which are in category but have no data
cat_list <- cat_list[ -c(7, 12, 14,19, 20) ]
#7: PERC male, 12:CAST female, 14:HMI female, 19:SPIC male

#df_list 
o=0
i=0

#split str.  # PWD - split by mouse,

new_dat = list()
for(i in 1:(length(cat_list))){
  new_dat[i] <- data.frame()
  new_dat[i]$pvals <- replicate(100, t.test(sample(cat_list[[i]]$nMLH1.foci, 5) ,    sample(cat_list[[i]]$nMLH1.foci, 5))$p.value )

new_dat$m <- cat_list[[i]]$mouse 
}

g <- as.data.frame(new_dat)

# ggplot only works with DFs

# g 7 columns, 
plot_list = list()
gcolnames <- colnames(g)
for(h in 1:length(gcolnames)){
  print(h)
  pf <- ggplot(g, aes(x = g[h])) + geom_histogram() + xlim(c(0,1))
  pf
  plot_list[[h]] =   pf
}

for (gg in 1:length(plot_list) ){
  print(gg)
  plot_list[gg][[1]]
  #show(plot_list[gg])
  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
}

```

The above code made a big dataframe with pvalues for small sample sizes

```{r}
plyMLH1_data <- MLH1_data[MLH1_data$mouse == "18may15_PWD_m1", ]

bb <- replicate(1000, t.test(sample(plyMLH1_data$nMLH1.foci, length(plyMLH1_data$nMLH1.foci)/2) ,    sample(plyMLH1_data$nMLH1.foci, 20))$p.value )

ggplot(data.frame(bb), aes(x=bb)) + 
  geom_histogram() #+ geom_tile()

```