### 2017-12-09 test4pvalues.Rmd

---
title: "test_for_p.values"
author: "April Peterson"
date: "December 6, 2017"
output: html_document
---

```{r}
#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
```

first the larger data 
seperate data into >20 cells or less

```{r}
twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,]
lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,]
#these will give the list of mice

#subset MLH1_data by these lists

twenty_data <- MLH1_data[(unique(MLH1_data$mouse) %in% twenty$mouse),]
lessTwenty_data <- MLH1_data[(unique(MLH1_data$mouse) %in% lessTwenty$mouse),]

```
I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.

```{r}

outlist = list()
for(o in 1:length(unique(twenty$mouse))){
  print(twenty$mouse[o])
  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[o],]

  outlist[o] <- list(replicate(100, t.test(sample(mus_data$adj_nMLH1.foci, 15, replace = TRUE),
                        sample(mus_data$adj_nMLH1.foci, 15, replace = TRUE) )$p.value ) )
  }
names(outlist) <- twenty$mouse
#outlist is not a list of p-value list. ()

#find a way to print out histogrms 
plot_list = list()
for (i in 1:length(twenty$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist[[i]])
  colnames(try_df) <- "pvals"
  p = ggplot(try_df) +
        geom_histogram(aes(x=pvals) )#xlim
    plot_list[[i]] = p
}

# Save plots to tiff. Makes a separate file for each plot.
for(i in 1:length(plot_list) ) {
    file_name = paste("histy_", i, ".tiff", sep="")
    tiff(file_name)
    print(plot_list[[i]])
    dev.off()
}

hist(outlist[[5]])
plot()

```

```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
#what is the purpose of this block
pMLH1_data <- MLH1_data[MLH1_data$category == "PWD male", ]

#4jan17_LEW_f3, 4jan17_LEW_f6
two_mouse_data <- MLH1_data[ (MLH1_data$mouse == "4jan17_LEW_f3" | MLH1_data$mouse == "4jan17_LEW_f6"), ]

cat_list <- split(pMLH1_data, (pMLH1_data$mouse))
#list of dfs, that are split by mouse
#cat_list$'4jan17_LEW_f3'$quality  AND  cat_list[[1]]$quality
#cat_list[1]$quality
#remove sections which are in category but have no data

#df_list 
o=0
i=0

#split str.  # PWD - split by mouse,

new_dat = list()
for(i in 1:(length(cat_list))){
  new_dat[i] <- data.frame()
  new_dat[i]$pvals <- replicate(100, t.test(sample(cat_list[[i]]$nMLH1.foci, 5) ,    sample(cat_list[[i]]$nMLH1.foci, 5))$p.value )

new_dat$m <- cat_list[[i]]$mouse 
}

g <- as.data.frame(new_dat)

# ggplot only works with DFs

# g 7 columns, 
plot_list = list()
gcolnames <- colnames(g)
for(h in 1:length(gcolnames)){
  print(h)
  pf <- ggplot(g, aes(x = g[h])) + geom_histogram() + xlim(c(0,1))
  pf
  plot_list[[h]] =   pf
}

for (gg in 1:length(plot_list) ){
  print(gg)
  plot_list[gg][[1]]
  #show(plot_list[gg])
  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
}

```

The above code made a big dataframe with pvalues for small sample sizes

```{r}
plyMLH1_data <- MLH1_data[MLH1_data$mouse == "18may15_PWD_m1", ]

bb <- replicate(1000, t.test(sample(plyMLH1_data$nMLH1.foci, length(plyMLH1_data$nMLH1.foci)/2) ,    sample(plyMLH1_data$nMLH1.foci, 20))$p.value )

ggplot(data.frame(bb), aes(x=bb)) + 
  geom_histogram() #+ geom_tile()

```