### 2017-12-10 test4pvalues.Rmd
working on pval script

---
title: "test_for_p.values"
author: "April Peterson"
date: "December 6, 2017"
output: html_document
---

```{r}
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)
```

first the larger data 
seperate data into >20 cells or less

```{r}
twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,]
lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,]
#these will give the list of mice

#subset MLH1_data by these lists. (fixed line below, )
twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]

twenty_data_table <- ddply(twenty_data, c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         cV = cv(adj_nMLH1.foci),
                          var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
)

lessTwenty_data <- MLH1_data[(unique(MLH1_data$mouse) %in% lessTwenty$mouse),]

#there -- the data counted in for AP_mouse_table doesn't match how 
#my smaller df made for each mouse in the loops.
#the table said Ncells was 21, but the smaller df is only 13 long

```
I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.

```{r}
sim_iterations <- 1000

outlist15 = list()
for(fif in 1:length(unique(twenty$mouse))){
  print(twenty$mouse[fif])
  #
  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[fif],]

  outlist15[fif] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE) )$p.value ) )
  }
names(outlist15) <- twenty$mouse
#outlist is not a list of p-value list. ()

outlist10 = list()
for(ten in 1:length(unique(twenty$mouse))){
  print(twenty$mouse[ten])
  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[ten],]

  outlist10[ten] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE) )$p.value ) )
  }
names(outlist10) <- twenty$mouse

outlist20 = list()
for(tw in 1:length(unique(twenty$mouse))){
  print(twenty$mouse[tw])
  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[tw],]

  outlist20[tw] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE),
                sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE) )$p.value ) )
  }
names(outlist20) <- twenty$mouse
#outlist is not a list of p-value list. ()

plot_list20 = list()
for (i in 1:length(twenty$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist20[[i]])
  colnames(try_df) <- "pvals"
  Title_mus_name <- paste("1000 iterations\n, Test of 2 Samples of 20, \n", names(outlist20[i]) )
  p20 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
    plot_list20[[i]] <- p20
}

plot_list10 = list()
for (i in 1:length(twenty$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist10[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 10, \n", names(outlist10[i]) )
  colnames(try_df) <- "pvals"
  p10 <- ggplot(try_df) + ggtitle( Title_mus_name ) +xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots

    plot_list10[[i]] <- p10
}

plot_list15 = list()
for (i in 1:length(twenty$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist15[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 15, \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"
  p15 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
    plot_list15[[i]] <- p15
}

#make list of scatter plots
sctr_plot = list()
for (i in 1:length(twenty$mouse)){
  print(i)  
  df10 <- as.data.frame(outlist10[[i]])
  df15 <- as.data.frame(outlist15[[i]])
  df20 <-as.data.frame(outlist20[[i]])
  names(df10) = "pvals"
  names(df15)= "pvals"
  names(df20)= "pvals"
  allDF <- data.frame(pvals = rbind(df10, df15, df20),
    DFsets = c (rep("samp10",length(df10[,1]) ), 
              rep("samp15",length(df10[,1])), 
              rep("samp20", length(df10[,1])) )
                      )
    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"

  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
        geom_jitter(aes(x=DFsets, y=pvals) )
    sctr_plot[[i]] <- scplot
}

source("src/MultiPlot.R")

multiplot(plot_list10[[8]],plot_list20[[8]], cols=1)

# Saves multiple plots to a tiff file for each mouse
for(i in 1:5) {
    file_name = paste("Sample_sizes_p_", i, ".tiff", sep="")
    tiff(file_name)
    #par( mfrow=c(1,2) )
    multiplot(sctr_plot[[i]], plot_list10[[i]], plot_list15[[i]], plot_list20[[i]],cols=2)
    dev.off()
}

```

```{r}

#scatterplot of pvales
# the x scales should be the different sample sizes
#so one ggplot per mouse

#mush all of the outlists into 1 DF
#

#make list of plots
sctr_plot = list()
for (i in 1:length(twenty$mouse)){
  print(i)  
  df10 <- as.data.frame(outlist10[[i]])
  df15 <- as.data.frame(outlist15[[i]])
  df20 <-as.data.frame(outlist20[[i]])
  names(df10) = "pvals"
  names(df15)= "pvals"
  names(df20)= "pvals"

  allDF <- data.frame(pvals = rbind(df10, df15, df20),
    DFsets = c (rep("samp10",length(df10[,1]) ), 
              rep("samp15",length(df10[,1])), 
              rep("samp20", length(df10[,1])) )
                      )

    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"

  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
        geom_jitter(aes(x=DFsets, y=pvals) )#xlim
  #tryfiting 3 plots
    sctr_plot[[i]] <- scplot
}

for(i in 1:5) {
    file_name = paste("Scatter_plots_", i, ".tiff", sep="")
    tiff(file_name)
    #par( mfrow=c(1,2) )
    print(sctr_plot[[i]] )
    #multiplot(plot_list10[[i]], plot_list15[[i]], plot_list20[[i]])
    #print(c(plot_list[[i]], plot_list20[[i]]) )

    dev.off()
}

```

```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
#what is the purpose of this block
pMLH1_data <- MLH1_data[MLH1_data$category == "PWD male", ]

#4jan17_LEW_f3, 4jan17_LEW_f6
two_mouse_data <- MLH1_data[ (MLH1_data$mouse == "4jan17_LEW_f3" | MLH1_data$mouse == "4jan17_LEW_f6"), ]

cat_list <- split(pMLH1_data, (pMLH1_data$mouse))
#list of dfs, that are split by mouse
#cat_list$'4jan17_LEW_f3'$quality  AND  cat_list[[1]]$quality
#cat_list[1]$quality
#remove sections which are in category but have no data

#df_list 
o=0
i=0

#split str.  # PWD - split by mouse,

new_dat = list()
for(i in 1:(length(cat_list))){
  new_dat[i] <- data.frame()
  new_dat[i]$pvals <- replicate(100, t.test(sample(cat_list[[i]]$nMLH1.foci, 5) ,    sample(cat_list[[i]]$nMLH1.foci, 5))$p.value )

new_dat$m <- cat_list[[i]]$mouse 
}

g <- as.data.frame(new_dat)

# ggplot only works with DFs

# g 7 columns, 
plot_list = list()
gcolnames <- colnames(g)
for(h in 1:length(gcolnames)){
  print(h)
  pf <- ggplot(g, aes(x = g[h])) + geom_histogram() + xlim(c(0,1))
  pf
  plot_list[[h]] =   pf
}

for (gg in 1:length(plot_list) ){
  print(gg)
  plot_list[gg][[1]]
  #show(plot_list[gg])
  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
}

```

The above code made a big dataframe with pvalues for small sample sizes

```{r}
plyMLH1_data <- MLH1_data[MLH1_data$mouse == "18may15_PWD_m1", ]

bb <- replicate(1000, t.test(sample(plyMLH1_data$nMLH1.foci, length(plyMLH1_data$nMLH1.foci)/2) ,    sample(plyMLH1_data$nMLH1.foci, 20))$p.value )

ggplot(data.frame(bb), aes(x=bb)) + 
  geom_histogram() #+ geom_tile()

```