### 2018-02-26 draft_scripts/pass-_report.Rmd

---
title: "passing mice script"
author: "April Peterson"
date: "February 8, 2018"
output: html_document
---

Guide / report document for showing which mice have passed - sufficient cells samples for showing good estimate of MLH1 count.

Outline
-dicription of graphs / data
-discription of simulations (sampling comparison)
(example images)

<code for making the graphs
-add mouse name, total number of cells,

- develope thresholds for grouping mice into passed and more images
- consider doing this by batch

- based on current plot outputs,

- think of ways to combine this sampling size data with quality distributions
```{r setup, include=FALSE}

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)

```

```{r group_mice}
#seperate large dataframe by how many cells are in the samples
twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,] #53
#are these assumed to have passed?

lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,] #28
#these will give the list of mice

#subset MLH1_data by these lists. (fixed line below, )
twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]
lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]

```

```{r pvalue_permutations, echo=FALSE}

```