### 2018-04-03 draft_scripts/MetaDataEffectReport.Rmd
updating  draft scripts

---
title: "MetaEffectReport"
author: "April Peterson"
date: "March 9, 2018"
output: rmarkdown::github_document
---

```{r load data, include=FALSE}
#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")

#load()metadata
meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1data/data/mouseDOBs.csv")

library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(png)
library(reshape2)
```

### Report for Examining Meta Effects

- age
- mat age
- G line
- female age
- litter

### files

- MLH1 data
- MetaData file

ToDo
- make sure metadata file is as complete as possible
- load and combine data

```{r dates, echo=FALSE}

#extract dates from file names

for( t in 1:length(meta_data$mouse)){
  euth_date = strsplit(meta_data$mouse[t], split="_")[[1]][1]
  meta_data$raw_euth_date[t] <- strsplit(meta_data$mouse[t], split="_")[[1]][1]
  fomt_euth <- as.Date(strsplit(meta_data$mouse[t], split="_")[[1]][1], format= '%d%b%y')
  fomt_euth_nrm <- as.Date(fomt_euth, "%Y-%m-%d")
  vv <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="weeks" )#whoo, this works
  hh <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="hours" )
  meta_data$age_weeks[t] <- vv
  meta_data$age_hours[t] <- hh

  #mat age, DOB -Mat_dob
  mat_dob <- as.Date(meta_data$maternal.age..DOB.[t], format= '%m/%d/%Y')
#  hh <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="hours" )
  gg <- as.numeric(difftime(as.Date(meta_data$DOB[t], '%m/%d/%Y'),mat_dob), units="weeks" )
  meta_data$mat_age_wk[t] <- gg
}

as.Date(first part of mouse,format='%d%b%y' )
euth_date <- as.Date(p[[1]][1],format='%d%b%y' )
as.Date('22JUN01',format='%d%b%y') ##this will read the mouse date format correctly.
#calculate age from dates
difftime()

```

```{r}
#merge data (MLH1 counts and ages)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
load("MLH1_data_setup.RData")
#add euth date and calc age
mergedDF = merge(MLH1_data, meta_data, by = "mouse") #4000 cells by table of 1178 mice

dissected_mice <- unique(meta_data$mouse)#180
imaged_mice <- unique(MLH1_data$mouse)#62
n.overlapping_mice <- length(imaged_mice)/length(dissected_mice)

dfn <- mergedDF[(mergedDF$DOB == '' ),]
dfm <- mergedDF[(is.na(mergedDF$DOB) ), ]

unique(dfn$mouse)

unique((dfm$mouse))

```