### 2018-04-03 draft_scripts/MixedModel_PnD_Report.Rmd

---
title: "SimulatedData_MixedModel"
author: "April Peterson"
date: "March 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(nlme)
library(plyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(raster)#for cV

#load data
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")
```

#### Mixed models

explain mixed model best I can

-- the main point of comparing these mixed models is to --- contrast female and male patterns of variation in MLH1 number--

<previous results show that males had more variaiton at the subspecies level 
< mouse level MLH1

-- redo the previous models -- compare the subsets with high at low quality --

Coding for the mixed model; base Dom male.  (in most of my dataframes female is ordered before males, for visualization purposes)

Displaying the comparisons of different models -- create a table with the ww$values -- so they can be compared..

values
-parameter estimates
-AICs ect

conditions
-quality scores
-batches
-simulated data? (randize strain)

-by sex

```{r, include=FALSE}
#make model

# just look at data of Musc and Dom 
mixmodmouse_data <- AP_mouse_table[(AP_mouse_table$subsp == "Musc" | AP_mouse_table$subsp == "Dom"),  ]

mixmodmouse_data$sex <- factor(mixmodmouse_data$sex, levels = c("male", "female"))

model1 <- lme(mean_co ~ sex * subsp, data=mixmodmouse_data, random=list(strain=pdDiag(~sex)) )

summary(model1)
ranef(model1)
```

To interpert the summary results, (Fixed effects)
- the intercept is the mean MLH1 number of the base factors -- which in this model is defined as Dom male (26.40).

- the next variable is sex being female, has an affect of increasing the mean MLH1 count by 1.7. (This is about the amount of sexual dimorphism)
1.7 (SD)

- the value for Musc is 3.935, meaning being male and Musc adds 3.935 to the mean MLH1 (in the previous model without KAZ, this was 5.5)

- the interaction of sex and subspecies is -3.7 meaning, musc females are 3 MLH1 foci lower than musc males (in previous models this was -5.1)

Random effects (strain and strain specific sexual dimorphism)
- the intercept is the average sd from strains 1.89, 
- the sex effec is the average sd from strain specific sexual dimorphism 2.07
- the random effects can be estimated; the intercepts are the average sd for the strain effect, ... from these esitmates is seems that G and Kaz have strain specific sexual dimorphism where females are larger
... if the strain specific SD effect is larger than the whole sex effect

```{r}
#make new AP mouse table with cells from 1-3
quality_data <- MLH1_data[MLH1_data$quality <= 3, ]
quality_data <- quality_data[(quality_data$subsp == "Musc" | quality_data$subsp == "Dom"),  ]

na.omit(quality_data)

qualAP_mouse_table <- ddply(quality_data, c("subsp", "sex", "strain", "mouse"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                         cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
                        #sex = sex,
                        #subsp = subsp,
                        #strain = strain
                      #quality?

)
qualAP_mouse_table <- na.omit(qualAP_mouse_table)

qualAP_mouse_table <- qualAP_mouse_table[qualAP_mouse_table$Ncells > 5,]

str(qualAP_mouse_table)

qualAP_mouse_table$sex <- factor(qualAP_mouse_table$sex, levels = c("male", "female"))
qualAP_mouse_table$strain <- factor(qualAP_mouse_table$strain)
qualAP_mouse_table$subsp <-  factor(qualAP_mouse_table$subsp, levels=c("Dom", "Musc"))

#model 2 -- asses if there are affects of quality
model2 <- lme(mean_co ~ sex * subsp, data=qualAP_mouse_table, random=list(strain=pdDiag(~sex)) )

summary(model2)
ranef(model2)

```

For this version of cleaned data, the effect of Musc on the mean MLH1 is lower.