### 2019-05-10 doc/pass_report.Rmd

---
title: "passing mice script"
author: "April Peterson"
date: `r format(Sys.time(), ''%d %B, %Y'')`
output: rmarkdown::github_document
---

Guide / report document for showing which mice have passed - sufficient cells samples for showing good estimate of MLH1 count.

Outline
-dicription of graphs / data
-discription of simulations (sampling comparison)
(example images)

<code for making the graphs
-add mouse name, total number of cells,

- develope thresholds for grouping mice into passed and more images
- consider doing this by batch

- based on current plot outputs,

- think of ways to combine this sampling size data with quality distributions
```{r setup, include=FALSE}

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)
```

```{r group_mice}
#seperate large dataframe by how many cells are in the samples
twenty <- AP_mouse_table[AP_mouse_table$Ncells > 20,] #53
#are these assumed to have passed?

lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,] #28
#these will give the list of mice

#subset MLH1_data by these lists. (fixed line below, )
twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]
lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]

```

### Power and sampling size effects
Create dataframes that iterate through different sample sizes of MLH1 counts 
 Outlists are dataframes, filled with pvalues from t-tests of different sample sizes.
 Q: How do the 20cells or greater mice observations compare with the less than 20 cell mouse observations.

Passing cells have a small effect from different sampling sizes.

```{r pvalue_permutations, echo=FALSE}
sim_iterations <- 10000

#list, DF, ncells, sim_num

source("src/Func_sampleCells.R")
lessTwenty_pvals_5_10 <- sample_cells_pvals(lessTwenty, lessTwenty_data, 5, 10)#this returns a list (of vectors?) list of pvalues for each mouse in a list

#nn[[4]] is the syntax to get just the pvalues
#names(nn[4]) is now you get the name of the vector / list

#sample >twenty mice

#REDO these, I think the function was wrong
Twenty_pvals_10_1k <- sample_cells_pvals(twenty, twenty_data, 10, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
Twenty_pvals_15_1k <- sample_cells_pvals(twenty, twenty_data, 15, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
Twenty_pvals_20_1k <- sample_cells_pvals(twenty, twenty_data, 20, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list

#decide on threshold for difference in pvalues

#below doesn't work
#guess <- Twenty_pvals_15_1k - Twenty_pvals_10_1k

guess <- mean(Twenty_pvals_15_1k[[1]]) - mean(Twenty_pvals_10_1k[[1]])

#make blank df
mean1<- vector()
mean2<- vector()
mean3<- vector()

onelist <- vector()
twolist <- vector()
threelist <- vector()
for(l in 1:length(Twenty_pvals_10_1k)){
  #var doesn't seem to be good
  mean1 <- mean(Twenty_pvals_10_1k[[l]])
  mean2 <- mean(Twenty_pvals_15_1k[[l]])
  mean3 <- mean(Twenty_pvals_20_1k[[l]])
  first <- ( mean(Twenty_pvals_10_1k[[l]]) - mean(Twenty_pvals_15_1k[[l]]) )
  sec <- mean(Twenty_pvals_15_1k[[l]]) - mean(Twenty_pvals_20_1k[[l]])
  thrid <- mean(Twenty_pvals_10_1k[[l]]) - mean(Twenty_pvals_20_1k[[l]])

  mean1.list[l] <- mean1
  mean2.list[l] <- mean2
  mean3.list[l] <- mean3

   onelist[l] <- first
  twolist[l] <- sec
  threelist[l] <- thrid
}

comp.pvals <- data.frame(mean1=mean1.list, mean2=mean2.list, mean3=mean3.list, dif1 = onelist, dif2 = twolist, dif3=threelist)
row.names(comp.pvals) <- names(Twenty_pvals_10_1k)

#comp.pvals now is a df that 
#rowSums(comp.pvals) 
ll <- rowSums(comp.pvals)

#this excludes ... mice/rows in data frame which have greater differences 
#between 15 and 20 samples
mm <- subset(comp.pvals, comp.pvals$dif2 >= -.05)

#some bad looking mice still get through (is this due to sample size)

# it seems like mice with between 20-25 cells, there is a large difference
# between the p values of 15 and 20

```

what does it mean that I excldude these mice?
I will infer differen mean MLH1 values due to sample size with mice with

10,15,20, might not be the best for mice with around 20 cells

less than 20
```{r}

lessTwenty <- subset(AP_mouse_table, Ncells >15 & Ncells <20)
lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]

source("src/Func_sampleCells.R")
lessTwenty_pvals_8_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 8, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
lessTwenty_pvals_11_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 11, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
lessTwenty_pvals_15_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 15, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list

#make blank df
mean1.list<- vector()
mean2.list<- vector()
mean3.list<- vector()

onelist <- vector()
twolist <- vector()
threelist <- vector()
for(l in 1:length(lessTwenty_pvals_8_1k)){
  #var doesn't seem to be good
  mean1 <- mean(lessTwenty_pvals_8_1k[[l]])
  mean2 <- mean(lessTwenty_pvals_11_1k[[l]])
  mean3 <- mean(lessTwenty_pvals_15_1k[[l]])

  first <- ( mean(lessTwenty_pvals_8_1k[[l]]) - mean(lessTwenty_pvals_11_1k[[l]]) )
  sec <- mean(lessTwenty_pvals_11_1k[[l]]) - mean(lessTwenty_pvals_15_1k[[l]])
  thrid <- mean(lessTwenty_pvals_8_1k[[l]]) - mean(lessTwenty_pvals_15_1k[[l]])

  mean1.list[l] <- mean1
  mean2.list[l] <- mean2
  mean3.list[l] <- mean3

   onelist[l] <- first
  twolist[l] <- sec
  threelist[l] <- thrid
}

LTcomp.pvals <- data.frame(mean1=mean1.list, mean2=mean2.list, mean3=mean3.list, dif1 = onelist, dif2 = twolist, dif3=threelist)
row.names(LTcomp.pvals) <- names(lessTwenty_pvals_8_1k)

#dif2 cut off -0.10, greater than -.1 mice should pass even though they have less than 20

pass_LessTwenty <- row.names(LTcomp.pvals[(LTcomp.pvals$dif2 > -.1),])

```

## what about the effects of quality on sample size effects