### 2019-05-10 doc/strain_reports/AST_Report.Rmd

---
title: "AST_Report"
author: "April Peterson"
date: "January 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(plyr)
library(dplyr)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#MLH1_data = read.csv("data/AnonData.csv", header=TRUE )
load(file="~./MLH1repo/MLH1_data_setup.RData")

#input
#output: Rmd file that can be read on github

knitr::opts_chunk$set(echo = TRUE)

mouse_list = read.csv("~./MLH1repo/data/MLH1q_AST.csv", header=TRUE )
# this is the list of total mice
#load big MLH1 Rdata file

#format the dates into standard format
mouse_list$DOB <- as.Date(mouse_list$DOB,format='%m/%d/%Y')

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
source("~./MLH1repo/src/CommonFunc.R")#this works
#source("src/Func_addCategory.R")

mouse_list <- add_mouse1(mouse_list)
mouse_list <- add_strain(mouse_list)
#source("src/Func_addSex.R")
mouse_list <- add_sex(mouse_list)
mouse_list <- add_category(mouse_list)
#set the order of categories (female, male) (cast, dom, musc)
#source("src/Func_addStrain.R")
mouse_list <- add_euth_date(mouse_list)
mouse_list <- add_age(mouse_list)

```

```{r}
#make a tables, sex, and ages
#count mice by categories of sex and ages

#ages in days -- 
#5 to 15 weeks
#35 days,  105days

#don't go row by row, for each mouse, 
#summarize the whole dataset...
#make new dataframe
mouse.report <- data.frame("total.Nmice" = length(unique(mouse_list$mouse)), 
      "Nmales"= length(mouse_list$mouse[mouse_list$sex == "male"]),
      "Nfemales"= length(mouse_list$mouse[mouse_list$sex == "female"]),
      "passed.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse)

     )
#imaged mice...

#sum( mouse_list$mouse  %in%  MLH1_data$mouse)
#ifelse(mouse_list$mouse %in% MLH1_data$mouse, "1", "")

```

```{r BasicStats, echo=FALSE}
#how many euthed?
#how many mice MLH1 stained and imaged?
#what is relevant meta data?

#how many cells?

#draw info from anonMLH1 data -- MLH1 rate, quality,

```

```{r}
#print out plots

```