### 2019-05-10 draft_scripts/pass_report.Rmd

---
title: "Sub-Sampling Effect"
author: "April Peterson"
date: "June 4, 2018"
output: rmarkdown::github_document
---

#### Past certeria for mouse level

BD: Approximately 20 cells were scored per animal. We were unable to obtain a sufficient number of high quality images for 39 of the 315 F2 animals.

RW: The average individual was represented by counts from >20 spermatocytes; we omitted individuals with counts from <5 spermatocytes.

AP: >20 cells with 3 highest quality (1,2,3).

Guide / report document for showing which mice have passed - sufficient cells samples for showing good estimate of MLH1 count.
Compare re-sampling simulations for 'Hard Pass / Strict Threshold' data to on the edge data, so assess if there is a large effect on mouse-MLH1 mean for cell sample size.

Maybe female and males will have different effects.

```{r "setup", include=FALSE}
require("knitr")
#opts_knit$set(root.dir = "~/gits/project/")

opts_knit$set(root.dir = "C:/Users/alpeterson7/Documents/MLH1repo")
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)

```

### Passing Mouse list

```{r mouse_lists, include=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above4 = sum(as.numeric(quality) <= 3, na.rm = TRUE ),
                        #add percentages
                        prop.lt5 = above4/total
)
length(q_cutoff_table$mouse)#116

pass_mus_table <- q_cutoff_table[q_cutoff_table$above4 >= 25,] #30
# boarder line
boarder_mice_list <- q_cutoff_table[ ( q_cutoff_table$prop.lt5 >= .6) & (q_cutoff_table$above4 < 20), ] #39

```

Mice with slides with over 25 cells of good quality: `r length(pass_mus_table$mouse)`
Mice with fewer than 20 cells, but have high proportion of good quality cells (>60%): `r length(boarder_mice_list$mouse)`.

What is the distribution of total cell counts for these? For the boarderline mice, total is around ~20 cells, ~15 cells above 4 quality.

Run sub-sampling simulations for data of all these mice ... This requires adjusting the sample sizes, over half the mice have 15 or more.

The chunks with the code for the sub-sampling simulations are skipped.

```{r pvalue_permutations HQ, echo=FALSE,include=FALSE, eval = FALSE}
HQ <- pass_mus_table
sim_iterations <- 1000

outlist15 = list()
for(fif in 1:length(unique(HQ$mouse))){
  #print(HQ$mouse[fif])
  mus_data <- MLH1_data[MLH1_data$mouse == HQ$mouse[fif],] #... make sure this is right

  outlist15[fif] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE) )$p.value ) )
  }
names(outlist15) <- HQ$mouse
#outlist is not a list of p-value list. ()

outlist10 = list()
for(ten in 1:length(unique(HQ$mouse))){
 # print(HQ$mouse[ten])
  mus_data <- MLH1_data[MLH1_data$mouse == HQ$mouse[ten],]
  outlist10[ten] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE) )$p.value ) )
  }
names(outlist10) <- HQ$mouse

outlist20 = list()
for(tw in 1:length(unique(HQ$mouse))){
 # print(HQ$mouse[tw])
  mus_data <- MLH1_data[MLH1_data$mouse == HQ$mouse[tw],]
  outlist20[tw] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE),
                sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE) )$p.value ) )
  }
names(outlist20) <- HQ$mouse
#outlist is not a list of p-value list. ()

#list for n=20 sample simulations
plot_list20 = list()
for (i in 1:length(HQ$mouse)){
#  print(i)  
  try_df <- as.data.frame(outlist20[[i]])
  colnames(try_df) <- "pvals"
  Title_mus_name <- paste("1000 iterations\n, Test of 2 Samples of 20, \n", names(outlist20[i]) )
  p20 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
    plot_list20[[i]] <- p20
}
#list for n=10 sample simulations
plot_list10 = list()
for (i in 1:length(HQ$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist10[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 10, \n", names(outlist10[i]) )
  colnames(try_df) <- "pvals"
  p10 <- ggplot(try_df) + ggtitle( Title_mus_name ) +xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
  plot_list10[[i]] <- p10
}

#list for n=15 sample simulations
plot_list15 = list()
for (i in 1:length(HQ$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist15[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 15, \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"
  p15 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
    plot_list15[[i]] <- p15
}

#make list of scatter plots
sctr_plot = list()
for (i in 1:length(HQ$mouse)){
  print(i)  
  df10 <- as.data.frame(outlist10[[i]])
  df15 <- as.data.frame(outlist15[[i]])
  df20 <-as.data.frame(outlist20[[i]])
  names(df10) = "pvals"
  names(df15)= "pvals"
  names(df20)= "pvals"
  allDF <- data.frame(pvals = rbind(df10, df15, df20),
    DFsets = c (rep("samp10",length(df10[,1]) ), 
              rep("samp15",length(df10[,1])), 
              rep("samp20", length(df10[,1])) )
                      )
    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"

  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
        geom_jitter(aes(x=DFsets, y=pvals) )
    sctr_plot[[i]] <- scplot
}

source("src/MultiPlot.R")
#multiplot(plot_list10[[8]],plot_list20[[8]], cols=1)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/outside_gitrepo/sampling_effect")

# Saves multiple plots to a tiff file for each mouse
for(i in 1:length(HQ$mouse)) {  #replace counter with mouse name
    file_name = paste("HQ_Sample_sizes_p_", HQ$mouse[i], ".png", sep="")
    #tiff(file_name)
    png(filename = file_name)
    #par( mfrow=c(1,2) )
    multiplot(sctr_plot[[i]], plot_list10[[i]], plot_list15[[i]], plot_list20[[i]],cols=2)
    dev.off()
}
```

eval = FALSE
-- errors at this section likely due to changes in working directories
```{r pvalue_permutations Boarderline, echo=FALSE, include=FALSE}
#HQ <- pass_mus_table
BL <- boarder_mice_list[boarder_mice_list$total >= 15,]
sim_iterations <- 1000

### adjust sampling sizes (limit this list to mice with )
# max 15,   subampling 12,10,8

outlist12 = list()
for(fif in 1:length(unique(BL$mouse))){
 # print(BL$mouse[fif])
  mus_data <- MLH1_data[MLH1_data$mouse == BL$mouse[fif],] #... make sure this is right

  outlist12[fif] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 12, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 12, replace = FALSE) )$p.value ) )
  }
names(outlist12) <- BL$mouse
#outlist is not a list of p-value list. ()

outlist10 = list()
for(ten in 1:length(unique(BL$mouse))){
  #print(BL$mouse[ten])
  mus_data <- MLH1_data[MLH1_data$mouse == BL$mouse[ten],]
  outlist10[ten] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE) )$p.value ) )
  }
names(outlist10) <- BL$mouse

outlist8 = list()
for(tw in 1:length(unique(BL$mouse))){
  #print(BL$mouse[tw])
  mus_data <- MLH1_data[MLH1_data$mouse == BL$mouse[tw],]
  outlist8[tw] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 8, replace = FALSE),
                sample(mus_data$adj_nMLH1.foci, 8, replace = FALSE) )$p.value ) )
  }
names(outlist8) <- BL$mouse
#outlist is not a list of p-value list. ()

#list for n=8 sample simulations
plot_list8 = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  try_df <- as.data.frame(outlist8[[i]])
  colnames(try_df) <- "pvals"
  Title_mus_name <- paste("1000 iterations\n, Test of 2 Samples of 8, \n", names(outlist8[i]) )
  p8 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
    plot_list8[[i]] <- p8
}
#list for n=10 sample simulations
plot_list10 = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  try_df <- as.data.frame(outlist10[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 10, \n", names(outlist10[i]) )
  colnames(try_df) <- "pvals"
  p10 <- ggplot(try_df) + ggtitle( Title_mus_name ) +xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
  plot_list10[[i]] <- p10
}

#list for n=12 sample simulations
plot_list12 = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  try_df <- as.data.frame(outlist12[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 12, \n", names(outlist12[i]) )
  colnames(try_df) <- "pvals"
  p12 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
    plot_list12[[i]] <- p12
}

#make list of scatter plots
sctr_plot = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  df12 <- as.data.frame(outlist12[[i]])
  df10 <- as.data.frame(outlist10[[i]])
  df8 <-as.data.frame(outlist8[[i]])
  names(df12) = "pvals"
  names(df10)= "pvals"
  names(df8)= "pvals"
  allDF <- data.frame(pvals = rbind(df8,df10, df12),
             DFsets = c( rep("samp8",length(df8[,1]) ),
                         rep("samp10",length(df10[,1]) ),
                         rep("samp12", length(df12[,1]) )
                      )
    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist12[i]) )
  colnames(try_df) <- "pvals"

  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
        geom_jitter(aes(x=DFsets, y=pvals) )
    sctr_plot[[i]] <- scplot
}

#source("src/MultiPlot.R")
#multiplot(plot_list10[[8]],plot_list12[[8]], cols=1)  #forget what this is for

setwd("./outside_gitrepo/sampling_effect") #the loop is not printing in the folders I want

# Saves multiple plots to a tiff file for each mouse
for(i in 1:length(BL$mouse)) {  #replace counter with mouse name
  file_name = paste("BL_Sample_sizes_p_", BL$mouse[i], ".png", sep="")
  #tiff(file_name)
  png(filename = file_name)
  #par( mfrow=c(1,2) )
  multiplot(sctr_plot[[i]], plot_list8[[i]], plot_list10[[i]], plot_list12[[i]],cols=2)
  dev.off()
}
```

```{r}
#![text](./outside_gitrepo/sampling_effect/BL_Sample_sizes_p_4dec17_HMI_m1.png)

#setwd("./outside_gitrepo/sampling_effect")
```

![text](C:/Users/alpeterson7/Documents/MLH1repo/outside_gitrepo/sampling_effect/BL_Sample_sizes_p_4dec17_HMI_m1.png)

![text](C:/Users/alpeterson7/Documents/MLH1repo/outside_gitrepo/sampling_effect/BL_Sample_sizes_p_4dec17_HMI_m1.png)

From comparing the sub-sampling plots from High quality (HQ) and Boarder line (BL),

7 mice out of 29 boardline mice, looked like there was a sampling effect. I will prioritie re-imaging those, and start re-imaging the rest with fewer than 20 cells.