### 2019-06-02 results/report.rmd
working on missing list

---
title: "MLH1_BivData_Report"
author: "April Peterson"
date: "May 30, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#set dir

#load data

```

goals for BivData Report

- count the number of observations from all categories
- identify best mice to make the subset for curated data

plots by categories. (which have longer SCs?)

```{r}

```