### 2019-06-04 results/PWD.KAZ.BivData.Rmd
started Rmd file for reporting/comparing PWD KAZ metrics

---
title: "PWD.KAZ"
author: "April Peterson"
date: "June 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load BivData

#load MLH1 data

```

## Background and Premise

Comparison of PWD and KAZ Biv Data

PWD blah

CZECHII blah

KAZ blah

```{r pressure, echo=FALSE}

#MLH1 difference

```

These are the MLH1 counts

### BivData dataset

Goal:  Count the Bivdata for pwd and KAZ males

start determining the subset of data

```{r}

```