### 2019-06-11 results/PWD.KAZ.BivData.Rmd
tried starting initial analysis of KAZ and PWD data, addressing bugs and found BivDataSet is incomplete

---
title: "PWD.KAZ"
author: "April Peterson"
date: "June 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~./MLH1repo/")
#load BivData
load(file="~./MLH1repo/data/MLH1_BivData.RData")

#load MLH1 data
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="~./MLH1repo/data/MLH1/MLH1_data_setup.RData")

```

## Background and Premise

Comparison of PWD and KAZ Biv Data

PWD blah

CZECHII blah

KAZ blah

```{r set.up.data, echo=FALSE}

#extract just the relavant PWD and KAZ
PWD.MLH1 <- MLH1_data[MLH1_data$category == "PWD male",]
KAZ.MLH1 <- MLH1_data[MLH1_data$category == "KAZ male",]

PWD.BivData.cells <- BivData_cells[BivData_cells$category == "PWD male",]
PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]

KAZ.BivData.cells <- BivData_cells[BivData_cells$category == "KAZ male",]
KAZ.BivData <- fullBivData[fullBivData$category == "KAZ male",]

PWD.KAZ_MLH1<- rbind(PWD.MLH1, KAZ.MLH1)
PWD.KAZ_BivData<- rbind(PWD.BivData, KAZ.BivData)

```

These are the MLH1 counts

### BivData dataset

Goal:  Count the Bivdata for pwd and KAZ males
start determining the subset of data

```{r mergeing.meta.data, echo=FALSE}
meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375

#find a way to remove the rows without the right format before setting as as.Date
#trying to remove these the extra rows..uggh
#yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]

#yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
#yy <- yy[subset(yy$DOB, !(yy$DOB == "e1")),]
#i think subseting is messing up the dataframes
meta.data.file99 <- meta.data.file2
meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#meta.data.file2 <- add_euth_date(meta.data.file2) #euthdate 1st
#meta.data.file2 <- add_age(meta.data.file2)  #age second
##got it to work ... maybe should save adding it to the merged DF

#reduce metadata down to 2 cols (mouse, DOB), merge this with MLH1
MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]

PWD.KAZ_MLH1.imprv <- merge(PWD.KAZ_MLH1, MouseMetaData.wrkin)#don't apply the all parameters
#report number with  and without

#add euth date for calculating age
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
PWD.KAZ_MLH1.imprv <- add_euth_date(PWD.KAZ_MLH1.imprv)
PWD.KAZ_MLH1.imprv <- add_age(PWD.KAZ_MLH1.imprv)

#yesDOB.check <- PWD.KAZ_MLH1.imprv[!(is.na(PWD.KAZ_MLH1.imprv$DOB)),]
#noDOB.check <- PWD.KAZ_MLH1.imprv[(is.na(PWD.KAZ_MLH1.imprv$DOB)),]#note, date type of data can't do the string matching!

#total, KAZ 144, PWD 139
#table(noDOB.check$category)#  no DOB, KAZ 67  PWD 133
#table(yesDOB.check$category) #yes DOB, KAZ 77  PWD 6

# I need to find the ages of more PWD males

```

```{r}
#BivData

MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]

PWD.KAZ_BivData.imprv <- merge(PWD.KAZ_BivData, MouseMetaData.wrkin)

table(fullBivData$category)# PWD 670, 
#there are no KAZ male data

```

```{r}
#distribution of ages

histys <- ggplot(PWD.KAZ_MLH1.imprv, aes(age.weeks, fill=strain))+geom_density()

```