### 2019-08-30 doc/Main_Report.rmd
updating list of mice to finish

---
title: "MLH1 Data Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
---

```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")

#load(file="data/MetaData.RData") #this was replacing most updated data file

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

# Research Description

Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice abd related rodents. Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.

## Description of Report

- Show basics of the current data file

- Display plots and figures of distributions of MLH1 counts to assess intial overall patterns.

-  Report basic statistics of results and to help keep track of the number of images for each category and mouse

*Main questions*

1. Which strains do I have enough mouse specific MLH1 data for?
2. What data will go into the FINAL dataset

## Discription of quantification process

Cell images are quantified in batches after file names are anonymized. The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified. A quality score is given; 1 to 5 (best to worst).

all cell image file names are anonimized and moved into bathces of a few hundred images.

Definition of passed quantified ect.

```{r organizing.data, echo=FALSE}
#set the order of another column, based on another variable (so that when)

MLH1_data <- MLH1_data %>%
  arrange(strain, sex, category, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
#use the above line for when order is out of wak

#count the non-quality measures,  #remove non qualit
#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]

#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]

#MLH1_data <- MLH1_data %>%
#  arrange(category) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])

```

*What were guidlines for quantification of lower quality cells / faint foci?*

-Count foci that look like other foci in the image.

## Missing Mice, finishing dataset

*Which mice are remaining and at what stage?*

This section shows / displays / catches mice which haven't been stained or quantified yet.

```{r missing.mice, echo=FALSE}

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
setwd("~./MLH1repo/")
#other lists

Image_dirs <- list.files(path = "C:/Users/alpeterson7/Documents/Images")
#write this out so that I can access it

#remove the non-mouse folders
Dissection.File = read.csv("~./MLH1repo/data/Disections_53019.csv", header = TRUE)
Dissection.File <- Dissection.File[!(is.na(Dissection.File$mouse)|Dissection.File$mouse==""),]

#print the name with the date of the file somewhere
Dissection.list <- Dissection.File$mouse
Dissection.list <- as.character(Dissection.list)

#load metadata .. start decerning which mice are missing information... and which should be used
Meta.Data.File = read.csv("~./MLH1repo/data/ALP_MouseMetadata.csv", header = TRUE)

#remove file name, and original name .. this is just a mouse level DF
Meta.Data.File <- Meta.Data.File[!(is.na(Meta.Data.File$mouse)|Meta.Data.File$mouse==""),]
#make sure that the DOB is a date

meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61019.csv", header = TRUE)
meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%y")

jl <- unique(Meta.Data.File$mouse[(Meta.Data.File$mouse %in% Dissection.File$mouse)])
BD.w.meta <- Meta.Data.File[(Meta.Data.File$mouse %in% Dissection.File$mouse), ]

#figure out a way to incorporate the staining... to get a better idea of which slides 
#to do next, #remove the pero mice?

#length(unique(MLH1_data$mouse) )

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
original_DF$fileName <- (original_DF$Original.Name)
original_DF <- add_mouse(original_DF)
#length(unique(original_DF$mouse) )
#131 mice have been quant total, 124 mice have passed into the final version

# make a list for the failed slides

#Dissection.list - Image/dirs
NonImaged.mice <- unique(Dissection.File$mouse[(Dissection.File$mouse %in% Image_dirs)])
#231 mice dissected which need images

#ImageFolder - MLH1.Quant
NonQuant.mice <- unique(Image_dirs[(Image_dirs %in% MLH1_data$mouse )])
#length(NonQuant.mice)
#123 mice that need to get into anon-batch

#make Dataframe to put into a table
mouse.number.DF <- data.frame( category = 
              c("Passed Quantified", "Total Quantified", "Non-Quantified", "Non-Imaged"),
              number.of.mice = c(length(unique(MLH1_data$mouse)),
                length(unique(original_DF$mouse) ),
                        length(NonQuant.mice),
                         length(NonImaged.mice)  )  )

kable(mouse.number.DF)
```

passed quantified - quantified mice (with enough cells)
total quantified -- all mice which have been (includes fail and passing mice)

non-quantified mice -- dissected mice / slides that haven't been quantified yet (all stained, but without image directory)

non-imaged -- mice without images, (might be stained and unstained)

This is the number of mice that are through different parts of the pathway or quantification process.

should I break down the categories into sex and strain?

```{r basic.numbers, echo=FALSE, eval=FALSE}

kable(table(AP_mouse_table$category), col.names = c("Category", "mice"), label = "Number of mice by category" )

kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )

```

These are the tables for the number of cells and mice which have been quantified.
The 'Other' are F1 crosses.

# Tests of dataset

Main metric is mean CO count per mouse. Assess the quality of this data with the following tests

*age effects* (male, maternal age)

*litter effect*

*mouse room effect* (Charmony, MSC)

*quantification repeatability*

*spread batch and staining effects*

*power analysis on number of cells*

*time of year effect (on spread)*

These aspects will be analyzed with a linear models on subsets of data (within category)

## Repeatability of Human quantification

*What is the Error in quantification across humans?*

A general ancedote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (useing some of Beth's data)

There is a subset of data which has been counted mulitple times -- use that to test the repeatbility.

```{r quant.repeatability, echo=FALSE}

```

## Age Effect

This section is meant to double check the age distributions of male mice.

```{r age.hists, echo=FALSE, eval=FALSE}

meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#make Mouse meta data work
MouseMetaData.wrkin <- meta.data[,c(1,4)]
colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")

MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]

#do this for the mouse table instead of the MLH1 table
MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters

#add euth date for calculating age

MLH1.imprv <- add_euth_date(MLH1.imprv)
#
MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
MLH1.imprv$euth.date <- as.Date(as.character(MLH1.imprv$euth.date2), format='%d%b%y')
#this is currently broken
MLH1.imprv <- add_age(MLH1.imprv)

#remove mice that are older than 20 weeks
MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]

pp <- ggplot(MLH1.imprv.male, aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Age Distributions")
pp

#what is this meta.data.file?
#meta.data.file2 <- add_strain(meta.data.file2)
#meta.data.file2 <- add_sex(meta.data.file2)
#meta.data.file2 <- add_category(meta.data.file2)
#meta.data.file2 <- add_euth_date(meta.data.file2)
#meta.data.file2 <- add_age(meta.data.file2)

#qq <- ggplot(meta.data.file2[meta.data.file2$sex == "male",], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("recorded ages of dissected mice")
#qq

```

 Is there an age effect?

make a subetted dataset for ploting mean mouse (quality < 5) MLH1 count and age.  Mice with enough ages:
WSB, KAZ, LEW, G, Cast, SPIC, MSM, MOLF, CZECH

```{r age.effect.model}
#run a age effect linear regression!

male.data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.y == "male",]

lm.me  <- lm(male.data$mean_co ~ male.data$age.weeks+male.data$strain.x)

summary(lm.me)

```

```{r age.effect.plot}

#make sure the right dataset 
male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
#remove NAs

male.data <- male.data[male.data$Ncells > 10,]

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(15,39))+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age Effect on CO mean")+facet_wrap(~subsp.y)
age.effect.plot
#todo add line to strain.y

```

Given this plot it seems that there isn't a age effect. I should double check some of these mice

#### Materinal Age effect

```{r mat.age.effect, echo=FALSE}
#compile data from female (and male?) mice with maternal ages.

#start making a general linear model...? for mouse averages

```

### litter effect?

## Spread and Stain Effects

*month dissected*

*staining batches*

## Room effect (envioronment)

3 different rooms used to house mice. This will be tested in X manner.

## cell number (power analysis results)

# Initial Results

```{r first.boxplots, echo=FALSE, fig.align='center'}

#the below ordering works!! keep it
MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
              "PWD","MSM","MOLF","CZECH","SKIVE",
              "KAZ","TOM","AST",
              "CAST","HMI","SPRET","SPIC", "other"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

#don't know why this won't do the right order...
#set the order of another column, based on another variable (so that when)

#try to make sure sacles are the same

#reduncant
#MLH1_data <- MLH1_data %>%
 # arrange(strain, sex, mouse) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
# sort your dataframe, by the focal categories
#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

#try this for ordering
#MLH1_data[order(MLH1_data$strain), ]

#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
#can't re-order the miceagg
#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
#        "grey","grey","grey",  "grey","grey","grey" )) +
# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
#don't show M.musculus by mouse

#order MLH1_data by strain!!

ff2 <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
  geom_boxplot(aes(x = as.character(mouse), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
        "green","green4","grey",  "grey","grey","grey" )) +
 ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
  facet_grid(subsp~ sex, scales="free_x")#scales="free_y"
#this is ok except for the set placement of mice on x axis
#if the full dataset was ordered correctly, this would kinda fix it

#boxplots by strain
ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey","green", "purple","grey1","grey","white", "blue","orange" )) +
  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )

ss <- ss   + facet_wrap(~ sex, scales="free")

#these plot currently look bad
ss
```

*What is the initial picture of Heterochiasmy across mouse strains?*

For the strain level plots
  -facet,sex by species

### Distributions of MLH1 counts

*Are the MLH1 distributions close to normal?*

```{r, MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
#consider making distribution with quality considered...

gg <- data.frame(weight=rep(x,nrow(df)),df)
gg$y <- with(gg,dnorm(x,mean,sd))
gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * diff(range(weight$weight))/30

mean.MLH1.obs <- ddply(.data=MLH1_data, 
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")+
  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))

  geom_smooth(method = lm, se = FALSE)

  geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

histM

#if i'm testing normalge

#function = glm(y ~ x, family = poisson)
#add Poission curve behind
#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)

#not sure how to add Expected poisson curve

```

(most all distributions are close to normal)

How do these compare to Poission? (mean == variance)
I should make 'expected' distributions behind these.

### Assesment of Quality Scores

> *How do the distributions change across different quality scores?*

> *How do the distributions change across mice with a good number of cells?*

```{r high.qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}

#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )

#are there no KAZ cells of high quality?
qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
```

```{r, passing.mice, echo=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table
passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)

not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )

#mnake dataframe with just passing mice -- so that it can be ploted
passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
```

*Answer to questions*

Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality. Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1. (CAST female data is fake)

```{r, scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means across quality scores"}
#make lattice / plots seperated by strain and sex
#facet

#quality scores are now 1:7?
qual_means <- MLH1_data %>% 
        group_by(category, quality) %>% 
        summarise(
          adj_nMLH1.foci = mean(adj_nMLH1.foci)
        )

p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
p <- p + geom_point(stat = "identity")
p <- p  + geom_point(data = qual_means, colour="red") 
p <- p + facet_wrap(~ category)
p

```

I ploted the mean of each quality bin with a red dot. From the pattern of the red dots, there is definately a negative relationship with quality and nMLH1 foci across the data. This is most pronounced in MSM males and least pronounced in G males. The CAST female data is dummy data.

####Distributions by mouse

*Is there mouse level MLH1 quality score bias?*

latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.

below are code chuncks that make a big dfs of CO.count by quality score plot. The code has been commented out, and moved to it's own doc.

```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
### to integrate 'pass'/fail
### this should be in setup data

#mice that don't have female data will break the chain-loop
#MOLF female, HMI, CZECH female, CAROLI female, Tom male

#instead of adding fake data, remove the categories from list

#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing

pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")

#this loop makes a quant_status col for a list of mice
for(i in 1:length(MLH1_data$mouse)){
  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
}

df_list <- split(MLH1_data, (MLH1_data$category))
#remove sections which are in category but have no data
#this is where you remove the non-data data!
df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]

#7 PERC, 12 MOLF female, 
#

#df list, is a list of df's which were split() from BIG df by category
df_list 
o=0
i=0

MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)

#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
for(i in 1:(length(df_list))){
 # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
                      mouse_nMLH1 = mean(adj_nMLH1.foci)
  )
# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
  for(o in 1:length(df_list[[i]]$Original.Name)){
    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
        }
}

plot_list = list()
for (h in 1:length(df_list)) {
  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
    facet_wrap(~ mouse, drop = TRUE) + 
    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
  #make an empty ggplot? for missing data

  plot_list[[h]] =   pf
}

#orig (I think this is the same as above)
#plot_list = list()
#for (h in 1:length(df_list)) {
#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)

#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +

 #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +

    #theme for indiviual mice, 
#theme(plot.background = element_rect(fill = "green"))
#  plot_list[[h]] =   pf
#}
#can't change the alpha/transparency of the red rec

#aes must be same length ---   #geom_point(stat = "identity") +
# the last one is correct....?

#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
```

```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
geom_hline(aes(yintercept = med, group = gr), colour = 'red')
#no plot is made that goes in the list -- so it can't be shown
for (gg in 1:length(plot_list) ){
  show(plot_list[gg])
 # invisible(lapply(obj, function(x) plot(x,main="some plot")))
}

#this stops at MSM
```

The mouse specific scatter plots aren't show because there are too many. (but they should be used for investigating specific mice).
Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse. The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.

PWD females and a very large range. This is driven by data from 17apr15_PWD_f4. These data should be looked at more closely.

In some mice there seems to be bimodel cell populations, those of clearly high quality and those of lower quality. This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.

# Research Questions

*-which strain has highest HetC?*

Start thinking of the direction for analysis

## within mouse

### female and male comparison

## within category (strain)

```{r HetC.plot, echo=FALSE,  fig.width=10, fig.height=5}
no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")

MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]

MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]

MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]

MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]

kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)

bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)

#add titles
#use multiplot
#remove X axis

multiplot(kuku,bubu, cols=2)
```

### female and male comparison

## Subspecies level comparison

Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.

*Does this hold up with the full data set?*
*What's the best visualization for these questions?*

Main focus is contrasting the level of difference within and between subspecies.

I'll do this with a mixed model framework approach (and analyze with ANOVA, a method to partition the variance of subset.)

```{r mixed.model, echo=FALSE}

#build mixed model

#run simulations for accounting for sampling error

```

### Subspecies examples / Data

list the strains/data used 
Mouse average MLH1 counts,

### Male Specific results

Not all strains have female data. Here are results from all strains wih male data.

### female and male comparison

## Meiotic Error

This section reviews all the data from achiasmy and 0CO bivalents.