### 2019-10-11 doc/Intro.Rmd
small updates for citations-bibliography

---
title: "Intro"
author: "April Peterson"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: yes
      toc_depth: '5'
      toc_float: yes
bibliography: refs_Intro.bib
---

# Outline

## Background:

### prev hetC lit:
(Lenormand, BurtBellOther, Shapley,)
HetC theories:   (Brandavain, Trivers ??? Lenormand, Sardell-Kirkpatrick, (hulten? Handel?)

wild pops genetic diversity and within species measures

my pero refs, honey bee RR, ritz noor

## gap, angle

# text

  Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products. The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes. Under one of the most common forms of sexual reproduction, anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates. Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.

# Notes and Ideas

# Background

## natural variation in RR

empiracal measures -- important / useful for understanding how rec evolves with narutal selection

sources from Ritz Noor,
Samuk et al 2019, (selection,)
(barley in north africa),  Dreissig et al 2019

humans (not very natural population)

## empiracal measures for hetC / theories for evo

Lenormand Duthiel,
Shapley / stapley et al
Sardell Kirkpatrick  
Gruhn + hassold

Hoffman papers -- strange inverted meiosis
Uniparental disomy
-- evidence that RR-constraint connection may be not straight forward.

### lack of female data / sig of female data

human annuploidy / fertility

# Bibliography