### 2019-11-12 doc/MouseLevelPlots.Rmd
updated the mouse level count by quality plots. started list of mice to exclude

---
title: "mouse level quality plots"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)

setwd("~./MLH1repo/")
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")

#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")#batch15 isn't showing up
#load(file="data/MetaData.RData")#this Rdata was replacing the most current batch.. agg!

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

knitr::opts_chunk$set(echo = TRUE)

#remove other!
MLH1_data <- MLH1_data[MLH1_data$strain != "other",]

MLH1_data <- droplevels(MLH1_data)
```

This rmd file shows the cell counts for mouse level for MLH1 count and quality. The black lines are the strain means and the red lines are the mouse level means. Points are colored by the quantification batch.

These plots are meant to asses which mice can be removed from the final dataset for too few or low quality count data.

Yellow colored boxes are mice that will be exclude from the final (HQ) dataset for low quality cells (5). Consider re-runing this with the final data frame to see how the strain and mouse means change.

### Updating plots

*1. check which category levels do not have data, and enter that index into 'df_list'. Otherwise the loop will be broken.*

*2. Run the code chunks.*

```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE}
### to integrate 'pass'/fail
### this should be in setup data

exclude_mice <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
                  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                  "3sep18_AST_m1",
                  "3jel18_TOM_m1","3sep18_TOM_m1",
                  "31jul17_HMI_m1",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )

#based on this plot make a list of mice to exclude, and figure out code to outline in red
#mostly excluding based on number of 5quality cells

#mice.to.exclude <-

#WSB; "10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
#G;  "8jun15_G_f3", "8jun15_G_f4"
#LEW; "4jan17_LEW_f2"
#PWD; "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5"
#MSM; "12sep16_MSM_f3",#"14jul17_MSM_f1"
#MOLF; "27oct18_MOLF_m3"
#SKIVE; "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3"
#CZECH; "27aug18_CZECH_m1","8jun18_CZECH_m1"
#AST: "3sep18_AST_m1"
#TOM; "3jel18_TOM_m1","3sep18_TOM_m1"
#HMI; "31jul17_HMI_m1"
#SPIC; "21aug17_SPI_f1", "30may18_SPIC_f1", #"30oct17_SPIC_m1", "8jun18_SPIC_m2",

#these plots are filled in yellow, but a different alpha/transparency is being applied - I don't know why

#this loop makes a quant_status col for a list of mice (THERE'S AN ARGUMENT FOR Drawing a box around the mouse blocks )
#switch this loop to ex
for(i in 1:length(MLH1_data$mouse)){
  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], exclude_mice), "exclude", NA)
}

#CZECH is not in this list!!!

df_list <- split(MLH1_data, (MLH1_data$category))
#remove sections which are in category but have no data
#this is where you remove the non-data data!
df_list <- df_list[ -c(24, 29) ]

#12 MOLF female, 18 CZECH female?, 24 HMI female, 29 CAROLI female? 31 other
#

#df list, is a list of df's which were split() from BIG df by category
df_list 
o=0
i=0

#MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)

#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
for(i in 1:(length(df_list))){
 # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
                      mouse_nMLH1 = mean(adj_nMLH1.foci)
  )
# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
  for(o in 1:length(df_list[[i]]$Original.Name)){
    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
        }
}

plot_list = list()
for (h in 1:length(df_list)) {
  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + xlim(c(0.5,5.5))+

    #put a box over strain plots based on the col, 'quant_status'
      geom_rect(data = df_list[[h]], aes(fill= quant_status), alpha= 0.045, xmin=-100, xmax=100, ymin=-100, ymax=100)+

      scale_fill_manual(values=c("yellow"), breaks=NULL) +

    facet_wrap(~ mouse, drop = TRUE) + 
    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +

      geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
  #make an empty ggplot? for missing data

  plot_list[[h]] =   pf
}

#orig (I think this is the same as above)
#plot_list = list()
#for (h in 1:length(df_list)) {
#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)

#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +

 #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +

    #theme for indiviual mice, 
#theme(plot.background = element_rect(fill = "green"))
#  plot_list[[h]] =   pf
#}
#can't change the alpha/transparency of the red rec

#aes must be same length ---   #geom_point(stat = "identity") +
# the last one is correct....?

#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
```

```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
geom_hline(aes(yintercept = med, group = gr), colour = 'red')
#no plot is made that goes in the list -- so it can't be shown
for (gg in 1:length(plot_list) ){
  show(plot_list[gg])
 # invisible(lapply(obj, function(x) plot(x,main="some plot")))
}
```