### 2019-12-05 results/Chapter3_ideas.Rmd
started writing permutations for in silico nuc oermutations

---
title: "Chapter3_ideas"
author: "April Peterson"
date: "November 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
#load biv data
#load MLH1 data

```

Take some mouse level distributions -- pull cells from top and bottom 10%?
(gather whole bivalent level data,)

-start making comparisons for the 2 pools
  how much technical noise is there? (staining differences?/ quality score?

Mice with many cells and BivData
17mar16_G_f4 - BivData
17mar16_G_f1

```{r choosing.mice, echo=FALSE}
#Choose mice with many cells

#G_f, G_m
#SKIVE_f, SKIVE_m #have BivData

#remove all the quality 5

#make plot/histogram of the MLH1 count distribution
counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]

mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]

ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
topsy <- ordered.MLH1[1:5,]
bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
top.n.bottom.bin <- rbind(topsy,bottomzy)

#write 
write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)

shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz
#get the filenames for this groups of cells

#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
#28feb15_PWD_f2, in PWD--female .xlsx file

#31may19_SKIVE_m2
male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
male.one.DF <- male.one.DF[male.one.DF$quality < 5,]

ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
male.high.bin.DF <- ordered.MLH1.male[1:5,]
male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
#write 
write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)

shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz.too

```

the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
Should I be comparing cells with the same number of COs?

```{r getting.cell.names, echo=FALSE}

#load in the cell dataset
#28feb15_PWD_f2  from Curated_BivData
#for a mouse, pull the 10 highest rec cells

PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
#"Foci3"                   "curation.notes"  
female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]

write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)

PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")

#merge with hand data
#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)

cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
                            .(fileName),
                            summarize, 
                            ncells = length(unique(fileName)),
                            nbivs = length(unique(Obj.ID)),
                            n.foci = sum(hand.foci.count),
                            total.SC = sum(chromosomeLength),
                            mean.SC = mean(chromosomeLength)

)
#add hand data
ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]

#I have 3 cells with full 20

```

Start discribing the Bivalent patterns for the PWD female patterns

```{r mini.biv, echo=FALSE}
#make mini biv plots

#PWD.female
#PWD.female_hand.measure

cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]

two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))

#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
#of SC.
MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
 #geom_hline(yintercept = c(male_IFD) ) +

  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
  #how to add points for 2foci
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
  #foci.1.position, two_groups$distal.foci.position)
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +

#add centromeres...  
  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+

  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
 # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
  ylim(c(0,200))+
     theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())  #for removeing x labels

#MiniBiv_plot1
#MiniBiv_plot2

#save the files

```

```{r biscribe.Biv.Pattern, echo=FALSE}
#start discribing the basic

#start merging curated bivData with handmeasured data

#BivData - Curated
```

# Whole Cell Measures

excel sheet Whole_cell_measures.xlsx has 840 bivalent measures.

colum

```{r whole.cell.data, echo=FALSE}

#read in csv of whole cell hand measures
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)

whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$fileName) | whole.cell.data$fileName==""), ]

whole.cell.data.XY <- whole.cell.data

whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

#find cells with the wrong number

biv.count_table <-  ddply(.data=whole.cell.data,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length)
)
#all have 20 or 19 (after removing XY)

for(musy in as.character(unique(whole.cell.data$fileName))){
  print(musy)

  #order the chrm lengths -- this is order
  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
}
#I think I saved the Excel with sorted SC lengths
#Yay I printed the rank in this

```

Try doing the permutations

```{r permutations, echo=FALSE}

#need to seperate sexes cause diff numbers
male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
#ncells
length(unique(male.whole.cell$fileName) ) #24 cells

female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456

#not sure how to make the initial DF
#replicate 24 rows with 19 cols
DF.for.perm  <- as.data.frame( replicate(24, "") )
colnames(DF.for.perm) <- c("blah")

DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),
                       rank3= replicate(24, ""),     rank4= replicate(24, ""),
                       rank5= replicate(24, ""),   rank6= replicate(24, ""),
                       rank7= replicate(24, ""), rank8= replicate(24, ""),
                       rank9= replicate(24, ""), rank10= replicate(24, ""),
                       rank11= replicate(24, ""),     rank12= replicate(24, ""),
                       rank13= replicate(24, ""),     rank14= replicate(24, ""),
                       rank15= replicate(24, ""),     rank16= replicate(24, ""),
                       rank17= replicate(24, ""),rank18= replicate(24, ""),rank19= replicate(24, "")
                       )

iterations = 24
variables = 20
DF.permutations <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

#ROW IS FIRST

for(row.num in 1:24){

  #print(c(row.num, "rowis") )

  for(col.num in 2:ncol(DF.permutations) ){

  #print(c(col.num, "colis") )

 DF.permutations[row.num,col.num] <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)

      }

  }
#yes this finally works!!!!

```