### 2019-12-05 results/Chapter3_ideas.Rmd
added in data to start recreating Wang et al findings

---
title: "Chapter3_ideas"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
#load biv data
#load MLH1 data
library(dplyr)
library(plyr)
library(ggplot2)
library(reshape2)
library(raster) #cv

#read files
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17

```

# Summary and recreation of Wang et al 2019

insert summary of the Wang paper

which data did they use? what summary stat were they comparing
What were their results?

human data- 
Gruhn et al 2013 (ploted below)
Sun et al 2006
Oliver-Bonet et al 2007 
Lian et al 2008
Lu et al 2012
Hou et al 2013

```{r human.male.data, echo=FALSE}
#read files
Hs.male_axis <- read.csv("~./MLH1repo/data/Human_male_axis_length.csv", header = TRUE)
Hs.male_MLH1 <- read.csv("~./MLH1repo/data/Human_male_MLH1.csv", header = TRUE)
#I assume that the rows of MLH1 and SC lengths are the linked observations, check or verify this assumption
#if they are using within cell -- they must

#melt DF -- how do I keep row.name info?
Melted.Hs.male_axis <- melt(Hs.male_axis, 
                            c("row.name", "chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22"), 
                         id=c("row.name") )
#remove the first part with of data frame, variable = "row.name""

#clean up dm1 (there are a bunch of rows without useful information)
Melted.Hs.male_axis <- Melted.Hs.male_axis[(!is.na(Melted.Hs.male_axis$value)),]
Melted.Hs.male_axis <- Melted.Hs.male_axis[((Melted.Hs.male_axis$variable != "row.name")),]

colnames(Melted.Hs.male_axis) <- c("row.name", "Chrm.id", "axis.length")

#Seems like there is an outlier, 
Melted.Hs.male_axis <- Melted.Hs.male_axis[(Melted.Hs.male_axis$axis.length < 700 ),]

hs.sc.length_plot <- ggplot(Melted.Hs.male_axis, aes(y=axis.length, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+geom_boxplot()+ 
  ggtitle("Human male SC Lengths")+theme(legend.position="none")
hs.sc.length_plot

```

Neat. cool to see human SC data plotted. One outlier was removed.

# Adding to whole Cells

Look/use the DF, Curated_BivData

Take some mouse level distributions -- pull cells from top and bottom 10%?
(gather whole bivalent level data,)

-start making comparisons for the 2 pools
  how much technical noise is there? (staining differences?/ quality score?

Mice with many cells and BivData
17mar16_G_f4 - BivData
17mar16_G_f1

Chunk below has code for choosing mice with a good number of cells (25?)

```{r choosing.mice, echo=FALSE, eval=FALSE}
#Choose mice with many cell observations 
#important for increased power for within mouse variance

#G_f, G_m
#SKIVE_f, SKIVE_m #have BivData

#remove all the quality 5

#make plot/histogram of the MLH1 count distribution
counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]

mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]

ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
topsy <- ordered.MLH1[1:5,]
bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
top.n.bottom.bin <- rbind(topsy,bottomzy)

#write 
write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)

shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
#highest of the low bin
#lowest value of the high bin

#get the filenames for this groups of cells

#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
#28feb15_PWD_f2, in PWD--female .xlsx file

#31may19_SKIVE_m2
male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
male.one.DF <- male.one.DF[male.one.DF$quality < 5,]

ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
male.high.bin.DF <- ordered.MLH1.male[1:5,]
male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
#write 
write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)

shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz.too

```

the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
Should I be comparing cells with the same number of COs?

```{r getting.cell.names, echo=FALSE,eval=FALSE}

#load in the cell dataset
#28feb15_PWD_f2  from Curated_BivData
#for a mouse, pull the 10 highest rec cells

PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
#"Foci3"                   "curation.notes"  
female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]

write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)

PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")

#merge with hand data
#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)

cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
                            .(fileName),
                            summarize, 
                            ncells = length(unique(fileName)),
                            nbivs = length(unique(Obj.ID)),
                            n.foci = sum(hand.foci.count),
                            total.SC = sum(chromosomeLength),
                            mean.SC = mean(chromosomeLength)

)
#add hand data
ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]

#I have 3 cells with full 20

```

Start discribing the Bivalent patterns for the PWD female patterns

```{r mini.biv, echo=FALSE,eval=FALSE}
#make mini biv plots

#PWD.female
#PWD.female_hand.measure

cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]

two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))

#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
#of SC.
MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
 #geom_hline(yintercept = c(male_IFD) ) +

  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
  #how to add points for 2foci
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
  #foci.1.position, two_groups$distal.foci.position)
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +

#add centromeres...  
  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+

  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
 # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
  ylim(c(0,200))+
     theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())  #for removeing x labels

#MiniBiv_plot1
#MiniBiv_plot2

#save the files

```

```{r biscribe.Biv.Pattern, echo=FALSE}
#start discribing the basic

#start merging curated bivData with handmeasured data

#BivData - Curated
```

from the curated BivData, figure out which can be added / into the whole cell data. I think there might only be a couple cell to add in.

```{r add.more.whole.cell, echo=FALSE}

#load data curatedBiv Data
#load in hand measures

```

# Whole Cell Measures

excel sheet Whole_cell_measures.xlsx has 840 bivalent measures.

```{r setup.whole.cell.data, echo=FALSE}

#read in csv of whole cell hand measures
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)

whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$fileName) | whole.cell.data$fileName==""), ]

whole.cell.data.XY <- whole.cell.data
whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

#make sure foci.count is number
whole.cell.data$hand.foci.count <- as.character(whole.cell.data$hand.foci.count)
whole.cell.data$hand.foci.count <- as.numeric(whole.cell.data$hand.foci.count)

#find cells with the wrong number

#ADD THE RANK
for(musy in as.character(unique(whole.cell.data$fileName))){
  #print(musy)

  #order the chrm lengths -- this is order
  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
}
#I think I saved the Excel with sorted SC lengths
#Yay I printed the rank in this

```

## Adding to whole cell measures

calculated the observed values in this chunk.

```{r cell.level.tables, echo=FALSE}

whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
#ncells
#length(unique(male.whole.cell$fileName) ) #24 cells

female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
#length(unique(female.whole.cell$fileName) ) #18 cells

biv.count_table <-  ddply(.data=whole.cell.data,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length)
)#all have 20 or 19 (after removing XY)

#real data
male.biv.count_table <-  ddply(.data=male.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                 mean_co = mean(hand.foci.count, rm.na = TRUE),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

female.biv.count_table <-  ddply(.data=female.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                 mean_co = mean(hand.foci.count),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)#one of the female cells is missing values

```

# Plots of whole cell data

show SC length (y) by rank  (x), facet (sex)

```{r play.plots, echo=FALSE}
#female.whole.cell
#male.whole.cell

female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$SC.length) | female.whole.cell$SC.length==""), ]

cow.hugger <- ggplot(female.whole.cell, aes(x=rank, y=SC.length, color=mouse) ) +geom_jitter()

rank.length.boxplot <- ggplot(data=female.whole.cell) + geom_boxplot( aes(x=as.factor(rank), y=SC.length, fill=rank) )+geom_jitter(aes(x=as.factor(rank), y=SC.length) )+ggtitle("blaah")

rank.length.boxplot.xx <- ggplot(data=female.whole.cell, aes(x=as.factor(rank), y=SC.length) ) + geom_boxplot()+ geom_jitter(aes( color=fileName))+
  ggtitle("Female SC Length Dist by ranks")+theme(legend.position="none")

rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=SC.length)) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")

rank.length.boxplot.xx
rank.length.boxplot.xy

```

# PERMUTATION SETUP

```{r per.co.count, echo=FALSE, eval=FALSE}
#I don't think ... I end up useing these (matrics / DFs)
iterations = 24
variables = 20
DF.permutations_CO.count <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations_CO.count) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

for(row.num in 1:24){

  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations_CO.count) ){

  rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations_CO.count[row.num,col.num] <- rand.rank.co
   }
  }
DF.permutations_CO.count <- as.data.frame(DF.permutations_CO.count)

#no rank - iterate / permute the sampling X times
its = 24
varis = 20
DF.permutations.CO.no_rank <- matrix(ncol=varis, nrow=its)

#make LIST of X DATAFRAMES

for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.CO.no_rank) ){
  rand.co <- sample(male.whole.cell$hand.foci.count, 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.CO.no_rank[row.num,col.num] <- rand.co
   }
  }#yes this finally works!!!!
DF.permutations.CO.no_rank <- as.data.frame(DF.permutations.CO.no_rank)

#
for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
  }

#merge with real
co.var.for.plot <- matrix(ncol=3, nrow=72)
co.var.for.plot <- as.data.frame(co.var.for.plot)
colnames(co.var.for.plot) <- c("cell.id","var","type")

co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
co.var.for.plot$type[1:24] <- "sim"

co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
co.var.for.plot$type[25:48] <- "real"

co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
co.var.for.plot$type[49:72] <- "sim_no_rank"

#plot
#whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
#whole.cell.var_CO.plot

```

## CO count, male

Make list of DFs to fill with permutations

```{r list.of.DF,echo=FALSE}

#dataframes to recreate -- and fill with 
#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)

#make subsampled piece of DF here -- these are taking the whole DF
#I will just take single values
#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
#male.whole.cell
#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)

#rank (sample all of the ranks then stich them together)
BIG.LIST_w.rank <- replicate(1000,
  list(bind_rows(list( 
  male.whole.cell[ sample( which(male.whole.cell$rank==1), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==2), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==3), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==4), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==5), 1, replace = FALSE ),],
   male.whole.cell[ sample( which(male.whole.cell$rank==6), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==7), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==8), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==9), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==10), 1, replace = FALSE ),],

   male.whole.cell[ sample( which(male.whole.cell$rank==11), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==12), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==13), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==14), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==15), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==16), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==17), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==18), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==19), 1, replace = FALSE ),]
))))
#this is replicating single cell (NOT a SAMPLE of 24)

#LIST of DF, no rank (include all), samples all 19 bivs at a time
#makes a list of 10 -- [[]]
BIG.DF.LIST_no.rank <- replicate(1000,
                  list(bind_rows(list( 
          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
) ) ) )

#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

#Calq cV for the rank perm
cv.val.DF_rank <- data.frame(  cV_co.count_rank = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  cv.val.DF_rank$cV_co.count_rank[j] <-cv(  BIG.LIST_w.rank[[j]]$hand.foci.count )
  }

#Calq cV for no rank
cv.val.DF_no.rank <- data.frame(  cV_co.count_no.rank = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  cv.val.DF_no.rank$cV_co.count_no.rank[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
  }
#extract the 
#next step make cell-level calqs of cV -- for plotting

#merge + plots of the cvs
#10 and 10 rows from the permuted
cv.val.DF_rank$type <- "sim w.rank"
colnames(cv.val.DF_rank) <- c("cV","type" )

cv.val.DF_no.rank$type <- "sim no rank"
colnames(cv.val.DF_no.rank) <- c("cV", "type")

#then add the real values
real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count)
real.vals.cv.co$type <- "real"
colnames(real.vals.cv.co) <- c("cV", "type")

#merge
cVs_permutations.DF <- rbind(cv.val.DF_rank, cv.val.DF_no.rank, real.vals.cv.co)

# need to melt the DFs before ploting?
#ploting --- plot with more sims
oo.plot <- ggplot(cVs_permutations.DF, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")

#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")

#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")

```

The above plot the real data might have more variance because it's a combination of strains

## CO count, female

```{r CO.count.female, echo=FALSE}

#dataframes to recreate -- and fill with 
#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)

#make subsampled piece of DF here -- these are taking the whole DF
#I will just take single values
#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
#male.whole.cell
#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)

#rank (sample all of the ranks then stich them together)
BIG.LIST_w.rank_female <- replicate(1000,
  list(bind_rows(list( 
  female.whole.cell[ sample( which(female.whole.cell$rank==1), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==2), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==3), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==4), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==5), 1, replace = FALSE ),],
   female.whole.cell[ sample( which(female.whole.cell$rank==6), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==7), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==8), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==9), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==10), 1, replace = FALSE ),],

   female.whole.cell[ sample( which(female.whole.cell$rank==11), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==12), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==13), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==14), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==15), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==16), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==17), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==18), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==19), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==20), 1, replace = FALSE ),]
  ))))

#LIST of DF, no rank (include all), samples all 19 bivs at a time
#makes a list of 10 -- [[]]
BIG.DF.LIST_no.rank_female <- replicate(1000,
                  list(bind_rows(list( 
          female.whole.cell[ sample( which( female.whole.cell$rank < 20), 20, replace = FALSE ),]
) ) ) )

```

These permuted dataframes can be reused for all the meiotic traits, (CO.count, SC length, and position).

```{r sum.stat_cv.CO, echo=FALSE}
#CALUCULATING SUMMARY STAT from the permuted DFs
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

#Calq cV for the rank perm
cv.val.DF_rank.female <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  cv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$hand.foci.count ) #for each
  }

#Calq cV for no rank
cv.val.DF_no.rank.female <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  cv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count ) #for each 
}

#merge + plots of the cvs
#10 and 10 rows from the permuted
cv.val.DF_rank.female$type <- "sim w.rank"
colnames(cv.val.DF_rank) <- c("cV","type" )

cv.val.DF_no.rank.female$type <- "sim no rank"
colnames(cv.val.DF_no.rank.female) <- c("cV", "type")

#then add the real values
real.vals.cv.co_female <- data.frame(female.biv.count_table$cv_co.count)
real.vals.cv.co_female$type <- "real"
colnames(real.vals.cv.co_female) <- c("cV", "type")

#merge
cVs_permutations.DF_female <- rbind(cv.val.DF_rank.female, cv.val.DF_no.rank.female, real.vals.cv.co_female)

# need to melt the DFs before ploting?
#ploting --- plot with more sims
Femaleplot_cV <- ggplot(cVs_permutations.DF_female, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count female, 1000 simulated")

#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")

#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")

cVs_permutations.DF$sex <- "male"
cVs_permutations.DF_female$sex <- "female"

all.cv <- rbind(cVs_permutations.DF, cVs_permutations.DF_female)
cV_plot <- ggplot(all.cv, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")+facet_wrap(~sex)
cV_plot

```

# SC length, bivalent level

don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.

```{r permutations.sc, echo=FALSE, eval=FALSE}

#merge with real
sc.var.for.plot <- matrix(ncol=3, nrow=72)
sc.var.for.plot <- as.data.frame(sc.var.for.plot)
colnames(sc.var.for.plot) <- c("cell.id","var","type")

sc.var.for.plot$var[1:24] <- DF.permutations.sc$var_sc
sc.var.for.plot$type[1:24] <- "sim"

sc.var.for.plot$var[25:48] <- male.biv.count_table$var_sc
sc.var.for.plot$type[25:48] <- "real"

sc.var.for.plot$var[49:72] <- DF.permutations.no_rank$var_rand.sc
sc.var.for.plot$type[49:72] <- "sim_no_rank"

#plot the forms
#plot within cell variance
whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
whole.cell.var_SC.plot
```

```{r sum.stat.SClength, echo=FALSE}

#Calq SC cV for the SC length FEMALE
SCcv.val.DF_rank.female <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(j)
  #fill in the cells of given index
  SCcv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$SC.length ) #for each
  }

#Calq SC cV for no rank
SCcv.val.DF_no.rank.female <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SCcv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$SC.length ) #for each 
}

#merge + plots of the cvs
#10 and 10 rows from the permuted
SCcv.val.DF_rank.female$type <- "sim w.rank"
colnames(SCcv.val.DF_rank.female) <- c("cV","type" )

SCcv.val.DF_no.rank.female$type <- "sim no rank"
colnames(SCcv.val.DF_no.rank.female) <- c("cV", "type")

#then add the real values
real.vals.cv.SC_female <- data.frame(female.biv.count_table$cv_SC.length)
real.vals.cv.SC_female$type <- "real"
colnames(real.vals.cv.SC_female) <- c("cV", "type")

#merge and plots
cV.SC_permutations.DF_female <- rbind(SCcv.val.DF_rank.female, SCcv.val.DF_no.rank.female, real.vals.cv.SC_female)

#MALE --BIG.LIST_w.rank, BIG.DF.LIST_no.rank,
#Calq cV for the SC length
SCcv.val.DF_rank.male <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
  #print(j)
  #fill in the cells of given index
  SCcv.val.DF_rank.male$cV[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
  }

#Calq cV for no rank
SCcv.val.DF_no.rank.male <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SCcv.val.DF_no.rank.male$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length ) #for each 
}

#merge + plots of the cvs
#10 and 10 rows from the permuted
SCcv.val.DF_rank.male$type <- "sim w.rank"
colnames(SCcv.val.DF_rank.male) <- c("cV","type" )

SCcv.val.DF_no.rank.male$type <- "sim no rank"
colnames(SCcv.val.DF_no.rank.male) <- c("cV", "type")

#then add the real values-- male.biv.count_table
real.vals.cv.SC_male <- data.frame(male.biv.count_table$cv_SC.length)
real.vals.cv.SC_male$type <- "real"
colnames(real.vals.cv.SC_male) <- c("cV", "type")

#merge and plots
cV.SC_permutations.DF_male <- rbind(SCcv.val.DF_rank.male, SCcv.val.DF_no.rank.male, real.vals.cv.SC_male)
cV.SC_permutations.DF_male$sex <- "male"
cV.SC_permutations.DF_female$sex <- "female"  #cV.SC_permutations.DF_female
all_cv.SC <- rbind(cV.SC_permutations.DF_female, cV.SC_permutations.DF_male)

cV.SC_plot <- ggplot(all_cv.SC, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv of within cell SC lengths, 1000 permutations")+facet_wrap(~sex)
```

```{r permuted.v.real, echo=FALSE, eval=FALSE}
for( o in 1:nrow(DF.permutations_CO.count) ) {
  DF.permutations_CO.count$var_CO.count[o] <- var(as.numeric(DF.permutations_CO.count[o,2:20]))
  }

#two data frames of within cell var
#cell.id, var, type
var.for.plot <- matrix(ncol=3, nrow=48)
var.for.plot <- as.data.frame(var.for.plot)
colnames(var.for.plot) <- c("cell.id","var","type")

var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
var.for.plot$type[1:24] <- "sim"

var.for.plot$var[24:48] <- male.biv.count_table$var_co
var.for.plot$type[24:48] <- "real"

#my permuted DFs are in a strange str

#plot within cell variance
whole.cell.var.plot  <- ggplot(var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell variance between types")

whole.cell.var.plot
```

# Junk

```{r junk, eval=FALSE}
DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),
                       rank3= replicate(24, ""),     rank4= replicate(24, ""),
                       rank5= replicate(24, ""),   rank6= replicate(24, ""),
                       rank7= replicate(24, ""), rank8= replicate(24, ""),
                       rank9= replicate(24, ""), rank10= replicate(24, ""),
                       rank11= replicate(24, ""),     rank12= replicate(24, ""),
                       rank13= replicate(24, ""),     rank14= replicate(24, ""),
                       rank15= replicate(24, ""),     rank16= replicate(24, ""),
                       rank17= replicate(24, ""),rank18= replicate(24, ""),rank19= replicate(24, "")
                       )

#not sure how to make the initial DF
#replicate 24 rows with 19 cols
iterations = 24
variables = 20
DF.permutations.sc <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

#permutations -- for SC

#permutations without accounting for rank
its = 24
varis = 20
DF.permutations.no_rank <- matrix(ncol=varis, nrow=its)
#colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
#                               "rank6","rank7","rank8","rank9","rank10","rank11",
#                               "rank12","rank13","rank14","rank15","rank16",
#                               "rank17","rank18","rank19" )

#ROW IS FIRST
for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.sc) ){
  rand.rank.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.sc[row.num,col.num] <- rand.rank.sc
   }
  }#yes this finally works!!!!
DF.permutations.sc <- as.data.frame(DF.permutations.sc)

#pair with real within cell SC variance
for( o in 1:nrow(DF.permutations.sc) ) {
  DF.permutations.sc$var_sc[o] <- var(as.numeric(DF.permutations.sc[o,2:20]))
  }

for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.no_rank) ){
  rand.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.no_rank[row.num,col.num] <- rand.sc
   }
  }#yes this finally works!!!!
DF.permutations.no_rank <- as.data.frame(DF.permutations.no_rank)

for( o in 1:nrow(DF.permutations.sc) ) {
  DF.permutations.no_rank$var_rand.sc[o] <- var(as.numeric(DF.permutations.no_rank[o,2:20]))
  }

```