### 2019-12-06 results/Chapter3_ideas.Rmd
working of randomizing foci position

---
title: "Chapter3_ideas"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
#load biv data
#load MLH1 data
library(dplyr)
library(plyr)
library(ggplot2)
library(reshape2)
library(raster) #cv

#read files
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
```

# Summary and of Wang et al 2019

Main take away from Wang et al 2019 is that chromosomes with the same nucleus covary in number of crossovers. Axis lengths / SC lengths show the same pattern and is known to be regulated at the cell wide level. The authors suggest that the within nucleus covariation in CO per chromosome is mediated through cell wide regulation of SC area. (the lengths of SC for chromosomes within a cell a correlated, there is more area for COs).

Data used in the paper
human data- Gruhn et al 2013 (ploted below, 755 cells)
Sun et al 2006
Oliver-Bonet et al 2007 
Lian et al 2008
Lu et al 2012
Hou et al 2013

Was there originall CO position data for this human data? (MLH1 spreads) (there is axis length and CO count per chrm -- this should mean there is position )

```{r human.male.data, echo=FALSE}
#read files
Hs.male_axis <- read.csv("~./MLH1repo/data/Human_male_axis_length.csv", header = TRUE)
#I assume that the rows of MLH1 and SC lengths are the linked observations, check or verify this assumption
#if they are using within cell -- they must

#melt DF -- how do I keep row.name info?
Melted.Hs.male_axis <- melt(Hs.male_axis, 
                            c("row.name", "chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22"), 
                         id=c("row.name") )
#remove the first part with of data frame, variable = "row.name""

#clean up dm1 (there are a bunch of rows without useful information)
Melted.Hs.male_axis <- Melted.Hs.male_axis[(!is.na(Melted.Hs.male_axis$value)),]
Melted.Hs.male_axis <- Melted.Hs.male_axis[((Melted.Hs.male_axis$variable != "row.name")),]

colnames(Melted.Hs.male_axis) <- c("row.name", "Chrm.id", "axis.length")

#Seems like there is an outlier, 
Melted.Hs.male_axis <- Melted.Hs.male_axis[(Melted.Hs.male_axis$axis.length < 700 ),]

hs.sc.length_plot <- ggplot(Melted.Hs.male_axis, aes(y=axis.length, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+geom_boxplot()+ 
  ggtitle("Human male SC Lengths")+theme(legend.position="none")
hs.sc.length_plot

#### .. Hs.male_MLH1
Hs.male_MLH1 <- read.csv("~./MLH1repo/data/Human_male_MLH1.csv", header = TRUE)

Melted.Hs.male_MLH1 <- melt(Hs.male_MLH1, 
                            c("row.name", "chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22"), 
                         id=c("row.name") )
#remove the first part with of data frame, variable = "row.name""

#clean up dm1 (there are a bunch of rows without useful information)
Melted.Hs.male_MLH1 <- Melted.Hs.male_MLH1[(!is.na(Melted.Hs.male_MLH1$value)),]
Melted.Hs.male_MLH1 <- Melted.Hs.male_MLH1[((Melted.Hs.male_MLH1$variable != "row.name")),]
colnames(Melted.Hs.male_MLH1) <- c("row.name", "Chrm.id", "CO")

#Seems like there is an outlier, 
#Melted.Hs.male_axis <- Melted.Hs.male_axis[(Melted.Hs.male_axis$axis.length < 700 #),]

hs.CO.length_plot <- ggplot(Melted.Hs.male_MLH1, aes(y=CO, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+ 
  ggtitle("Human male CO Counts")+theme(legend.position="none")
hs.CO.length_plot
```

approach -- they compared covariation (cV) of per chrm CO count and SC length for permuted/randomized groupd of chromosome observations representing 'in silico' nuclei.

## Redoing the permutaitons

Cell summaries below

```{r Hs.cell_table, echo=FALSE}
#make table for cell wide measures
Hs.CO_cell_wide <- ddply(.data=Melted.Hs.male_MLH1, 
                            .(row.name),
                            summarize, 
                            nbivs = length(unique(Chrm.id)),
                            n.CO = sum(CO),
                            mean.CO = mean(CO),
                            cV.CO = cv(CO)
)
head(Hs.CO_cell_wide)

Hs.SC_cell_wide <- ddply(.data=Melted.Hs.male_axis, 
                            .(row.name),
                            summarize,

                            nbivs = length(unique(Chrm.id)),
                            total.SC = sum(axis.length),
                            mean.SC = mean(axis.length),
                            cV.SC = cv(axis.length)
)
head(Hs.SC_cell_wide)

```

```{r Hs.Perm_noRank, echo=FALSE}

Melted.Hs.male_MLH1$chrm.num <- (Melted.Hs.male_MLH1$Chrm.id)
levels(Melted.Hs.male_MLH1$chrm.num) <- 1:23

Melted.Hs.male_MLH1$chrm.num <- as.character(Melted.Hs.male_MLH1$chrm.num)
Melted.Hs.male_MLH1$chrm.num <- as.numeric(Melted.Hs.male_MLH1$chrm.num)
#chrm.num is now 1:23

#LIST of DF, no rank (include all), samples all 19 bivs at a time
Hs_male.BIG.DF.LIST_no.rank <- replicate(755,
                  list(bind_rows(list( 
          #can't make do logic on factors
          Melted.Hs.male_MLH1[ sample( which( Melted.Hs.male_MLH1$chrm.num < 24), 22),]
) ) ) )
#now have randomized DF of 1000 cells
```

```{r Hs.Perm_wRank, echo=FALSE}
#rank (sample all of the ranks then stich them together)
Hs.CO_BIG.LIST_w.rank <- replicate(755,
  list(bind_rows(list(

  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==2), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==3), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==4), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==5), 1, replace = FALSE ),],
   Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==6), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==7), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==8), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==9), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==10), 1, replace = FALSE ),],

   Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==11), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==12), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==13), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==14), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==15), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==16), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==17), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==18), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==19), 1, replace = FALSE ),],
    Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==20), 1, replace = FALSE ),],
  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==21), 1, replace = FALSE ),],
    Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==22), 1, replace = FALSE ),],
      Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==23), 1, replace = FALSE ),]
))
)) #list of DFs which are 'cells' with observations drawn from pooled chrm bins
```

```{r sum.stat.display, echo=FALSE}
#after making big randomize / permuted data sets, CALCULATE the metrics / sum.stat
#Hs.CO_BIG.LIST_w.rank
#Hs_male.BIG.DF.LIST_no.rank

#Calq cV for the rank perm
#1. make empty DF
Hs.CO_cv_rank <- data.frame(  cV.CO = rep(0, length(Hs.CO_BIG.LIST_w.rank) ) )

for(j in 1:length(Hs.CO_BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
  #print(j)
  #fill in the cells of given index
  Hs.CO_cv_rank$cV.CO[j] <-cv(  Hs.CO_BIG.LIST_w.rank[[j]]$CO ) #but j should be row, row should be a cell
  #add CO per nuc
  Hs.CO_cv_rank$co.per.nuc[j] <- sum(Hs.CO_BIG.LIST_w.rank[[j]]$CO)
  }

#Calq cV for no rank
Hs.CO_cv_no.rank <- data.frame(  cV.CO = rep(0, length(Hs_male.BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(Hs_male.BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  Hs.CO_cv_no.rank$cV.CO[i] <-cv(  Hs_male.BIG.DF.LIST_no.rank[[i]]$CO )
  Hs.CO_cv_no.rank$co.per.nuc[i] <- sum(Hs_male.BIG.DF.LIST_no.rank[[i]]$CO) #these are the total sum, not divided out by nuc!
  }

#merge simulated and real values into 1DF
Hs.CO_cv_rank$type <- "sim w.rank"
colnames(Hs.CO_cv_rank) <- c("cV","COs","type" )

Hs.CO_cv_no.rank$type <- "sim no rank"
colnames(Hs.CO_cv_no.rank) <- c("cV", "COs","type")

#then add the real values   Hs.CO_cell_wide
Hs_real.cv.co <- data.frame(cV = Hs.CO_cell_wide$cV.CO, Hs.CO_cell_wide$n.CO)
Hs_real.cv.co$type <- "real"
colnames(Hs_real.cv.co) <- c("cV","COs", "type")

#merge
Hs.CO_cVs.DF <- rbind(Hs.CO_cv_rank, Hs.CO_cv_no.rank, Hs_real.cv.co)

# need to melt the DFs before ploting?
#ploting --- plot with more sims
#HS.plot <- ggplot(Hs.CO_cVs.DF, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("Cell-wide cv for CO, Human male")
#HS.plot
```

```{r HS.plots, echo=FALSE}

HS.scatter.plot <- ggplot(Hs.CO_cVs.DF, aes(y = COs, x=type) )+ geom_jitter()+ ggtitle("COs per cell, Human male")
HS.scatter.plot

#make density distributions
#percent of cells

#quantilles of COs

hist <- ggplot(Hs.CO_cVs.DF, aes(x = COs)) + geom_histogram(aes(fill=type), binwidth = 2 )
hist

HS.density.plot <- ggplot(Hs.CO_cVs.DF, aes(x = COs, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Human male")+xlim(c(0,100))
HS.density.plot
```

To calculate the variability index, authors take percent of cells(integrated curve area) which differ between the empiracal and simulated distributions.

I changed the number of permutations to 200, which should be the same number of oberved cells, but it looks like the real data is much larger....the sheet has 755 cell observations, not 200. This might be due to how histograms are presented.

The rank distinction is to test if sampling within pools of specific chromsomes makes a difference. The smaller/tighter distribution between CO count per nuc with the ranked and non-ranked distributions indecate there is a small chromosome specific effect.

Not sure how to calculate the differences in distributions yet.

this current data shows random samples, not permuted random samples.

# Mus whole Cell Data Set

```{r setup.whole.cell.data, echo=FALSE}

#read in csv of whole cell hand measures
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)
whole.cell.data <- add_strain(whole.cell.data)
whole.cell.data <- add_category(whole.cell.data)

whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$fileName) | whole.cell.data$fileName==""), ]
#this doesn't remove all of the data that are missing values

whole.cell.data$Obj.ID <- paste(whole.cell.data$fileName, whole.cell.data$boxNumber, sep = "_")

whole.cell.data.XY <- whole.cell.data
whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

#make sure foci.count is number
whole.cell.data$hand.foci.count <- as.character(whole.cell.data$hand.foci.count)
whole.cell.data$hand.foci.count <- as.numeric(whole.cell.data$hand.foci.count)

whole.cell.data$Foci1 <- whole.cell.data$F1
whole.cell.data$Foci2 <- whole.cell.data$F2
whole.cell.data$Foci3 <- whole.cell.data$F3

#add IFDs and normalized positions 
whole.cell.data$Foci1.PER <- whole.cell.data$F1 / whole.cell.data$SC.length
whole.cell.data$Foci2.PER <- whole.cell.data$F2 / whole.cell.data$SC.length
whole.cell.data$Foci3.PER <- whole.cell.data$F3 / whole.cell.data$SC.length

whole.cell.data <- add_IFD(whole.cell.data)
whole.cell.data <-  add_IFD.2(whole.cell.data)

Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
#IFD.2 does mulitple IFDs (just absolute)
Curated_BivData <- add_IFD.2(Curated_BivData)

#find cells with the wrong number
#whole.cell.data_cell.table <-

#ADD THE RANK
for(musy in as.character(unique(whole.cell.data$fileName))){
  #print(musy)

  #order the chrm lengths -- this is order
  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
}
#I think I saved the Excel with sorted SC lengths
#Yay I printed the rank in this

```

To make the same comparisons, I need chromosome/bivalent level observations for whole cells (all chromosomes). 
I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations

```{r whole.cell.table, echo=FALSE}

table(whole.cell.data$category)

#table(cell.table$category)

```

```{r cell.level.tables, echo=FALSE}

whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
#ncells
#length(unique(male.whole.cell$fileName) ) #24 cells

female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
#length(unique(female.whole.cell$fileName) ) #18 cells

biv.count_table <-  ddply(.data=whole.cell.data,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length)
)#all have 20 or 19 (after removing XY)

#real data
male.biv.count_table <-  ddply(.data=male.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),

                 CO = sum(hand.foci.count),
                 total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count, rm.na = TRUE),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

female.biv.count_table <-  ddply(.data=female.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),

                  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),

                 CO = sum(hand.foci.count),
                  total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)#one of the female cells is missing values

```

I calulate within cell ranked SC length to use rank as a proxy for chromsome ID for bivalents.

Below are plots of the SC distributions across the rank

```{r play.plots, echo=FALSE}
#female.whole.cell
#male.whole.cell
female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$SC.length) | female.whole.cell$SC.length==""), ]

cow.hugger <- ggplot(female.whole.cell, aes(x=rank, y=SC.length, color=mouse) ) +geom_jitter()

rank.length.boxplot <- ggplot(data=female.whole.cell) + geom_boxplot( aes(x=as.factor(rank), y=SC.length, fill=rank) )+geom_jitter(aes(x=as.factor(rank), y=SC.length) )+ggtitle("blaah")

rank.length.boxplot.xx <- ggplot(data=female.whole.cell, aes(x=as.factor(rank), y=SC.length) ) + geom_boxplot()+ geom_jitter(aes( color=fileName))+
  ggtitle("Female SC Length Dist by ranks")+theme(legend.position="none")

rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=SC.length)) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")

rank.length.boxplot.xx
rank.length.boxplot.xy

```

# PERMUTATION SETUP

```{r per.co.count, echo=FALSE, eval=FALSE}
#I don't think ... I end up useing these (matrics / DFs)
iterations = 24
variables = 20
DF.permutations_CO.count <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations_CO.count) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

for(row.num in 1:24){

  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations_CO.count) ){

  rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations_CO.count[row.num,col.num] <- rand.rank.co
   }
  }
DF.permutations_CO.count <- as.data.frame(DF.permutations_CO.count)

#no rank - iterate / permute the sampling X times
its = 24
varis = 20
DF.permutations.CO.no_rank <- matrix(ncol=varis, nrow=its)

#make LIST of X DATAFRAMES

for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.CO.no_rank) ){
  rand.co <- sample(male.whole.cell$hand.foci.count, 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.CO.no_rank[row.num,col.num] <- rand.co
   }
  }#yes this finally works!!!!
DF.permutations.CO.no_rank <- as.data.frame(DF.permutations.CO.no_rank)

#
for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
  }

#merge with real
#co.var.for.plot <- matrix(ncol=3, nrow=72)
#co.var.for.plot <- as.data.frame(co.var.for.plot)
#colnames(co.var.for.plot) <- c("cell.id","var","type")

#co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
#co.var.for.plot$type[1:24] <- "sim"

#co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
#co.var.for.plot$type[25:48] <- "real"

#co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
#co.var.for.plot$type[49:72] <- "sim_no_rank"

#plot
#whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
#whole.cell.var_CO.plot

```

## CO count, male

Make list of DFs to fill with permutations

```{r list.of.DF, echo=FALSE}

#dataframes to recreate -- and fill with 
#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)

#make subsampled piece of DF here -- these are taking the whole DF
#I will just take single values
#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
#male.whole.cell
#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)

#rank (sample all of the ranks then stich them together)
#replication should match the num
BIG.LIST_w.rank <- replicate(18,
  list(bind_rows(list( 
  male.whole.cell[ sample( which(male.whole.cell$rank==1), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==2), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==3), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==4), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==5), 1, replace = FALSE ),],
   male.whole.cell[ sample( which(male.whole.cell$rank==6), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==7), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==8), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==9), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==10), 1, replace = FALSE ),],

   male.whole.cell[ sample( which(male.whole.cell$rank==11), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==12), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==13), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==14), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==15), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==16), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==17), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==18), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==19), 1, replace = FALSE ),]
))))
#this is replicating single cell (NOT a SAMPLE of 24)

#LIST of DF, no rank (include all), samples all 19 bivs at a time
#makes a list of 10 -- [[]]
BIG.DF.LIST_no.rank <- replicate(18,
                  list(bind_rows(list( 
          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
) ) ) )
```

```{r sum.stat, echo=FALSE}
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

#Calq cV for the rank perm
cv.val.DF_rank <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  cv.val.DF_rank$cV[j] <-cv(  BIG.LIST_w.rank[[j]]$hand.foci.count )
    cv.val.DF_rank$CO[j] <-sum(  BIG.LIST_w.rank[[j]]$hand.foci.count )

    }

#Calq cV for no rank
cv.val.DF_no.rank <- data.frame(  cV_CO = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  cv.val.DF_no.rank$cV_CO[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
   cv.val.DF_no.rank$CO[i] <- sum(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
  }
#extract the 
#next step make cell-level calqs of cV -- for plotting

#merge + plots of the cvs
#10 and 10 rows from the permuted
cv.val.DF_rank$type <- "sim w.rank"
colnames(cv.val.DF_rank) <- c("cV", "CO", "type" )

cv.val.DF_no.rank$type <- "sim no rank"
colnames(cv.val.DF_no.rank) <- c("cV", "CO", "type")

#then add the real values
real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count,male.biv.count_table$CO )
real.vals.cv.co$type <- "real"
colnames(real.vals.cv.co) <- c("cV","CO", "type")

#merge
cVs_permutations.DF <- rbind(cv.val.DF_rank, cv.val.DF_no.rank, real.vals.cv.co)

#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")

#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")

```

```{r perm.plots, echo=FALSE}
#plots for permutations

# need to melt the DFs before ploting?
#ploting --- plot with more sims
oo.plot <- ggplot(cVs_permutations.DF, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")

Mus_hist <- ggplot(cVs_permutations.DF, aes(x = CO)) + geom_histogram(aes(fill=type), binwidth = 2 )+ggtitle("COs per cell, Mus male")

HS.density.plot <- ggplot(cVs_permutations.DF, aes(x = CO, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Mus male")+xlim(c(0,100))
HS.density.plot

```

Above plot shows overdispersion of CO counts per cell for my Mus male dataset.
The above plot the real data might have more variance because it's a combination of strains

## CO count, female

```{r rand.DF_CO.count.female, echo=FALSE}

#dataframes to recreate -- and fill with 
#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)

#make subsampled piece of DF here -- these are taking the whole DF
#I will just take single values
#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
#male.whole.cell
#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)

#rank (sample all of the ranks then stich them together)
BIG.LIST_w.rank_female <- replicate(24,
  list(bind_rows(list( 
  female.whole.cell[ sample( which(female.whole.cell$rank==1), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==2), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==3), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==4), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==5), 1, replace = FALSE ),],
   female.whole.cell[ sample( which(female.whole.cell$rank==6), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==7), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==8), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==9), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==10), 1, replace = FALSE ),],

   female.whole.cell[ sample( which(female.whole.cell$rank==11), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==12), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==13), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==14), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==15), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==16), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==17), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==18), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==19), 1, replace = FALSE ),],
  female.whole.cell[ sample( which(female.whole.cell$rank==20), 1, replace = FALSE ),]
  ))))

#LIST of DF, no rank (include all), samples all 19 bivs at a time
BIG.DF.LIST_no.rank_female <- replicate(24,
                  list(bind_rows(list( 
          female.whole.cell[ sample( which( female.whole.cell$rank < 20), 20, replace = FALSE ),]
) ) ) )

```

These permuted dataframes can be reused for all the meiotic traits, (CO.count, SC length, and position).

```{r sum.stat_cv.CO, echo=FALSE}
#CALUCULATING SUMMARY STAT from the permuted DFs
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

#Calq cV for the rank perm
cv.val.DF_rank.female <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  cv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$hand.foci.count ) #for each
  cv.val.DF_rank.female$CO[j] <-sum(  BIG.LIST_w.rank_female[[j]]$hand.foci.count )
  }

#Calq cV for no rank
cv.val.DF_no.rank.female <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  cv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count )
  cv.val.DF_no.rank.female$CO[i] <-sum(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count )
}

#merge + plots of the cvs
#10 and 10 rows from the permuted
cv.val.DF_rank.female$type <- "sim w.rank"
colnames(cv.val.DF_rank) <- c("cV","CO","type" )

cv.val.DF_no.rank.female$type <- "sim no rank"
colnames(cv.val.DF_no.rank.female) <- c("cV","CO", "type")

#then add the real values
real.vals.cv.co_female <- data.frame(cv = female.biv.count_table$cv_co.count, CO = female.biv.count_table$CO)
real.vals.cv.co_female$type <- "real"
colnames(real.vals.cv.co_female) <- c("cV","CO", "type")

#merge
cVs_permutations.DF_female <- rbind(cv.val.DF_rank.female, cv.val.DF_no.rank.female, real.vals.cv.co_female)

# need to melt the DFs before ploting?
#ploting --- plot with more sims
Femaleplot_cV <- ggplot(cVs_permutations.DF_female, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count female, 1000 simulated")

Mus_hist_F <- ggplot(cVs_permutations.DF_female, aes(x = CO)) + geom_histogram(aes(fill=type), binwidth = 2 )+ggtitle("COs per cell, Mus female")

Mus.density.plot_F <- ggplot(cVs_permutations.DF_female, aes(x = CO, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Mus female")+xlim(c(0,50))
Mus.density.plot_F

#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")

#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")

cVs_permutations.DF$sex <- "male"
cVs_permutations.DF_female$sex <- "female"

all.cv <- rbind(cVs_permutations.DF, cVs_permutations.DF_female)
cV_plot <- ggplot(all.cv, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")+facet_wrap(~sex)

```

```{r female.plot, echo=FALSE}

```

Density plot above also shows that the female CO count data is overdispersed compered to a randomized distribution.

# SC length, bivalent level

Start doing the same thing but for bivalent SC lengths.
Use the same randomized data sets!
summary stats - mean SC length, cV SC length, total SC area

don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.

```{r SC_Sum.Stats.male, echo=FALSE}
# calq the sum stats
SC_Mus.male_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  SC_Mus.male_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
  SC_Mus.male_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank[[j]]$SC.length )
  SC_Mus.male_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank[[j]]$SC.length )
  }

#Calq cV for no rank
SC_Mus.male_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SC_Mus.male_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length )
  SC_Mus.male_no.rank$total_SC[i] <-sum(  BIG.DF.LIST_no.rank[[i]]$SC.length )
  SC_Mus.male_no.rank$mean_SC[i] <-mean(  BIG.DF.LIST_no.rank[[i]]$SC.length )
}

#10 and 10 rows from the permuted
SC_Mus.male_w.rank$type <- "sim w.rank"
colnames(SC_Mus.male_w.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )

SC_Mus.male_no.rank$type <- "sim_no.rank"
colnames(SC_Mus.male_no.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )

real_SC.Mus <- data.frame(male.biv.count_table$var_sc,  male.biv.count_table$total.SC, male.biv.count_table$mean_sc )
real_SC.Mus$type <- "real"

colnames(real_SC.Mus) <- c("SC.cV", "tot.SC", "mean.SC", "type")
#merge
SC_Mus_male <- rbind(SC_Mus.male_w.rank, SC_Mus.male_no.rank,real_SC.Mus)

```

```{r SC.plots_Mus.male, echo=FALSE}
Mus_SC_hist <- ggplot(SC_Mus_male, aes(x = tot.SC)) + geom_histogram(aes(fill=type), binwidth = 50 )+ggtitle("total SC per cell, Mus male")

Mus_SC_density <- ggplot(SC_Mus_male, aes(x = tot.SC, fill=type) )+geom_density(alpha=.6) + ggtitle("Total SC per cell, Mus male")+xlim(c(800,2000))
Mus_SC_density

#plot the forms
#plot within cell variance
#whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
```

```{r SC_sum.stats.female, echo=FALSE, eval=FALSE}

# calq the sum stats
SC_Mus.female_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  SC_Mus.female_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank_female[[j]]$SC.length ) #for each
  SC_Mus.female_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank_female[[j]]$SC.length )
  SC_Mus.female_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank_female[[j]]$SC.length )
  }

#Calq cV for no rank
SC_Mus.female_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female.) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SC_Mus.female_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank_female.[[i]]$SC.length )
  SC_Mus.female_no.rank$total_SC[j] <-sum(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
  SC_Mus.female_no.rank$mean_SC[j] <-mean(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
}

```

```{r sum.stat.SClength, echo=FALSE}

#Calq SC cV for the SC length FEMALE
SCcv.val.DF_rank.female <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(j)
  #fill in the cells of given index
  SCcv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$SC.length ) #for each
  }

#Calq SC cV for no rank
SCcv.val.DF_no.rank.female <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SCcv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$SC.length ) #for each 
}

#merge + plots of the cvs
#10 and 10 rows from the permuted
SCcv.val.DF_rank.female$type <- "sim w.rank"
colnames(SCcv.val.DF_rank.female) <- c("cV","type" )

SCcv.val.DF_no.rank.female$type <- "sim no rank"
colnames(SCcv.val.DF_no.rank.female) <- c("cV", "type")

#then add the real values
real.vals.cv.SC_female <- data.frame(female.biv.count_table$cv_SC.length)
real.vals.cv.SC_female$type <- "real"
colnames(real.vals.cv.SC_female) <- c("cV", "type")

#merge and plots
cV.SC_permutations.DF_female <- rbind(SCcv.val.DF_rank.female, SCcv.val.DF_no.rank.female, real.vals.cv.SC_female)

#MALE --BIG.LIST_w.rank, BIG.DF.LIST_no.rank,
#Calq cV for the SC length
SCcv.val.DF_rank.male <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
  #print(j)
  #fill in the cells of given index
  SCcv.val.DF_rank.male$cV[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
  }

#Calq cV for no rank
SCcv.val.DF_no.rank.male <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SCcv.val.DF_no.rank.male$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length ) #for each 
}

#merge + plots of the cvs
#10 and 10 rows from the permuted
SCcv.val.DF_rank.male$type <- "sim w.rank"
colnames(SCcv.val.DF_rank.male) <- c("cV","type" )

SCcv.val.DF_no.rank.male$type <- "sim no rank"
colnames(SCcv.val.DF_no.rank.male) <- c("cV", "type")

#then add the real values-- male.biv.count_table
real.vals.cv.SC_male <- data.frame(male.biv.count_table$cv_SC.length)
real.vals.cv.SC_male$type <- "real"
colnames(real.vals.cv.SC_male) <- c("cV", "type")

#merge and plots
cV.SC_permutations.DF_male <- rbind(SCcv.val.DF_rank.male, SCcv.val.DF_no.rank.male, real.vals.cv.SC_male)
cV.SC_permutations.DF_male$sex <- "male"
cV.SC_permutations.DF_female$sex <- "female"  #cV.SC_permutations.DF_female
all_cv.SC <- rbind(cV.SC_permutations.DF_female, cV.SC_permutations.DF_male)

cV.SC_plot <- ggplot(all_cv.SC, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv of within cell SC lengths")+facet_wrap(~sex)
```

```{r permuted.v.real, echo=FALSE, eval=FALSE}
for( o in 1:nrow(DF.permutations_CO.count) ) {
  DF.permutations_CO.count$var_CO.count[o] <- var(as.numeric(DF.permutations_CO.count[o,2:20]))
  }

#two data frames of within cell var
#cell.id, var, type
var.for.plot <- matrix(ncol=3, nrow=48)
var.for.plot <- as.data.frame(var.for.plot)
colnames(var.for.plot) <- c("cell.id","var","type")

var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
var.for.plot$type[1:24] <- "sim"

var.for.plot$var[24:48] <- male.biv.count_table$var_co
var.for.plot$type[24:48] <- "real"

#my permuted DFs are in a strange str

#plot within cell variance
whole.cell.var.plot  <- ggplot(var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell variance between types")

whole.cell.var.plot
```

# CO Position

don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.

look at CO position metrics.
Do I need to melt the plots 
the position col names are also, messed up I need to fix them --

1. use the same randomized dataframe
2. use the summary stat

How do I randomize CO pos for a chrm?
I want to draw positions from chrm_classes

For each cell observation, note the classes,  I want to 
make bins for positions of all

```{r CO.pos, echo=FALSE}
#male.whole.cell
#female.whole.cell
#fix the data DF for PERMUTING CO

whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359

male.biv.count_table <- male.biv.count_table[!(is.na(male.biv.count_table$fileName) | male.biv.count_table$fileName==""), ]

pos.pool_1CO <- male.whole.cell[male.whole.cell$hand.foci.count == 1,]
pos.pool_2CO <- male.whole.cell[male.whole.cell$hand.foci.count == 2,]
pos.pool_3CO <- male.whole.cell[male.whole.cell$hand.foci.count == 3,]

#now make the randomized data set for CO pos
#do whole row -- so I can cen n

random_pos <- data.frame(male.whole.cell$fileName, male.whole.cell$SC.length, male.whole.cell$hand.foci.count, male.whole.cell$Obj.ID  )

#INITIALIZE EMPTY list of DFs
#fill in with full DF

LIST.DFs <- setNames( replicate( length(male.biv.count_table$fileName), data.frame( obj.ID = c())), male.biv.count_table$fileName)

ids <-    male.whole.cell$Obj.ID[which(male.whole.cell$fileName == '14oct16_30jun16_CAST_m3_sp1_3.2')]

LIST.DFs[1]$obj.ID[1:19] <- ids[1:19]

#INITALIZE empty LIST (make DF in the loop)

LIST <- list()

#Loop for rand sampling  
for( male.cell in male.biv.count_table$fileName){

  #for each fileName, make a new DF
  DF.in.loop <- data.frame(matrix(ncol = 1, nrow = 19))
  #making the Within loop DF
  DF.in.loop$X1 <- male.whole.cell$fileName[male.whole.cell$fileName ==male.cell]
  DF.in.loop$obj.ID <- male.whole.cell$Obj.ID[male.whole.cell$fileName ==male.cell]
  DF.in.loop$SC.length <- male.whole.cell$SC.length[male.whole.cell$fileName ==male.cell]
  DF.in.loop$hand.foci.count <- male.whole.cell$hand.foci.count[male.whole.cell$fileName ==male.cell]

  #LIST.DFs[[male.cell]]$obj.ID[1:19] <- male.whole.cell$Obj.ID[male.whole.cell$fileName == male.cell]
    #print(male.cell)
    n0CO = male.biv.count_table$nCO0[male.biv.count_table$fileName == male.cell] #this reference needs to be from male.biv.count_table
    n1CO = male.biv.count_table$nCO1[male.biv.count_table$fileName == male.cell]
    n2CO = male.biv.count_table$nCO2[male.biv.count_table$fileName == male.cell]
    n3CO = male.biv.count_table$nCO3[male.biv.count_table$fileName == male.cell]

   #rand samples for foci.pos
    #JUST DO THIS FOR 1CO bivs
    DF.in.loop$rand_F1
    sample.n1CO <- sample(pos.pool_1CO$Foci1.PER, n1CO, replace = FALSE )  #sample from the cell.table

    #sample a random 2CO chrm (obj.ID),    #doesn't make sense to sample a random position from 2CO pool, need to sample the linked positions, or get the 
    #chrm observation
   sample.n2CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n2CO, replace = FALSE )
   sample.n3CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n3CO, replace = FALSE )

   #MAP normalized positions onto SC lengths/chrms of the fileName cell obs
   #map the 1CO rand Foci position onto 1CO SCLength

   #for i in n1CO
#   which(random_pos$fileName == male.cell)
   random_pos$Rand.F1[which(random_pos$fileName == male.cell)][i] <- sample(pos.pool_1CO$Foci1.PER, 1, replace = FALSE)

   #list of the chrm lengths for all the 1COs for the current cell obs
   #SC.Length_1COs <-

}

#need to define number of chrm classes within each nuc
n1CO = 9

BIG.DF.LIST_COpos_no.rank <- replicate(10,
                  list(bind_rows(list(

  pos.pool_1CO[ sample( which( pos.pool_1CO$hand.foci.count == 1), n1CO, replace = FALSE ),]

) ) ) )

BIG.DF.LIST_no.rank <- replicate(18,
                  list(bind_rows(list( 
          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
) ) ) )

```

```{r sum.stat.CO.pos, echo=FALSE}
#remove F1 without values
#and 1CO with 2 values

SC_Mus.male_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  SC_Mus.male_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
  SC_Mus.male_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank[[j]]$SC.length )
  SC_Mus.male_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank[[j]]$SC.length )
  }

#Calq cV for no rank
SC_Mus.male_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SC_Mus.male_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length )
  SC_Mus.male_no.rank$total_SC[i] <-sum(  BIG.DF.LIST_no.rank[[i]]$SC.length )
  SC_Mus.male_no.rank$mean_SC[i] <-mean(  BIG.DF.LIST_no.rank[[i]]$SC.length )
}

```

## Adding to Data set

how to best increase the size of the dataset, with my bivalent data. From the curated BivData, figure out which can be added / into the whole cell data. I think there might only be a couple cell to add in.

```{r getting.cell.names, echo=FALSE,eval=FALSE}

#load in the cell dataset
#28feb15_PWD_f2  from Curated_BivData
#for a mouse, pull the 10 highest rec cells

PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
#"Foci3"                   "curation.notes"  
female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]

write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)

PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")

#merge with hand data
#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)

cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
                            .(fileName),
                            summarize, 
                            ncells = length(unique(fileName)),
                            nbivs = length(unique(Obj.ID)),
                            n.foci = sum(hand.foci.count),
                            total.SC = sum(chromosomeLength),
                            mean.SC = mean(chromosomeLength)

)
#add hand data
ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]

#I have 3 cells with full 20

```

```{r mini.biv, echo=FALSE,eval=FALSE}
#make mini biv plots

#PWD.female
#PWD.female_hand.measure

cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]

two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))

#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
#of SC.
MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
 #geom_hline(yintercept = c(male_IFD) ) +

  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
  #how to add points for 2foci
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
  #foci.1.position, two_groups$distal.foci.position)
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +

#add centromeres...  
  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+

  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
 # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
  ylim(c(0,200))+
     theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())  #for removeing x labels

#MiniBiv_plot1
#MiniBiv_plot2

#save the files

```

```{r discribe.Biv.Pattern, echo=FALSE}
#start discribing the basic

#start merging curated bivData with handmeasured data

#BivData - Curated
```

```{r add.more.whole.cell, echo=FALSE}

#load data curatedBiv Data
#load in hand measures

```

Look/use the DF, Curated_BivData data frame for mice with many cell measures and cells with whole cell measures (all bivalents have been measured).

Mice with many cells and BivData
17mar16_G_f4 - BivData
17mar16_G_f1

Chunk below has code for choosing mice with a good number of cells (25?)

```{r choosing.mice, echo=FALSE, eval=FALSE}
#Choose mice with many cell observations 
#important for increased power for within mouse variance

#G_f, G_m
#SKIVE_f, SKIVE_m #have BivData

#remove all the quality 5

#make plot/histogram of the MLH1 count distribution
counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]

mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]

ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
topsy <- ordered.MLH1[1:5,]
bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
top.n.bottom.bin <- rbind(topsy,bottomzy)

#write 
write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)

shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
#highest of the low bin
#lowest value of the high bin

#get the filenames for this groups of cells

#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
#28feb15_PWD_f2, in PWD--female .xlsx file

#31may19_SKIVE_m2
male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
male.one.DF <- male.one.DF[male.one.DF$quality < 5,]

ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
male.high.bin.DF <- ordered.MLH1.male[1:5,]
male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
#write 
write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)

shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz.too

```

the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
Should I be comparing cells with the same number of COs?

# Junk

```{r junk, eval=FALSE, echo=FALSE}
DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),
                       rank3= replicate(24, ""),     rank4= replicate(24, ""),
                       rank5= replicate(24, ""),   rank6= replicate(24, ""),
                       rank7= replicate(24, ""), rank8= replicate(24, ""),
                       rank9= replicate(24, ""), rank10= replicate(24, ""),
                       rank11= replicate(24, ""),     rank12= replicate(24, ""),
                       rank13= replicate(24, ""),     rank14= replicate(24, ""),
                       rank15= replicate(24, ""),     rank16= replicate(24, ""),
                       rank17= replicate(24, ""),rank18= replicate(24, ""),rank19= replicate(24, "")
                       )

#not sure how to make the initial DF
#replicate 24 rows with 19 cols
iterations = 24
variables = 20
DF.permutations.sc <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

#permutations -- for SC

#permutations without accounting for rank
its = 24
varis = 20
DF.permutations.no_rank <- matrix(ncol=varis, nrow=its)
#colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
#                               "rank6","rank7","rank8","rank9","rank10","rank11",
#                               "rank12","rank13","rank14","rank15","rank16",
#                               "rank17","rank18","rank19" )

#ROW IS FIRST
for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.sc) ){
  rand.rank.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.sc[row.num,col.num] <- rand.rank.sc
   }
  }#yes this finally works!!!!
DF.permutations.sc <- as.data.frame(DF.permutations.sc)

#pair with real within cell SC variance
for( o in 1:nrow(DF.permutations.sc) ) {
  DF.permutations.sc$var_sc[o] <- var(as.numeric(DF.permutations.sc[o,2:20]))
  }

for(row.num in 1:24){
  #print(c(row.num, "rowis") )
  for(col.num in 2:ncol(DF.permutations.no_rank) ){
  rand.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
  #print(SCll)   
  DF.permutations.no_rank[row.num,col.num] <- rand.sc
   }
  }#yes this finally works!!!!
DF.permutations.no_rank <- as.data.frame(DF.permutations.no_rank)

for( o in 1:nrow(DF.permutations.sc) ) {
  DF.permutations.no_rank$var_rand.sc[o] <- var(as.numeric(DF.permutations.no_rank[o,2:20]))
  }

```