### 2019-12-17 doc/figureFiles/figures.Rmd
working rmd files for drafts of main figures

---
title: "Chapter 2 Figures"
author: "April Peterson"
date: "December 16, 2019"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, include=FALSE}
#load data ect
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(dplyr)
library(lattice)
library(dplyr)
library("cowplot")
library(raster)#for cV
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

```{r strain.color, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= 'indianred4',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

```

```{r setting.data, echo=FALSE, warning=FALSE}
#remove F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "F1"),]

#remove outgroups
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPRET"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPIC"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "CAROLI"),]

#quality control - remove outlier mice
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$mouse != "18nov17_WSB_f5"),] #this mouse only had 1 cell

#I need cell level data

uni.mice <- unique(AP_mouse_table.HQ$mouse)
fig.data<- MLH1_data[ MLH1_data$mouse %in% uni.mice, ]

#take all mouse levels from AP.table,
#subset MLH1 data with those mice
```

```{r fig1_scatter_basic, echo=FALSE, warning=FALSE}
#mouse averages
#AP_mouse_table.HQ
#fig.data  cell data

fig.data <- fig.data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
fig.data$mouse <- factor(fig.data$mouse, levels=unique(fig.data$mouse))

fig.data.male <- subset(fig.data, sex== "male")

#male
Fig1_A_mouse.male <- ggplot(fig.data.male[order(fig.data.male$strain), ],
  aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+  #geom_jitter()+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+geom_boxplot(alpha = 0.6)+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),

  #lines -- these might be tricky with the faceting
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())  +
  #Dom
  annotate("segment", x= 10, xend=20.5, y=15, yend=15, colour="black", size=1)+
  #Musc
  annotate("segment", x= 50, xend=55, y=15, yend=15, colour="black", size=1)+
  #Cast
  annotate("segment", x=70, xend=80, y=15, yend=15, colour="black", size=1)

#female
#the data is mouse mean 
fig.data.female <- subset(fig.data, sex== "female")

Fig1_A_mouse.female <- ggplot(fig.data.female[order(fig.data.female$strain), ], aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+ #geom_jitter(width = 0.25)+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #lines -- these might be tricky with the faceting
 axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  #Dom
  annotate("segment", x= 10, xend=20.5, y=18, yend=18, colour="black", size=1)+
  #Musc
  annotate("segment", x= 50, xend=55, y=18, yend=18, colour="black", size=1)+
  #Cast
  annotate("segment", x=70, xend=80, y=18, yend=18, colour="black", size=1)

```

# Figure 1

```{r strain.table, echo=FALSE, warning=FALSE}
#fig.data -- (only mice from better list)

#there are still quality 5 cells in this dataframe
fig.data <- subset(fig.data, quality < 5)

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Nmice), 3)
    )

#cv, sd and se are all NA / empty
strain.table.HQ <- strain.table.HQ[!(is.na(strain.table.HQ$cV) | strain.table.HQ$cV== "" ), ]

#remove outgroups
#remove F1
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "F1"),]

#remove outgroups
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPRET"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPIC"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAROLI"),]

```

```{r reorder.strain, echo=FALSE}

#re-order strains
#Musc
#PWD, SKIVE, KAZ, CZECH, TOM, AST

#MOl
#MSM, MOLF
strain.table.HQ$strain<- factor(strain.table.HQ$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                                                                   "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                                  "MSM", "MOLF",
                                                                    "CAST", "HMI",
                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
```

I think some strain observations can be removed, since they only have

```{r strain.table.plot, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_B_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male

  #segments
  annotate("segment", x= 0, xend=5, y=21, yend=21, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5, xend=8, y=20, yend=20, colour="black", size=1)+
  #Cast
  annotate("segment", x=11, xend=15, y=18, yend=18, colour="black", size=1)+
  #label
  annotate("text", x = c(1.5,8,10.5, 11), #list of all x points 
           y = c(20,18, 20,18),

           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus", "M.m.castaneus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

# 2
Fig1_B_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=1, xend=3.5, y=22, yend=22, colour="black", size=1)+
  #Musc
  annotate("segment", x=5, xend=8.5, y=20, yend=20, colour="black", size=1)+
  #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(1.5,8), #list of all x points 
           y = c(20,18),
           label = c("M. m. domesticus","M. m. musculus"),
           fontface = 'italic', hjust = 0.5, size=3)

```

```{r Strain.fig.merge, echo=FALSE}
fig1_B  <- ggdraw() +
  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.48, height=.88) +
  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.48, height=.88) +

  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )
fig1_B
```

add molossinus

re-arrange Muusc strain level order to highlight the polymorphism to musc

Clean up segments
There are many more male strains represented.

The male plot is squished because of the legend

## Figure 1, mouse levels shown

```{r cow.plot.fig1, echo=FALSE}

fig1  <- ggdraw() +
  draw_plot(Fig1_A_mouse.female, x=0, y = 0, width=.48, height=.88) +
  draw_plot(Fig1_A_mouse.male, x=.5, y=0, width=.48, height=.88) +
  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )

fig1
```

USE boxplots --- this is a bit too busy with single mouse level observations. (so forget the cell level data I just coded. What if I sub-sampled?)

Draft first fig. MLH1 distributions will be the main first figure
I think a strain level would look better.  Add white space between

I still need to work on some colors
add line segments, they are good at anchoring

add line segments for strain means?, do boxplots?

Note that this figure knitted in html will look different than a saved png or tif.
change the ylabel 'Mean Number of MLH1 Foci per Cell'