### 2019-12-18 doc/figureFiles/figures.Rmd
started v2 of main figure

---
title: "Chapter 2 Figures"
author: "April Peterson"
date: "December 16, 2019"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, include=FALSE}
#load data ect
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(dplyr)
library(lattice)
library(dplyr)
library("cowplot")
library(raster)#for cV

#setwd("~./MLH1repo/")
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

```{r strain.color, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= 'indianred4',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

```

```{r setting.data, echo=FALSE, warning=FALSE}
#remove F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "F1"),]

#remove outgroups
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPRET"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPIC"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "CAROLI"),]

#quality control - remove outlier mice
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$mouse != "18nov17_WSB_f5"),] #this mouse only had 1 cell

#remove male only data

#I need cell level data 
uni.mice <- unique(AP_mouse_table.HQ$mouse)
fig.data<- MLH1_data[ MLH1_data$mouse %in% uni.mice, ]

#take all mouse levels from AP.table,
#subset MLH1 data with those mice
```

# Option 1, strain averages, sexes matched

```{r strain.table.v2, echo=FALSE, warning=FALSE}

#there are still quality 5 cells in this dataframe
fig.data <- subset(fig.data, quality < 5)

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Nmice), 3)
    )

#cv, sd and se are all NA / empty
#this removes MOLF female
#strain.table.HQ <- strain.table.HQ[!(is.na(strain.table.HQ$cV) | strain.table.HQ$cV== "" ), ]

#remove outgroups
#remove F1
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "F1"),]

#remove outgroups
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPRET"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPIC"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAROLI"),]

#REMOVE ALL STRAINS WITHOUT FEMALE DATA
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]

strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "PERC"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "TOM"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "AST"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CZECH"),]

strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]

```

```{r reorder.strain.v2, echo=FALSE}

#re-order strains
#Musc
#PWD, SKIVE, KAZ, CZECH, TOM, AST

#MOl
#MSM, MOLF
strain.table.HQ$strain<- factor(strain.table.HQ$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                                                                   "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                                  "MSM", "MOLF",
                                                                    "CAST", "HMI",
                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
```

```{r strain.table.plot.v2, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_Bv2_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #Dom
  annotate("segment", x=.75, xend=3.6, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=3.9, xend=6.15, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=6.5, xend=8.1, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(2,5,7.15), #list of all x points 
           y = c(18.5,18.5, 18.5),

   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

# 2
Fig1_Bv2_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
#mol
 annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
    #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
           y = c(18.5,18.5,18.5),
           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

```

```{r Strain.fig.merge.v2, echo=FALSE}

fig1_B_v2  <- ggdraw() +
  draw_plot(Fig1_Bv2_strain.female, x=0, y = 0, width=.5, height=.97) +
  draw_plot(Fig1_Bv2_strain.male, x=.45, y=0, width=.5, height=.97) +

  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .45), y = c(1, 1) )
fig1_B_v2
```

Molf female data point is missing because there is only 1 female, so NAs were generated.
Mol text should be moved back above the points.

## option 2, strain averages, more strains

```{r strain.table, echo=FALSE, warning=FALSE}
#fig.data -- (only mice from better list)

#there are still quality 5 cells in this dataframe
fig.data <- subset(fig.data, quality < 5)

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ.backup <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Nmice), 3)
    )

#cv, sd and se are all NA / empty
#strain.table.HQ.backup <- strain.table.HQ.backup[!(is.na(strain.table.HQ.backup$cV) | strain.table.HQ.backup$cV== "" ), ]

#this removes MOLF female

#remove outgroups
#remove F1
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "F1"),]

#remove outgroups
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "SPRET"),]
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "SPIC"),]
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "CAROLI"),]

strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "CAST"),]
strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "HMI"),]

#Keep more of the strains
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "PERC"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "TOM"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "AST"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CZECH"),]

#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]

```

```{r reorder.strain, echo=FALSE}

#re-order strains
#Musc
#PWD, SKIVE, KAZ, CZECH, TOM, AST

#MOl
#MSM, MOLF
strain.table.HQ.backup$strain<- factor(strain.table.HQ.backup$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                           "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                    "MSM", "MOLF",
                                          "CAST", "HMI",
                             "SPRET", "SPIC", "CAROLI", "F1", "other") )
```

```{r strain.table.plot, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_B_strain.male <- ggplot(strain.table.HQ.backup[strain.table.HQ.backup$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10), #list of all x points 
           y = c(18.5,18.5,33),

   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

# 2
Fig1_B_strain.female <- ggplot(strain.table.HQ.backup[strain.table.HQ.backup$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
#mol
 annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
    #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
           y = c(18.5,18.5,33),
           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

```

```{r Strain.fig.merge, echo=FALSE}
fig1_B_v1  <- ggdraw() +
  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.5, height=.97) +
  draw_plot(Fig1_B_strain.male, x=.45, y=0, width=.5, height=.97) +

  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )
fig1_B_v1
```

This is currently backup version.

I removed the cast and outgroup measures. I left off the color legend, but I think I could add one to the bottom of the figure. These show the strain averages, but I can try adding the mouse averages, if you think having mulitple points would show the data better.

## Option 3, mouse boxplots

```{r fig1_scatter_basic, echo=FALSE, warning=FALSE}
#mouse averages
#AP_mouse_table.HQ
#fig.data  cell data

fig.data <- fig.data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
fig.data$mouse <- factor(fig.data$mouse, levels=unique(fig.data$mouse))

fig.data.male <- subset(fig.data, sex== "male")

#male
Fig1_A_mouse.male <- ggplot(fig.data.male[order(fig.data.male$strain), ],
  aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+  #geom_jitter()+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+geom_boxplot(alpha = 0.6)+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #lines -- these might be tricky with the faceting
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
  #Dom
#  annotate("segment", x= 10, xend=20.5, y=15, yend=15, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 50, xend=55, y=15, yend=15, colour="black", size=1)+
  #Cast
#  annotate("segment", x=70, xend=80, y=15, yend=15, colour="black", size=1)

#female
#the data is mouse mean 
fig.data.female <- subset(fig.data, sex== "female")

Fig1_A_mouse.female <- ggplot(fig.data.female[order(fig.data.female$strain), ], aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+ #geom_jitter(width = 0.25)+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #lines -- these might be tricky with the faceting
 axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
  #Dom
#  annotate("segment", x= 10, xend=20.5, y=18, yend=18, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 50, xend=55, y=18, yend=18, colour="black", size=1)+
  #Cast
#  annotate("segment", x=70, xend=80, y=18, yend=18, colour="black", size=1)

```

```{r cow.plot.fig1, echo=FALSE}

fig1  <- ggdraw() +
  draw_plot(Fig1_A_mouse.female, x=0, y = 0, width=.48, height=.88) +
  draw_plot(Fig1_A_mouse.male, x=.48, y=0, width=.48, height=.88) +
  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )

fig1
```

This combination of points and box plots is similar to the final peromyscus figure 1, but the number of individual mice observations makes it very busy.

This might be a good format if fewer mice were used for each strain.