### 2019-12-27 doc/Effects.Rmd
re-starting effect quan measurements

---
title: "Effects"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

library(knitr)
library(ggplot2)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)

#load main data file

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")

```

```{r Effects.on.the.dataset, eval=FALSE, echo=FALSE}

#*age effects* (male, maternal age)

#*litter effect*

#*mouse room effect* (Charmony, MSC)

#*quantification repeatability*

#*spread batch and staining effects*

#*power analysis on number of cells*

#*time of year effect (on spread)*

#*dominant effect* (F1s)

```

## Quality Effects

below is a chunk to looking at histogram/distribution and means for the MLH1 data across quality bins

```{r MLH1.histogram.2, echo=FALSE}
#remake the pooled MLH1 histgoram plot from above but facet by quality

#PWD male, MSM male
#LEW female, WSB female
mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
                 .(category, quality),
                 summarize, 
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
#what's the point of the above table? -- add vlines?

#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect

histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
histM_qual.KAZ_male

histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
histM_qual.PWD_male

histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
histM_qual.MSM_male

histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
histM_qual.WSB_female

histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
histM_qual.LEW_female

#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")

#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))

#ad
#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

```

I'm not sure sure how effective the above plots are.

### Age Effect (male)

Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age].

The initial analysis uses plots. This should be double checked with a general linear model as well.

```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#make Mouse meta data work .. rename these
MouseMetaData.wrkin <- meta.data[,c(1,4)]
colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
#do this for the mouse table instead of the MLH1 table
MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
#DOB is date
MLH1.imprv <- add_euth_date(MLH1.imprv)
#good euth date..

#make backup date col #add euth date for calculating age
MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
MLH1.imprv <- add_age(MLH1.imprv)
#age works

#remove mice that are older than 20 weeks
MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]

age.hist <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
age.hist

```

I like this plot and find it effective. Can I do a similar set up for other factors (quality, ncells?)

```{r age.effect.model, include=FALSE}
#run a age effect linear regression!

male.data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.y == "male",]

male.data.dom <- male.data[male.data$subsp.y == "Dom",]

male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST"),]

male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "MOLF" | male.data$strain.x == "SKIVE"),]

male.data.outgroup <- male.data[(male.data$subsp.y == "Spretus" | male.data$subsp.y == "Spic" | male.data$subsp.y == "Caroli"),]

lm.me.dom  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.dom)

lm.me.musc.low  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.low)

lm.me.musc.high  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.high)
#significant differences across strains in musc high

lm.me.OG  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.outgroup)

summary(lm.me.OG)
#for all models, mostly it was that the intercept was significant, and mot much else

#good knitting output
#as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )

```

```{r age.effect.plot, warning=FALSE, echo=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
#remove NAs
AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]

#make sure the right dataset 
male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
#lots of NA rows... ugh .. remove them
male.data <- male.data[!(is.na(male.data$Nmice) | male.data$Nmice==""), ]

#other removed, no points,
male.data <- male.data[male.data$Ncells > 10,]

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
age.effect.plot
#no longer has 'other', #yay I added purtty standard colors
#there is an NA
#todo add line to strain.y

```

While many slopes are different than 0, I am not convinced there is a clear age effect. Most strains don't have more than 2 ages. I will double check some of these mice and check a general linear model.

### Age Effect (female)

Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.

```{r female.age, eval=FALSE}
#compile this dataset
```

### Maternal Age effect

Check if there is a maternal age effect for females and males.

```{r mat.age.effect, echo=FALSE, eval=FALSE}
#pause evaluation -- needs debugging
meta.data.file = read.csv("~./MLH1repo/data/DandM_8.30.19.csv", header = TRUE)

meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
#445? -- probably extras at bottom not good

#how many mice have maternal DOB
meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
#151

#format this meta.data.file
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

meta.data.file.w.mat.DOB <- add_mouse(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_category(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_strain(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_sex(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_subsp(meta.data.file.w.mat.DOB)

#check if date format's been applied before tryung the age functions
#str(meta.data.file.w.mat.DOB)

#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")

#str(meta.data.file.w.mat.DOB)

meta.data.file.w.mat.DOB <- add_euth_date(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_age(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_mat_age(meta.data.file.w.mat.DOB)

#merge meta data with MLH1 -- rename this
see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
#remove anything without mean_CO
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
#what did I do with this DF...?

#visualize the data
meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]

#hist of data
mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
mat.age.hist.female

mat.age.hist.male <- ggplot(meta.data.file.w.mat.DOB.male, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of males")
mat.age.hist.male

#remove mice that might have low quality
see.whatcha.do.HQ <- see.whatcha.do[see.whatcha.do$Ncells > 14,]

#age by mean_co (these aren't working)
mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
mat.age.effect.plot.female

mat.age.effect.plot.male <- ggplot(see.whatcha.do.HQ[see.whatcha.do.HQ$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")+facet_wrap(~subsp.y)

#
ALLmat.age.effect.plot.male <- ggplot(see.whatcha.do[see.whatcha.do$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")
+facet_wrap(~subsp.y)

#is there a correlation with proband age and maternal age?
#there is a big lack of maternal age data...
#some of the values reference the original breeding pair..
#from notes on original BPs(males), assigning maternal ages

```

### Litter effect

Compare female spreads made on the same day, but from different mothers/litters.
```{r litter.effect, eval=FALSE}
#compile this data set

```

# Dissection Effect

# Spread and Stain Effects

> *month dissected*

For mice strains that have a similar MLH1 level (control for strain MLH1 effect), check if a month effect for MLH1 count (as seen in Peromyscus).

> *staining batches*

Compare observations from the same mouse, but different MLH1 stains.

```{r stain.effects, eval=FALSE}

#compile this dataset

```

# Room effect

```{r ROOM.effect, eval=FALSE}

#COMPILE THIS datasets
#3 different rooms used to house mice. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD

```