### 2020-02-03 src/BivData_Setup.Rmd

---
title: "BivData Setup"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

# Data setup and basic cleaning

- remove BivData! (these are all Peromyscus!!) (not sure where this DF comes from)

calq the nromalized.siscoten, double check the negative values; 13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14, 27aug15_12jun15_WSB_m2_sp1_18_rev_18, 21nov17_16nov17_MSM_f3_sp1_3_rev_1

foci.pass = 2  5aug17_14jul17_MSM_f1_sp1_13_rev_3

hand foci =0 but there is a foci position

too high hand foci count

declare the Strain specific BivData_DFs  (for convinice)

check the numbers of bivalents with a SC length, but no hand.foci.count

add mean.telomere, mean cent.dist and mean sis.coten to the mouse level table

update IFD functions, to add IFD2,3 -- report the summaries

format manual data set so SC.length = Chrmlength (matches the bivData)

add Q1 and Q4 to the cell summary stats table

ToDo

- Go back and clean up the original excel data

- clean the bottom, fill in missing data

- check the order of foci positions for 5aug17_14jul17_LEW_f1_sp1_11_rev_blob4.

edit / clean up the lines with bad hand.foci.count (61)

ToDo:

1.clean up the non-numbered curated values (for the og df)

2.Goal of this doc is to make DFs that are ready for the analysis.

3.remove the 2 main sources of duplicate lines; 6mar16_PWD_m1, 1mar15_PWD_m1

4. decide on the right number of passing curated bivalents per category I should settle for
Justify why the current number is enough -- or subsample -- to random equivilent sized data sets

5. make sure that the saved Rdata has all the right cols and tables for the next analyses

6. consider integrating manual hand / whole cell measures (check for DUPs ect.)

```{r setup, include=FALSE}
library(plyr)
library(lattice)
library(dplyr)
library(ggplot2)
library(reshape2)
library(raster)

setwd("~./MLH1repo/")
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

whole.cell.data_manual.OG <- read.csv("~./MLH1repo/whole_cell_measures.csv")

#BIVDATA MASTER SHEET
new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
#83976
#83009
#80898
#80919
#61084
#60241
#60830

#Clean away all the empty rows
new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]

new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
#make OG copy
new.Curate.FULL.org <- new.Curate.FULL
new.Curate.FULL.org <- add_category(new.Curate.FULL.org)

#write files for basic stats to csv files
#OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)

OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)

#write the curated file
#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_21.1.20.csv", sep=",",
#            row.names = TRUE)
#write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_21.1.20.csv", sep=",",
#            row.names = TRUE)
```

```{r split.up.DF, echo=FALSE}

#new.Curate.FULL to Curated_BivData

#SC PASS
new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#10480 #9819, #9769

#fix these duplicated ones
#remove duplicates
Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151

Curated_BivData <- add_mouse(Curated_BivData)
Curated_BivData <- add_category(Curated_BivData)
Curated_BivData <- add_strain(Curated_BivData)
Curated_BivData <- add_sex(Curated_BivData)
Curated_BivData <- add_subsp(Curated_BivData)
Curated_BivData <- add_species(Curated_BivData)

Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
```

```{r no.Foci.Pos, echo=FALSE}

just.SC.length <- Curated_BivData[is.na(Curated_BivData$hand.foci.count),]#650 lines

#table(just.SC.length$category)
#decide what to do with these lines
#how will they be represented in Mat Methods?
```

```{r input.check,echo=FALSE}

#check for mistakes that might have been made when inputing / curation process

#hand foci > 4
#foci pos > SC lengths
#negative PER.foci.pos

#it hink most (ALL) of these came from bug in my function (ugh)

#Curated_BivData <- add_IFD(Curated_BivData)
#negative.IFD <- Curated_BivData[which(Curated_BivData$IFD1_ABS < 0),]
#caught 2
#15mar16_16jan16_G_f1_sp1_4_rev
#27mar16_6mar16_G_m3_sp1_12_rev

#neg Nrm F1
#negative.IFD2 <- Curated_BivData[which(Curated_BivData$PER_Foci_1 < 0),]
#0 really?
```

Main files from this plot:

  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures

  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information

  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)

  4. MELT_ABS: make melt version of foci positions for abs measuers

  5. CELL and

  6. mouse table

  7. **Strain DFs** for Het models

  8. **male Subsp DFs** for Polymorphism models

```{r CLEAN.UP.1, echo=FALSE, include=FALSE}

#check for duplicates, and track them down
anyDuplicated(new.Curate.FULL$Obj.ID)  #5335, #3467

dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
dup.curated <- add_mouse(dup.curated)

#table(dup.curated$mouse)
#1mar15_PWD_m1

#which lines are duplicated
#lots of PWD_female duplicated -- probably how I put together the curated files

#new.Curate.FULL <- new.Curate.FULL.org[(new.Curate.FULL.org$SC.pass == 1 | new.Curate.FULL.org$SC.pass == 0),]
#new.Curate.FULL <- droplevels(new.Curate.FULL)

#new.Curate.FULL <- add_mouse(new.Curate.FULL)
#new.Curate.FULL <- add_category(new.Curate.FULL)

#pass.fail.table <- table(new.Curate.FULL$SC.pass, new.Curate.FULL$category)
#this outsput shows the relative ratios of pass to fail biv segmentation

```

```{r remove.mice, echo=FALSE, eval=FALSE}

#mice to remove from the Curated_BivData -- should be similar to the mice exclude from the MLH1

#poor quality?

#reaxamine these mice for if they should be included in 
exclude_mice_list_MLH1 <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",

#    Musc males
    "27aug18_CZECH_m1","8jun18_CZECH_m1",

   "3sep18_AST_m1",

    "3jul18_TOM_m1","3sep18_TOM_m1",

                "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
#REaxamine and rename to exclude mice_BivData

#Curated_BivData <- Curated_BivData[ ! Curated_BivData$mouse %in% exclude_mice_list, ]

exmice.Curated_BivData <- Curated_BivData[ Curated_BivData$mouse %in% exclude_mice_list, ]

#juvi males
#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
#juvi.mice_DF <- juvi.mice[!(is.na(juvi.mice$mouse) | juvi.mice$mouse==""), ]

#old males

```

I'm considering removing the same mice from the MLH1 counts, from the BivData, but this might -- limit the comparisons -- might be too stict. IF foci are bright and clear, SC/bivalents intact -- include.

IF it seems like primary reason for excluding is low cell number include bivalents

Pro remove:

- more similarity across data sets

Con remove:

- the reasons for poor quality of the exclude mice in above list, might be from 2-3 bad bivalents -- which are removed from the biv data during the curation process.

```{r set.type, echo=FALSE}
##change factor cols to numbers (factor -> chr -> number)
Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)

Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)

Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)

##Numeric
Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)

Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)

Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
#I think the number of sig figs and rounding changes slightly from the blocks above

#adding more feature calculations (IFD, ect)
Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength

#IFD.2 does mulitple IFDs (just absolute)
#Curated_BivData <- add_IFD.2(Curated_BivData)

Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength

Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)

#not working
Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength

#add telomere.dist
Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )

Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength

#sis.co.ten
Curated_BivData <- add_SisCoTen(Curated_BivData)

```

Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.

- HetC_MM  : subset only mice

IFD functions

  - IFD: currently I think just does 2CO bivalents

  - IFD.2: for mulitple IFDs (3CO ect.)

## Strain specific data sets

Compile dataframes for both sexes of each strain.

```{r WSB.SC, echo=FALSE}
#make the dataset
WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)

#plot the differences
wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
wsb.ploty

#make a table of the means and n ect
wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r LEW.SC, echo=FALSE}
#make the dataset
LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)

#plot the differences
LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

LEW.ploty
#remove the NAs

#make a table of the means and n ect
LEW.SC.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r G.SC, echo=FALSE}
#make the dataset
G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)

#plot the differences
G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

G.ploty
#remove the NAs

#make a table of the means and n ect
G.SC.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r PWD.SC, echo=FALSE}
#make the dataset
PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)

#plot the differences
PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

PWD.ploty
#remove the NAs

#make a table of the means and n ect
PWD.SC.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r KAZ.SC, echo=FALSE}
#make the dataset
KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)

#plot the differences
KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

KAZ.ploty
#remove the NAs

#make a table of the means and n ect
KAZ.SC.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r SKIVE.Sc, echo=FALSE}
#make the dataset
SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)

#plot the differences
SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

SKIVE.ploty
#remove the NAs

#make a table of the means and n ect
SKIVE.SC.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r MSM.SC, echo=FALSE}
#make the dataset
MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)

#plot the differences
MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

MSM.ploty
#remove the NAs

#make a table of the means and n ect
MSM.SC.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

```{r MOLF.SC, eval=FALSE}
#make the dataset
MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)

#plot the differences
MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

MOLF.ploty
#remove the NAs

#make a table of the means and n ect
MOLF.SC.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

```

## Male specific data set

```{r male.sc.length.strain, echo=FALSE}

#isolate all musc and male

Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == ),]

#make table
SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)

```

# Manual Whole Cell

Cleaning and integrating the dataframes for manual measured data.

```{r whole.cell.manual.clean, echo=FALSE}
#compare and clean up whole cell measures

whole.cell.data_manual <- whole.cell.data_manual.OG[whole.cell.data_manual.OG$chrm.class != "XY",]

colnames(whole.cell.data_manual.OG)
#"return5"         "fileName"        "boxNumber"       "centromere"      "SC.length"  #    
 #[6] "rank"            "chrm.class"      "hand.foci.count" "F1"              "F2"    #         
#[11] "F3"              "notes"           "X"               "X.1"            
#remove, X, X.1
whole.cell.data_manual <- whole.cell.data_manual[-c(1, 13:14)]
whole.cell.data_manual$Foci4 <- ""

#rename cols to match BivData
whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3
whole.cell.data_manual$Foci4 <- whole.cell.data_manual$F4  #currently Foci4 is chr
whole.cell.data_manual$chromosomeLength <- whole.cell.data_manual$SC.length

whole.cell.data_manual$PER_Foci_1 <- whole.cell.data_manual$Foci1 / whole.cell.data_manual$chromosomeLength
whole.cell.data_manual$PER_Foci_2 <- whole.cell.data_manual$Foci2 / whole.cell.data_manual$chromosomeLength
whole.cell.data_manual$PER_Foci_3 <- whole.cell.data_manual$Foci3 / whole.cell.data_manual$chromosomeLength
#whole.cell.data_manual$PER_Foci_4 <- whole.cell.data_manual$Foci4 / whole.cell.data_manual$chromosomeLength

#mouse, strain, sex, category
whole.cell.data_manual <- add_mouse(whole.cell.data_manual)
whole.cell.data_manual <- add_category(whole.cell.data_manual)
whole.cell.data_manual <- add_strain(whole.cell.data_manual)
whole.cell.data_manual <- add_sex(whole.cell.data_manual)
whole.cell.data_manual <- add_subsp(whole.cell.data_manual)

#whole.cell.data_manual <-   whole.cell.data_manual[(!is.na(whole.cell.data_manual$fileName)),]
whole.cell.data_manual <- whole.cell.data_manual[!(is.na(whole.cell.data_manual$fileName) | whole.cell.data_manual$fileName==""), ]

#obj.id
whole.cell.data_manual$Obj.ID <- paste(whole.cell.data_manual$fileName, whole.cell.data_manual$boxNumber, sep = "_")

#add all the extra columns, blah
#percent Foci position

#ADD RANK
for(cellys in as.character(unique(whole.cell.data_manual$fileName))){

 # print(cellys)
  chrm.order = rank(whole.cell.data_manual$SC.length[whole.cell.data_manual$fileName == cellys])
  whole.cell.data_manual$rank[whole.cell.data_manual$fileName == cellys] <- chrm.order  
  }

#IFD 
whole.cell.data_manual <- add_IFD(whole.cell.data_manual)

#PER Foci

```

```{r clean.whole.cell, echo = FALSE}

#CHECKED FOR MISSING bivs -- also check for wrong directions, ect

#negative IFD
#5aug17_14jul17_LEW_f1_sp1_11_rev_blob4

#PER foci positions greater th

#check the numbers of bivalents with a SC length, but no hand.foci.count

```

# Cell Level table

- pass rate (don't need the OG.DF file, because I can count the number of boxes with the max index number)

- number bivs per mice?

- number cells represented per mouse

- number of pass.SC w/out foci position data

- number of foci with 'faint' mentioned in the notes

The below cell level table is used for making the **long.biv.dataset**, (any others?)

```{r cell.level.table, echo=FALSE}

cell.level.BivData_table <- ddply(.data=Curated_BivData, 
                       .(fileName),
                       summarize, 
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                       ncells = length(unique(fileName)),
                       nbivs = length(unique(Obj.ID)),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number

                       cv.SC = cv(chromosomeLength),

#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),

                       mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                      var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)

#cent_dis
                      mean_cent_dist = mean(dis.cent, na.rm=TRUE)
#telomere dist

#can't think of a why to calq this by just 2CO

)
                       #add the centromere and ifd here?
                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),

#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
 #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),

#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
#                       var.IFD = var(IFD1, na.rm=TRUE),
#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),

 #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)

cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
cell.level.BivData_table <- add_strain(cell.level.BivData_table)
cell.level.BivData_table <- add_category(cell.level.BivData_table)

as.data.frame( head(cell.level.BivData_table) )

```

```{r whole.cell_merging, echo=FALSE, eval=FALSE}
#I think this block is redundant, could delete

#check if there are fileNames
double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% Curated_BivData$fileName, ]

#which cell-levels fileNames have close to 15-17 chrms?
almost_whole_cell <- cell.level.BivData_table[(cell.level.BivData_table$Nbiv < 21),] #cut off cells with too many
almost_whole_cell <- almost_whole_cell[almost_whole_cell$Nbiv > 15,]

#there is some discrpancy with the Nbiv and nbivs columns! ugh why, some have 20 and 10

#get the Biv data 
almost_whole_cell_BivData <- Curated_BivData[ Curated_BivData$fileName %in% almost_whole_cell$fileName, ]

as.data.frame(  table( droplevels( almost_whole_cell_BivData$fileName) ) )
#length(unique(almost_whole_cell_BivData$fileName))

#what is almost whole cell BivData used for...
```

3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.  
8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev

## Long Biv Data Set

I want to make a dataset of 'longest' 25% chromosomes from cells with at least 16 segmented bivalents from a single cell.

```{r long.chrm.dataset, echo=FALSE}

#1. how many biv's are from cells with >15 bivalents
over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

#check for blank obj.id
BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")

#check for dup
#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]

#remove duplicate obj.ids
BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]

#make sure obj.id and fileName are charcter for str stuff
BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)

#make sure 
missing_from_cell_table <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% Biv_Data_mean_table$fileName, ]#there are 75, objs 3 cells
#these are bivalents which don't appear the in cell level table, which can break the loop
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
#final cleaned BivData table, 2428

#loop for greping the mouse from file name ect
for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
  #the above can produe mulitple Q4

  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
  #these are just using the last index,

  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]

  BivData_from_15biv_cells$Q4.FN[i] <- the.FN
   # print(c(i, the.FN, the.Q4))

}

BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= BivData_from_15biv_cells$cell.Q4), "yes", "no" )

real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]

```

```{r table.long.biv, echo=FALSE}

#real.long.bivData
#calq the mouse averages for

Long_biv_mouse.avs_4MM <- ddply(real.long.bivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),
                       #IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

)

Long_biv_mouse.avs_4MM <- add_sex(Long_biv_mouse.avs_4MM)
Long_biv_mouse.avs_4MM <- add_strain(Long_biv_mouse.avs_4MM)
Long_biv_mouse.avs_4MM <- add_subsp(Long_biv_mouse.avs_4MM)

Long_biv_mouse.avs_4MM$strain <- unclass(Long_biv_mouse.avs_4MM$strain)
Long_biv_mouse.avs_4MM$strain <- as.factor(Long_biv_mouse.avs_4MM$strain)

Long_biv_mouse.avs_4MM$sex <- unclass(Long_biv_mouse.avs_4MM$sex)
Long_biv_mouse.avs_4MM$sex <- as.factor(Long_biv_mouse.avs_4MM$sex)

Long_biv_mouse.avs_4MM$subsp <- unclass(Long_biv_mouse.avs_4MM$v)
Long_biv_mouse.avs_4MM$subsp <- as.factor(Long_biv_mouse.avs_4MM$subsp)

```

# Mouse leve table

the mouse table will be the pooled stats / summary of all SC.passing +

```{r mouse.table, echo=FALSE}

mouse.level.BivData_table <- ddply(.data=Curated_BivData, 
                       .(mouse),
                       summarize,

                       ncells = length(unique(fileName)),

                       boxs.IDd = max(boxNumber),  #
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                       ncells = length(unique(fileName)),
                       nbivs = length(unique(Obj.ID)),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3),

                       #sis.co.ten
                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
)

as.data.frame( head(mouse.level.BivData_table) )
#kable didn't work

```

```{r table.HetC.MM, echo=FALSE}

```

```{r Q1Q2.mouse.table, echo=FALSE}

```

analysis using the mouse level table -- to justify that using mouse averages is a good idea.

Plots

1. Histograms of N.bivs per cell, facet category

2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)

3.

Is there a way to check my assumptions for these measues?

What are the sources of error fot these data?

```{r telo.cent_dist, echo=FALSE}

#code for cal1 the telomere distance
#space between last foci and telomere end

#telo_dist <- if(hand.foci = 1){
          #SC.Length - F1
#}
#elseif(hand_foci == 2){ #SC.length - F2}

#code for centromere
#space between first foci and centromere/centromere-end

#cent_end_dist
#cent_dist
```

# melt position table

make the melt DF for the abs (raw) foci positions and then calculated the normalized (PER) foci positions from that table

```{r MELT.block1, echo=FALSE}

col.names.list <- colnames(Curated_BivData)
#"fileName" ,  "chromosomeLength",        "centromere_ABS_Position", "centromere_PER_Position",
#"boxNumber",               "Obj.ID",                  "SC.pass",                 "foci.pass",              
#"curated.1pass.0fail",     "curation.notes",          "hand.foci.count",        "Foci1",                  
#"Foci2",       "Foci3",                "notes",                 "mouse",                  
#"category",                "strain",                  "sex",                    "subsp",                  
#"IFD1_ABS",                "IFD1_PER",

#all of the cols to use
MELT_Curated_BivData <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
              #leave out the Foci
                "Foci1","Foci2","Foci3", "Foci4")],

            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]

#MELT, -- 15
colnames(MELT_Curated_BivData) <- c(col.names.list[c(1:11, 16:17)], "Foci_Type", "Pos") #too many foci

MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)

#calq normalized melt pos
MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]

```

```{r general.foci.pos, echo=FALSE, eval=FALSE}

#make the normalized foci density plots
#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]

#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]#

#remove 0 and NA
#remove 4foci count?

#check for values over 1

#density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")

#density
#table(the4$mouse)
#something wrong with the 4 bivalents
#many of the lines just have 3 lines

```

This is proof that the plot codes work, but move the figures in the analysis doc. remeber this is just data-setup.

think about weighing the density plots by the number of bivalents per category...

# Assemble final BivData DF

Main files from this plot:

  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures

  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)

  3. MELT_ABS: make melt version of foci positions for abs measuers

  4. CELL and mouse table

  5.

MAIN BIVDATA_DF
decide on the colnames that I want for the final DF

fileName, Object_ID, box.number, centromere_ABS, centromere_PER
chromosome length, SC.pass, Foci.pass, hand.foci.count, 
F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD

telomere_dist, cent_dist,

mouse, strain, category, subsp,

merge:
cell.MLH1.count, cell.quality, mouse.age, 
cell.total_skel, cell.total_binary

# Save

```{r save.file}

#save Rdata
save.image("~./MLH1repo/data/CleanBivData_1.28.20.RData")
#save.image("data/BivData_4_PCA.RData")

```

# Analysis

planned analyses for later docs, but this is the list for getting everything ready

Use the Bivalent_Analysis.Rmd for analysis pieces
BivPattern.Rmd started to analyze this but, not a very clear setup

-chrom class proportions

-SC lengths, by CO number

<!--
trash / junk
#Apply the cleaning steps below to the FULL-OG DF

#remove non number charaecters for the FULL / OG DF
#Curated_BivData <- Curated_BivData[ !grepl("[?]", Curated_BivData$SC.pass) , ]
#Curated_BivData <- Curated_BivData[ !grepl("[?]", Curated_BivData$hand.foci.count) , ]

#remove data which has paint marks
#P_fileNames <- Curated_BivData$fileName[(grep('p_rev', Curated_BivData$fileName))]
#length(P_fileNames)

#remove anything with 'DUPLICATE'
#DUP_list <- Curated_BivData$fileName[(grep('DUP', Curated_BivData$fileName))]
#DUP_num = length(DUP_list)#0

#dup_list= c()
#for(i in Curated_BivData$fileName){  #change the file name header,  
#  # print(i)
#  ifelse(grepl("\\.[0-9]_rev", i),
#         (dup_list = c(dup_list, i)), NA)
#}
#MultiCells = c()
#for(j in dup_list){
#  stringLIST = strsplit(j, split="\\.")[[1]]
#  new_string = paste(stringLIST[[1]][1], "rev", sep="_")
#  MultiCells = c(MultiCells, new_string)
#}
#MultiCells = unique(MultiCells)
#multi_cell_num = length(MultiCells)#118

#new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
#colnames(new.Curate.FULL)

```{r clean.data, echo=FALSE, eval=FALSE}
#Keep these biv data for the male polymorphism version

#Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
#Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
#Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]

#Curated_BivData$strain <- droplevels(Curated_BivData$strain)
```

-->