### 2020-02-04 doc/BivPattern.Rmd
summarizing all results, updating saving files

---
title: "BivData Patterns"
author: "April Peterson"
date: "October 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup.data, echo=FALSE}

library(ggplot2)

setwd("~./MLH1repo/")
#load curated data.frame
load(file="data/BivData_4_PCA.RData")
```

first discribe data (table functions)

Test everything from chrms of the same class across strains
How similar are 1COs bivalents in MSM,PWD, and KAZ, LEW ect?
(male and female?)

1CO bivalents;

> position
> centromere position
> SC length

```{r blah}

master_curated_BivData <- curated.oneCO.male[!(is.na(curated.oneCO.male$fileName) | curated.oneCO.male$fileName ==""), ]
master_curated_BivData <- master_curated_BivData[master_curated_BivData$hand.foci.count != 8,]

all1CO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 1,]
all1CO <- all1CO[!(is.na(all1CO$fileName) | all1CO$fileName ==""), ]

blahblah <- ggplot(all1CO, aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))

all2CO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 2,]
all2CO <- all2CO[!(is.na(all2CO$fileName) | all2CO$fileName ==""), ]

blahblah2 <- ggplot(all2CO, aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("2COs male and females")+ ylim(c(30,250))

## (try IFD)
staggered <- ggplot(master_curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("")+theme(legend.position="none") #+ ylim(c(30,250))

fill.by.ifd <- ggplot(master_curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=IFD1_PER))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("SC length Fill by IFD")#+theme(legend.position="none") #+ ylim(c(30,250))

ifd.rand <- ggplot(master_curated_BivData, aes(y = IFD1_PER, x=hand.foci.count, color=strain))+geom_jitter(aes(alpha=0.2))+ facet_wrap(~sex)+ ggtitle("SC length Fill by IFD")#+theme(legend.position="none") #+ ylim(c(30,250))

IFD.comp <- ggplot(master_curated_BivData, aes(y = IFD1_PER, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("IFD1 distributions")+theme(legend.position="none")

```

```{r contrast.chrm.class, echo=FALSE}

curated.oneCO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 1,]
curated.oneCO.male <- curated.oneCO[curated.oneCO$sex == "male",]

curated.oneCO.male$Foci1.PER <- curated.oneCO.male$Foci1 / curated.oneCO.male$chromosomeLength

curated.oneCO.male <- curated.oneCO.male[!(is.na(curated.oneCO.male$fileName) | curated.oneCO.male$fileName ==""), ]

#lm - but what trait do i want to predict
#one.co.lm <- lm(data = curated.oneCO.male)

#ggplots
oocc <- ggplot(curated.oneCO.male, aes(y = dis.cent, x = Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)+ylim(c(0,120))

mabeh <- ggplot(curated.oneCO.male, aes(x= dis.cent, y =chromosomeLength ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)#+ ylim(c(0,125))+xlim(c(30,150))

ghost <- ggplot(curated.oneCO.male, aes(y = chromosomeLength, x =Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)+ylim(c(30,150))

ghost.ssdgm <- ggplot(curated.oneCO.male, aes(y = chromosomeLength, x = Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ylim(c(30,150))

Length.1CO <- ggplot(curated.oneCO.male, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter()+
  ggtitle("1CO chromosome Length") #+facet_wrap(~strain) #+ylim(c(0,150))
#check the outlier

oocc
mabeh
ghost
```

KAZ has more terminal 1CO. 
are 1COs longer?

MSM has higher mean, the PWD and MSM have stronger floor. LEW has stronger ceiling?

MSM and PWD have long chromosome length. (and stronger floor - what is the biological significance of the floor?)

predict that rank would have strata. (there is true bivalent rank and realized rank)

```{r two.co.biv, echo=FALSE}

curated.twoCO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 2,]
curated.twoCO.male <- curated.twoCO[curated.twoCO$sex == "male",]

curated.twoCO.male$Foci1.PER <- curated.twoCO.male$Foci1 / curated.twoCO.male$chromosomeLength
curated.twoCO.male$Foci2.PER <- curated.twoCO.male$Foci2 / curated.twoCO.male$chromosomeLength

curated.twoCO.male <- curated.twoCO.male[!(is.na(curated.twoCO.male$fileName) | curated.twoCO.male$fileName ==""), ]

#ggplots
ttoo.oo <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x = IFD1 ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ylim(c(0,150))

per.ifd.oo <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter()+
  ggtitle("2CO chromosome Length") #+facet_wrap(~strain) #+ylim(c(0,150))

mabeh.two <- ggplot(curated.twoCO.male, aes(x = dis.cent, y =chromosomeLength ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ ylim(c(0,125))+xlim(c(30,120))

ghost.two <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x = Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)+ylim(c(30,150))

ghost.two2 <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x = Foci2.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ylim(c(30,150))

```

there seems to be less variance in the MSM and PWD IFD percent. the lengths are similar, but the PER.IFDs are tighter.
Also, in PWD and MSM there are stronger 'floors' for the 2CO bivalents

```{r female.patterns, echo=FALSE}

curated.female.bivs <- master_curated_BivData[master_curated_BivData$sex == "female",]
curated.female.1CO <- curated.female.bivs[curated.female.bivs$hand.foci.count == 1,]
curated.female.2CO <- curated.female.bivs[curated.female.bivs$hand.foci.count == 2,]

curated.female.1CO <- curated.female.1CO[!(is.na(curated.female.1CO$fileName) | curated.female.1CO$fileName ==""), ]
curated.female.2CO <- curated.female.2CO[!(is.na(curated.female.2CO$fileName) | curated.female.2CO$fileName ==""), ]

curated.female.1CO$dist.cent.per <- curated.female.1CO$dis.cent / curated.female.1CO$chromosomeLength
curated.female.2CO$dist.cent.per <- curated.female.2CO$dis.cent / curated.female.2CO$chromosomeLength

Length.1CO.female <- ggplot(curated.female.1CO, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(color=dis.cent))+
  ggtitle("1CO chromosome Length Female")

Length.2CO.female <- ggplot(curated.female.2CO, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(color=dist.cent.per))+
  ggtitle("2CO chromosome Length Female")

```

weaker basements? (how can I test) violin plots?
KAZ and LEW have short 1CO bivalents.

Lew shorter 2CO chrms, 
weaker basments in 2CO chrms

within category variation

```{r kaz.male, echo=FALSE}
#master_curated_BivData
#KAZ males
kaz.male <- master_curated_BivData[master_curated_BivData$category == "KAZ male",]
#3 mice, 73 cells

kaz.male.1CO <- kaz.male[kaz.male$hand.foci.count == 1,]

first.model <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=kaz.male.1CO)
summary(first.model)
#1CO foci position is driven by SC length

kaz.male.2CO <- kaz.male[kaz.male$hand.foci.count == 2,]
kaz.male.2CO <- kaz.male.2CO[!(is.na(kaz.male.2CO$fileName) | kaz.male.2CO$fileName==""), ]

two.model <- lm(Foci2 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+Foci1,  data=kaz.male.2CO)
summary(two.model)
#Foci1 not predictable, but foci2 position (raw), predicted by chrm length and mouse

SC.length.lm <- lm(chromosomeLength ~ mouse +centromere_ABS_Position+centromere_PER_Position+avergeRED+Foci1,  data=kaz.male.1CO)
summary(SC.length.lm)

SC.length.lm2 <- lm(chromosomeLength ~  IFD1_ABS+centromere_ABS_Position+Foci1+Foci2+dis.cent+IFD1,  data=kaz.male.2CO)
summary(SC.length.lm2)
#the significance depends on the order, 
#centromere position is significant for predicting SC length

length(unique(kaz.male$fileName))

```

```{r KAZ.female}

kaz.female <- master_curated_BivData[master_curated_BivData$category == "KAZ female",]
kaz.female.1CO <- kaz.female[kaz.female$hand.foci.count == 1,]
kaz.female.2CO <- kaz.female[kaz.female$hand.foci.count == 2,]

kaz.female.1CO <- kaz.female.1CO[!(is.na(kaz.female.1CO$fileName) | kaz.female.1CO$fileName==""), ]
kaz.female.2CO <- kaz.female.2CO[!(is.na(kaz.female.2CO$fileName) | kaz.female.2CO$fileName==""), ]

first.model <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=kaz.female.1CO)
summary(first.model)
#chrm length is strongest predictor

two.model <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+IFD1+Foci2,  data=kaz.female.2CO)
summary(two.model)
#chrm length and foci1, predict foci2, and IFD1
#to predict foci1, IFD, chrm length, and foci 2 are strong predictors

length.model <- lm(chromosomeLength ~ mouse +Foci1 +centromere_ABS_Position+centromere_PER_Position+avergeRED+hand.foci.count,  data=kaz.female)
summary(length.model)
#centromere position is sig, IFD1, FOci1 is sig -- this is true for 1 and 2CO
#for all chrms, hand.foci.count is significant

```

```{r MSM.male, echo=FALSE}
#MSM male
MSM.male <- master_curated_BivData[master_curated_BivData$category == "MSM male",]
droplevels(MSM.male)
MSM.male.1CO <- MSM.male[MSM.male$hand.foci.count == 1,]
MSM.male.2CO <- MSM.male[MSM.male$hand.foci.count == 2,]

MSM.male.1CO <- MSM.male.1CO[!(is.na(MSM.male.1CO$fileName) | MSM.male.1CO$fileName==""), ]
MSM.male.2CO <- MSM.male.2CO[!(is.na(MSM.male.2CO$fileName) | MSM.male.2CO$fileName==""), ]

first.mo <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=MSM.male.1CO)
summary(first.mo)
#chrm length is most sig

two.mo <- lm(Foci2 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+IFD1+Foci1+SisCoTen,  data=MSM.male.2CO)
summary(two.mo)
#one of the males is sig, chrm length and IFD1, foci2 are signicant

length.mo <- lm(chromosomeLength ~ centromere_ABS_Position+centromere_PER_Position+avergeRED+hand.foci.count+mouse,  data=MSM.male)
summary(length.mo)

#foci1, and centromere positions are sig for 2CO
#centromeres sig
#for all bivs, centromere and hand.foci are the sig factors

```

If all bivalents included for a strain, and sex included, sex will be significant for chrm (for foci as well?) -- what does that add -- 
models don't predict foci, centromere doesn't predict.
I think the chromosome length predicted by centromere position might be related to resolution.

```{r MSM.female}

MSM.female <- master_curated_BivData[master_curated_BivData$category == "MSM female",]
droplevels(MSM.female)
MSM.female.1CO <- MSM.female[MSM.female$hand.foci.count == 1,]
MSM.female.2CO <- MSM.female[MSM.female$hand.foci.count == 2,]

MSM.female.1CO <- MSM.female.1CO[!(is.na(MSM.female.1CO$fileName) | MSM.female.1CO$fileName==""), ]
MSM.female.2CO <- MSM.female.2CO[!(is.na(MSM.female.2CO$fileName) | MSM.female.2CO$fileName==""), ]

first.mo <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=MSM.female.1CO)
summary(first.mo)
#chrm length is most sig

two.mo <- lm(Foci2 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+IFD1+Foci1+SisCoTen,  data=MSM.female.2CO)
summary(two.mo)
#one of the males is sig, chrm length and IFD1, foci2 are signicant

length.mo <- lm(chromosomeLength ~ centromere_ABS_Position+centromere_PER_Position+avergeRED+hand.foci.count+mouse,  data=MSM.female)
summary(length.mo)
#for chrm length in 2COs, intercept, and centromere postions are sig (hand foci count is not significant if i only use 1 chrm class), it is sig for all bivs
#1 of the mice is significant

```

what does a significant intercept for foci1 mean? (it's different than 0?)

```{r melt.foci.positions, echo=FALSE}

```