### 2020-02-19 doc/Debug_Rmd.Rmd
messy work from home

---
title: "MLH1 Results Outline"
output:
  html_notebook: default
  html_document: default
---

```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)

options(digits=2)

setwd("~.")

load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

#2.2.20
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")

load(file="~./MLH1repo/data/CleanBivData_13.2.20.RData")

load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")

#setwd("~./MLH1repo/doc/")
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
#setwd("C:/Users/April/Documents/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")

#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

#load(file="data/MetaData.RData") #this was replacing most updated data file

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

# Data set ups

```{r organizing.data, echo=FALSE }
#set the order of another column, based on another variable. consistant order for ploting

#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
#-ene endings refer to the cells/units within the cell stage? (I think)

#apply this before ploting
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, category, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
#use the above line for when order is out of wak

#count the non-quality measures,  #remove non qualit
#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]

#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]

#MLH1_data <- MLH1_data %>%
#  arrange(category) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
                 .(strain,sex),
                 summarize,
                 sum.ncells = sum(Ncells),
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co),
                 AV.var_MLH1 = mean(var, na.rm=TRUE)

)

#there might be an error for changing wd
#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)

AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]

var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()

```

```{r misc.setup, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4',

                       'TOM'= 'indianred1','AST' = 'tan2',
                       'CZECH'= 'brown1',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
#4!!

#show_col(hue_pal()(5))

#add Rec group,
#reclassified SKIVE as intermediate
Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))

Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

#apply this before ploting, to display strains in desired order
#MLH1_data <- MLH1_data %>%
#  arrange(strain, sex, category, mouse) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

# sort your dataframe, by the focal categories
#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

```

```{r MLH1.FEMALE.TABLE, echo=FALSE}

female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

female_mouse.av <- add_sex(female_mouse.av)
female_mouse.av <- add_strain(female_mouse.av)
female_mouse.av <- add_subsp(female_mouse.av)
female_mouse.av <- add_species(female_mouse.av)

female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")

female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)

Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

#Mol
#Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")

#Dom 
Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")

```

```{r MLH1.MALE.TABLE, echo=FALSE}
male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)

fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)

Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
#Dom version 
Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
```

```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ

#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

#What are the main DFs for this code chunk
#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
#DF.HetC.MixedModel   <- only Mmusculus
#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded

#variance is still chr
AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

#remove the extra merged cols
DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )
#DONT NEED CATEGORY

#REMOVE EXTRA STRAINS
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#main df with variables coded as chr
DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))

all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
                 .(sex, strain),
                 summarize,
                 all.cells = sum(Ncells)
)

```

```{r CurBiv.Mouse.av.TABLE, echo=FALSE}

#this block should be in the data set up
#the mouse averages will be used

mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
mouse.avs_4MM <- add_category(mouse.avs_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)

#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
#code if males are high or low?

#code intp chr

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))

```

```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}

#add this to setup

#remove female, 
#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons

#Curated_BivData_male

male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#vari
male.av_4MM <- add_sex(male.av_4MM)
male.av_4MM <- add_strain(male.av_4MM)
male.av_4MM <- add_subsp(male.av_4MM)
male.av_4MM <- add_category(male.av_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
```

```{r totSC_setup, echo=FALSE}
#SC.skel
#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included

SC.skel.OG <- SC.skel #5683

#MERGEING MLH1 with SKEL
MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)

MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
#5110,

OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF

#check for DUPs, #check for duplicatations of the random name
MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]

MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")

MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
#table(MLH1.merge.Skel$category)

#remove outliers
MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)

```

```{r totSC.viz4outliers, echo=FALSE}

# -- EXAMINE the ggplot scatter plots for outliers
tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")

#round 1
SC.skel.outliers_round1 <- c(
  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",

  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",

       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",

   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
   "6may15_17apr15_PWD_f1_sp1_7_rev",

      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",

    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
 "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
 "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
 "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,

  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",

# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",

#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
"16aug19_27aug18_AST_m1_sp1_23_rev.tif",

#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",

#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
 "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
 "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
 "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
 "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
 "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
 "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
 "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )

MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed

remove.mice_round2 <- c(
    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",

  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",

# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",

#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
#SPIC male  x
#SPIC female  x
#TOM x
#AST x
#CZECH male x

#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
 "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
#SPRET female X
#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
)
#3remove.mice_round2

MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]

all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
```

<!-- 
two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
-->

```{r tot.skel.table, echo=FALSE}

totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                #se for skel

                 mean.bin =  mean(bin_size, na.rm = TRUE)

)

totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
totSC.mouse.av.table <- add_species(totSC.mouse.av.table)

#unorder the factors
totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)

#add rec group
totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))

OG.totSC.mouse.av.table <- totSC.mouse.av.table

totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av. total SC area")

mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")

```

```{r, tot.SC.setup, echo = FALSE}

totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
                 .(sex,strain),
                 summarize, 
                 ncells = sum(ncells),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(mean.skel, na.rm = TRUE),
                 mean.bin =  mean(mean.bin, na.rm = TRUE)

)

totSC.strain.av.table <- add_subsp(totSC.strain.av.table)

```

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1

INSERT TABLE FOR THE TOTAL SC means ect

The difference between

bin is short for binary, it reports the binarized area of the SC

skel -- takes the skeletonized midline of the bin objects
-->
# Main Figure

```{r table.for.plot, echo=FALSE, warning=FALSE}
#why do I add in a different table?

#this move this to the setup file
#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed

#pooled
strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)

                       #range_mouse_means = range()
                   #consider adding crhm proportions?
    )

strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")

#adjust order of plot here
strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ",
  "WSB","G","LEW","PERC",
                     "CAST",
                         "SPRET","SPIC" ), order=T )

```

```{r Fig1, echo=FALSE, warning=FALSE}

#might need to nudge the X values

Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#remove y labfor male

   #segments
#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
 # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label

 #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
   #        y = c(18.5,18.5,33,18.5, 18.5),

  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
    #       fontface = 'italic', hjust = 0.5, size=3.5)

Fig1_v1_strain
```

This is a draft of the main figure. Outgroups are not included in the analysis below.

<!--  insert Details of data set cleaning etc is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd -->

```{r main.fig1s, echo=FALSE, include=FALSE}
strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )

Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point( size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

#strain.table.HQ_Fig.og

strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")

strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
  "WSB","G","LEW"), order=T )

Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

```

```{r basic.numbers, echo=FALSE, eval=FALSE}

kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )

kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )

```

```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
#str(AP_mouse_table_w.Ages)

AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

#remove sex.x == na, and strain.x=na or F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
#plot variance and mean (facet sex)

AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]

VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")

cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")

str(AP_mouse_table.HQ)

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
#[1] 0.5750399
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
#0.321601

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
# 0.3515544
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
#-0.03713536

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
# p-value = 8.623e-09

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )

#remove mice with fewer than 10 cells

AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]

VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")

cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])

cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )

#why did I male just a female table?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]

female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)
```

# Genome wide recombiantion rate estimates

```{r MLH1.summ.stats, echo=FALSE}

#strain.table.HQ_Fig.og  all strains
#strain.table.HQ_Fig  extra strains removed

main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")

main.table <- add_subsp(main.table)
main.table <- add_species(main.table)

kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")

#strain.table.HQ_Fig

#sum( (main.table  %>% filter(sex == "male"))$Ncells)
#the kable version is bulkier than the other versaion
```

In order to estimate the genome wide recombination rates across our panel of three subspecies from the house mouse species complex, _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. 
`r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains were used with reflect samples of natural genetic variation across the terretories of these subspecies.

MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes. The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.

1. -female strain means are v close  -male strain means are more varied!!
-Comparisons to references / literatures
-house mouse close to minimum of 1  obligate CO per chrm

2. The largest female difference is G, 1.07X  the largest male difference is PWD and MSM which are 1.3 and 1.2 respectively.

<!-- female -->
In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.

```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

female_mouse.av <- add_sex(female_mouse.av)
female_mouse.av <- add_strain(female_mouse.av)
female_mouse.av <- add_subsp(female_mouse.av)
female_mouse.av <- add_species(female_mouse.av)

female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")

female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)

#table( droplevels(female_mouse.avs_for.analysis$strain) )
#there are cast strains too

Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Female mouse av. MLH1 count per cell ")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))

#female.specific.plot

musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))

#Mol
Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")

Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))

#Dom version 
Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")

Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))

```

## female section

```{r female.strain.table, echo=FALSE}

#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
F.strain.table

# I don't know why most of the table is blank
```

```{r female.dif, echo=FALSE}

#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co)

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)

```

##showing values

'tostring()  didn't work

The overall pattern for the female mean MLH1 counts per cell is low variation. ( in the most different strains
In females specifically _domesticus^G^_ is `r toString(G.female.diff)` higher the the other means. 
while _ domesticu^LEW^_  is `r LEW.female.diff` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.

In the _domesticus_ subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to