### 2020-02-20 doc/Results.rmd
more slow progress

---
title: "MLH1 Results Outline"
date: 
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  html_notebook: default
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)

options(digits=2)

setwd("~.")

load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

#2.2.20
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")

load(file="~./MLH1repo/data/CleanBivData_19.2.20.RData")

load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")

#setwd("~./MLH1repo/doc/")
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
#setwd("C:/Users/April/Documents/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")

#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

#load(file="data/MetaData.RData") #this was replacing most updated data file

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

# Data Set Up Notes

```{r organizing.data, echo=FALSE }
#set the order of another column, based on another variable. consistant order for ploting

#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
#-ene endings refer to the cells/units within the cell stage? (I think)

#apply this before ploting
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, category, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
#use the above line for when order is out of wak

#count the non-quality measures,  #remove non qualit
#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]

#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]

#MLH1_data <- MLH1_data %>%
#  arrange(category) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
                 .(strain,sex),
                 summarize,
                 sum.ncells = sum(Ncells),
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co),
                 AV.var_MLH1 = mean(var, na.rm=TRUE)

)

#there might be an error for changing wd
#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)

AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]

var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()

```

```{r misc.setup, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4',

                       'TOM'= 'indianred1','AST' = 'tan2',
                       'CZECH'= 'brown1',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
#4!!

#show_col(hue_pal()(5))

#add Rec group,
#reclassified SKIVE as intermediate
Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))

Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

#apply this before ploting, to display strains in desired order
#MLH1_data <- MLH1_data %>%
#  arrange(strain, sex, category, mouse) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

# sort your dataframe, by the focal categories
#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

```

```{r MLH1.FEMALE.TABLE, echo=FALSE}

female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

female_mouse.av <- add_sex(female_mouse.av)
female_mouse.av <- add_strain(female_mouse.av)
female_mouse.av <- add_subsp(female_mouse.av)
female_mouse.av <- add_species(female_mouse.av)

female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")

female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)

Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

#Mol
#Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")

#Dom 
Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")

```

```{r MLH1.MALE.TABLE, echo=FALSE}
male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)

fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)

Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
#Dom version 
Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
```

```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ

#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

#What are the main DFs for this code chunk
#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
#DF.HetC.MixedModel   <- only Mmusculus
#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded

#variance is still chr
AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

#remove the extra merged cols
DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )
#DONT NEED CATEGORY

#REMOVE EXTRA STRAINS
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#main df with variables coded as chr
DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))

all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
                 .(sex, strain),
                 summarize,
                 all.cells = sum(Ncells)
)

```

```{r juvimice.list, echo=FALSE}
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )

#REMOVE JUVI mice
#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]

```

```{r CurBiv.Mouse.av.TABLE, echo=FALSE}

#this block should be in the data set up
#the mouse averages will be used

#has 
#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
Curated_BivData <- Curated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")

mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
mouse.avs_4MM <- add_category(mouse.avs_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)

#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
#code if males are high or low?

#REMOVE JUVI mice
juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))

mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))

```

```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}

#add this to setup

#remove female, 
#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons

#Curated_BivData_male

male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#vari
male.av_4MM <- add_sex(male.av_4MM)
male.av_4MM <- add_strain(male.av_4MM)
male.av_4MM <- add_subsp(male.av_4MM)
male.av_4MM <- add_category(male.av_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)

#REMOVE JUVI mice
juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]

```

```{r totSC_setup, echo=FALSE}
#SC.skel
#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included

SC.skel.OG <- SC.skel #5683

#MERGEING MLH1 with SKEL
MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)

MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
#5110,

OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF

#check for DUPs, #check for duplicatations of the random name
MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]

MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")

MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
#table(MLH1.merge.Skel$category)

#remove outliers
MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)

```

```{r totSC.viz4outliers, echo=FALSE}

# -- EXAMINE the ggplot scatter plots for outliers
tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")

#round 1
SC.skel.outliers_round1 <- c(
  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",

  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",

       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",

   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
   "6may15_17apr15_PWD_f1_sp1_7_rev",

      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",

    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
 "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
 "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
 "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,

  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",

# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",

#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
"16aug19_27aug18_AST_m1_sp1_23_rev.tif",

#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",

#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
 "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
 "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
 "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
 "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
 "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
 "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
 "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )

MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed

remove.mice_round2 <- c(
    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",

  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",

# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",

#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
#SPIC male  x
#SPIC female  x
#TOM x
#AST x
#CZECH male x

#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
 "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
#SPRET female X
#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
)
#3remove.mice_round2

MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]

all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
```

<!-- 
two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
-->

```{r tot.skel.table, echo=FALSE}

totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                #se for skel

                 mean.bin =  mean(bin_size, na.rm = TRUE)

)

totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
totSC.mouse.av.table <- add_species(totSC.mouse.av.table)

#unorder the factors
totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)

#add rec group
totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))

OG.totSC.mouse.av.table <- totSC.mouse.av.table

#REMOVE JUVI mice
tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]

totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]

totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av. total SC area")

mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")

```

```{r, tot.SC.setup, echo = FALSE}

totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
                 .(sex,strain),
                 summarize, 
                 ncells = sum(ncells),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(mean.skel, na.rm = TRUE),
                 mean.bin =  mean(mean.bin, na.rm = TRUE)

)

totSC.strain.av.table <- add_subsp(totSC.strain.av.table)

```

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1

INSERT TABLE FOR THE TOTAL SC means ect

The difference between

bin is short for binary, it reports the binarized area of the SC

skel -- takes the skeletonized midline of the bin objects
-->
# Main Figure

```{r table.for.plot, echo=FALSE, warning=FALSE}
#why do I add in a different table?

#this move this to the setup file
#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed

#pooled
strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)

                       #range_mouse_means = range()
                   #consider adding crhm proportions?
    )

strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")

#adjust order of plot here
strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ",
  "WSB","G","LEW","PERC",
                     "CAST",
                         "SPRET","SPIC" ), order=T )

```

```{r Fig1, echo=FALSE, warning=FALSE}

#might need to nudge the X values

Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#remove y labfor male

   #segments
#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
 # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label

 #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
   #        y = c(18.5,18.5,33,18.5, 18.5),

  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
    #       fontface = 'italic', hjust = 0.5, size=3.5)

Fig1_v1_strain
```

This is a draft of the main figure. Outgroups are not included in the analysis below.

<!--  insert Details of data set cleaning etc is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd -->

```{r main.fig1s, echo=FALSE, include=FALSE}
strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )

Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point( size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

#strain.table.HQ_Fig.og

strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")

strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
  "WSB","G","LEW"), order=T )

Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

```

```{r basic.numbers, echo=FALSE, eval=FALSE}

kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )

kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )

```

```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
#str(AP_mouse_table_w.Ages)

AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

#remove sex.x == na, and strain.x=na or F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
#plot variance and mean (facet sex)

AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]

VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")

cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")

str(AP_mouse_table.HQ)

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
#[1] 0.5750399
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
#0.321601

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
# 0.3515544
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
#-0.03713536

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
# p-value = 8.623e-09

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )

#remove mice with fewer than 10 cells

AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]

VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")

cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])

cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )

#why did I male just a female table?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]

female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)
```

# Genome wide recombiantion rate estimates

```{r MLH1.summ.stats, echo=FALSE}

#strain.table.HQ_Fig.og  all strains
#strain.table.HQ_Fig  extra strains removed

main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")

main.table <- add_subsp(main.table)
main.table <- add_species(main.table)

kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")

#strain.table.HQ_Fig

#sum( (main.table  %>% filter(sex == "male"))$Ncells)
#the kable version is bulkier than the other versaion
```

We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from thress subspecies within the house mouse species complex;  _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes. The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.

<!--
more basic stats, error, comparison to other measures

transition first aragraph to strain specific
Main points
1. -female strain means are v close  -male strain means are more varied!!
-Comparisons to references / literatures
-house mouse close to minimum of 1  obligate CO per chrm

2. The largest female difference is G, 1.07X  the largest male difference is PWD and MSM which are 1.3 and 1.2 respectively.  
-->

<!-- transition from first part into female -->

```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

female_mouse.av <- add_sex(female_mouse.av)
female_mouse.av <- add_strain(female_mouse.av)
female_mouse.av <- add_subsp(female_mouse.av)
female_mouse.av <- add_species(female_mouse.av)

female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")

female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)

#table( droplevels(female_mouse.avs_for.analysis$strain) )
#there are cast strains too

Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Female mouse av. MLH1 count per cell ")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))

#female.specific.plot

musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))

#Mol
Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")

Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))

#Dom version 
Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")

Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))

```

<!-- female section -->

```{r female.strain.table, echo=FALSE}

female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#F.strain.table

#F.strain.table

# I don't know why most of the table is blank
```

```{r female.dif, echo=FALSE}

#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)
```

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

#variance plots
musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.female.VAR.plot

#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
#Mol.female.VAR.plot

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.female.VAR.plot

```

```{r Female.specific.glm, echo=FALSE}

glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)

summry.female.strain.effects <- summary(glm.female.strain.subsp)
#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)

MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)

#round(lm.female.strain_sum$coefficients[,c(1,4)],5)

#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
```

<!--
The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (glm; p= ).
 -->

```{r FEMALE.VAR.MODELS, echo=FALSE}
## within mouse variance models 
lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)

lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
lmVAR.female.strain_sum <- summary(lmVAR.female.strain)

#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]
#Q12
Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")

lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)

lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD

lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)

#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
#LEW sig

```

<!-- male section -->

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
#try customize the plot size

male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)

#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")

just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)

#table( droplevels(just.male_mouse.av$strain) )
#there are cast strains too

Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

#fomo.male_mouse.av

male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
male.plot

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))

# + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +
#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +

#make the Mol and
Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")

Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))

#Dom version 
Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")

Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
```

```{r male.strain.table, echo=FALSE}
M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        totNcells  = sum(Ncells, na.rm=TRUE),

                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm=TRUE),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#M.strain.table
#same issue as the female table .. ugh
```

```{r male.diff, echo=FALSE}
just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)

#how much greater are the High Rec gwRRs
just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))

#how much bigger are the high rec male means?
sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
#mean(sans_LewRec_male$mean_co)

PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)

MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)

SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
#mean(SKIVE_male$mean_co) /  
```

```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}

#variance plots
musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.male.VAR.plot

Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Mol.male.VAR.plot

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.male.VAR.plot
```

```{r Male.specific.glm, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect
glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)

#summry.male.strain.effects <- summary(glm.male.strain.subsp)
#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)

#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)

#1. determine which strain effects are sig

#M2
MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)

#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)

#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)

#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

```{r MALE.SPECIFIC.variance.model, echo=FALSE}
## within mouse variance models 
lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)

#all p values kinda high, MSM and TOM are slightly sig
lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
#non sig

lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
lmcv.male.strain_sum <- summary(lmcv.male.strain)
#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant

#Q12
Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")

lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
#NS

lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
#NS
```

Intial examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females. In females, _domesticus^G^_, _musculus^MSM^_, and _domesticu^LEW^_ had the greatest differences in means compared to other strains, `r G.female.diff`, `r round(LEW.female.diff,4)`, and 
`r MSM.female.diff` respectively. While in the male data the strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`, `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.

<notabley these strains come from Musc and Mol -- Mol natural hybrid of musculus and castaneus. this comments out text?>

The musculus and molossinus subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.

<!-- dom, low variation and close to the expected in
within the male means, there is a low degree of strain varaince in _Domesticus_ subspecies with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
-->

<!-- ranges of mouse means
In the _domesticus_ subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` 
in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_

The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_. 
-->

<!--
$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$ -->

<!--
The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively. 
-->

While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively. In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively. In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.

<!--(the figure is made from pooling MLH1 cells)
the male 
- the ranges of mouse means are displayed in Sup table X
- How the means were calculated
-The male gwRR is adjusted for the PAR
-The means for each strains are report in Table 1
-Normality was assessed for the MLH1 counts (include in Supplement histograms)
-->

# Evolutionary Framework

<to better understand this varation in RR within an evolutionary framework>

In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.

<!--
(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )

Check the DF of the number cited above
This is the break down of mice used within the Mixed model
`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`

add number of cells
`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
 $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

Setting up the data set for using the mixed model (base data set, HQ data set)

- only Mus musculus strains with sex matched observations  
- Male gwRR have been adjusted for the PAR
- Q5 excluded
- All the effects are un-ordered factors  
- male ages within 5 - 20 weeks
- Molossinus is coded as a subsp level

subspecies levels: 3, Musc, Dom and Molossinus
strain levels: Dom: 3, Musc: 3  Mol: 2
sex: 2 female male  
-->

**M1.Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

Our model, M1., was designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.

```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3, message=FALSE}
#THIS IS THE NEW MIXED MODEL
#should be able to use the loaded 
#DF.HetC.MixedModel.HQ

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1.MM.M1)$coefficients

MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)
```

All of the effects within the mixed model were significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).

```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
#strain = fixed
MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)

summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)

#table of just coeff anf pvalue
#round(summ$coefficients[,c(1,4)],5 )

#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
```

```{r MLH1.glm.M3, echo=FALSE, digits = 3}

MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)

summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)

#round(summ.strain.fixed$coefficients[,c(1,4)],5)

#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
```

 **M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$

 **M3. LInear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$

Following up with linear models which tested for (subsecies M2), and strain effects alone (M3)
both showed similar -- results -- **so we just report those from M3. also just report effects which were consisantly significant across both M2 and M3**

<they also have sig strain effects .. PWD strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`>

which identified several signifcant sex by strain interaction effects, ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`
)

<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.

<M3 also has a few less significant effects>
<not the reverse coefficient indexing -- this will give the name >

Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect). These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.

_domesticus^G^_ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`

<!--
also g strain effect/ dom^G^ strain effect 
`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[12,][4]` 
In both models strain by sex interaction effects were significant for (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively. 
-->
<!-- glms for sex specific data also run -->

A similar glm models (M4) was run on the sex specific data set, (just containing subspecies, strain and the interaction).

 **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$

For the male specific data, three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ). (with effect sizes ranging from 5, 7, and 2 foci respectively). Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.

While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.

<!-- transition gwRR to within mouse variance -->

## Within Mouse Variance in CO Count per Cell

In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).

Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))

Following the same basic framework we used for analyzing mean number of MLH1 foci per cell, we applied a set of mixed and linear models -- to test effects (sex differences for the within mouse variance in MLH1 foci per cell).

 **M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

 **M6. Linear Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex * strain + \varepsilon$$

 **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$

```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}

MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1var_M1)
LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value

LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)

#CV
MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value

LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)

#USE Q12 (Q123) table
#Q12_mouse_table_w.Ages
#Q12 var
MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )

LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value

LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)

#Q12 cv
MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )

LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value

LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
```

```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )

#summary(MLH1var_M3)
anova(MLH1var_M3)  
as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))

MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)

#summary(MLH1cv_M3)

anova(MLH1cv_M3)
as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))

#Q12 var
MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )

#summary(MLH1var_M3_Q12)
anova(MLH1var_M3_Q12)  
as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))

#Q12 cv
MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )

#summary(MLH1cv_M3_Q12)
anova(MLH1cv_M3_Q12)#sex large F

as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))

```

```{r M3.MLH1.var, echo=FALSE}

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)

### Q12 var
MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)

#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```
<!--
These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
-->

For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively). Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e. background noise). For this reasons, we replicated the model analysis using a subset of cells with higher quality scores. These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ). Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males. While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`) for the model with the full dataset. While in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes. For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell. This limits the power for detecting significance differences across groups, <but the main/general pattern, is that females have significantly higher within mouse variance compared to males. This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.

<!--
motivation / background: maybe checkpoint / regulation
limitations: observations at the mouse level, decreased power to estiamate the variance --

$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
<!--
RW's main qtl paper used the within animal variance,

@article{wang2017genetics,
  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
  author={Wang, Richard J and Payseur, Bret A},
  journal={Genetics},
  volume={206},
  number={4},
  pages={1841--1852},
  year={2017},
  publisher={Genetics Soc America}
}
references for phenotypic variance connected to robustness to environment, increase in developmental consistency

Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
phenotypic variance leading to adaptation - 
(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-->

< SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE >

# Variation in DSB Number Per Cell

```{r DMC1.setup, echo=FALSE}
#load DMC1 csv
#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)

DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_category(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
                 .(category),
                 summarize,

                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.L$stage <- "L"

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "Z"

DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)

#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
#            row.names = FALSE)

DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
```

```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
#think of better color scheme
#make the jitter col thinner or make the stacked dots.
#remember that box-plots add outlier dots.
#change y labels

#experimenting with ggboplot
everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
          color = "stage", palette = "jco",
          add = "jitter", short.panel.labs = FALSE)
  #library gpubr has functions to add in sig/test bars.

#change order

DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

DMC1.boxplot
```

```{r plot.stage.ttest, echo=FALSE, warning=FALSE}

anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
#

DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=rec_group), position=position_jitterdodge()) +

  #facet_wrap(~stage, nrow=1) +

  ggsignif::geom_signif(
    data=anno_df_stage, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")

ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )

DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")

ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )

DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")

ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
```

```{r plot.t.test, echo=FALSE, include=FALSE}

  #this will add neat bars, but still need to figure out what groups ect it's actually using
#THE stat_compare function adds a arm thing and pvalue for 
#  stat_compare_means(comparisons = my_comparisons)

#status - legend, expression y-axis, facet, facet_wrap SYMBOL
#legend - stage. y=total count, FACet strain

#add column to distinguish high vs low rec strains
DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))

DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)

# annotation table with adjusted pvals and y-position of the labels

anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
#

DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=stage), position=position_jitterdodge()) + 
  facet_wrap(~stage, nrow=1) + 
  ggsignif::geom_signif(
    data=anno_df, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

#example for ggplot+ test from here
#https://github.com/kassambara/ggpubr/issues/65

#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/

```

```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
#2. t-tests for the DMC1 counts 
#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo

#high vs low group (L)
ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value =  0.002145

#high vs low group (Z)
ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value = 0.6628

#dom vs musc (L) (2 subsp vs)
ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
# p-value = 0.003016

#dom vs musc (Z)
ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
#p-value = 0.09588

#KAZ vs Dom (Z) (only 1 L cell)
ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )

#p-value = 0.004148
#Kaz Z are higher than Dom
```

```{r DMC1_anova, echo=FALSE, warning=FALSE}
#1. ANOVA, is there a difference in DMC1 counts across strains?

#unorder strain
DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )

mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])

l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
```

```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
#make category averages from AP_mouse_table.HQ
mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co)
)

DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
                   category == "WSB male" |
                     category == "KAZ male" |
                  category == "MSM male" |
                    category == "PWD male")
#attach this mini DF to the DMC1 data

DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count

DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")

DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")

MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor

MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237

##RATIOS
DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z

#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")

#DMC1.mean.table
DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))

colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")

WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1

G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1

MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1

PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1

DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )

DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]

kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")

DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )

DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )

```

<!--
explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
-->

In an attempt to locatize the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes. DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days). A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_. One mouse was sampled for each strain.

For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = ), which reflects the expected patterns across the timing of meiosis.

Confirming the prediction that (more crossovers are associated with more initial DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r format(l.anova.pval, scientific=F)`).

The differences in DMC1 foci were not significant for zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).  <!-- which maybe be attributed to noise effects

The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->

The ratios of average number of MLH1 and DMC1 foci per cell -- have  .... biological significance

The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).

These results indicate that the changes -- which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).

<!--
The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`. The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`. `r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M. m. domesticus^WSB^_ , _M. m. domesticus^G^_, _M. m. molossinus^MSM^_, and  _M. m. musculus^PWD^_ respecitively. 
-->

```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}

#Subsampling code -- don't report

#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells. Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups. Permutations were done for both cell stages. The actual p value is in red.

#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.

#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
# the 2 groups

#make the 2 pools
High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]

Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]

#permute drawing / subsampling and performing the t-test
#put the p.value results into a DF!

#permut.SC.cv.diff  <- as.data.frame(
#sample(THE.DF, SAMPLE.SIZE, REPLACE?)

#permuting samples -- and re-licating t-tests

#nrep <- 1000

#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
bunchOpvals.L <- replicate(1000, 
    # t.test
    t.test( sample(High.L.pool, 25, replace = FALSE),
            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
          )
bunchOpvals.L <- as.data.frame(bunchOpvals.L)
bunchOpvals.L$stage <- "L"

bunchOpvals.Z <- replicate(1000, 
    # t.test
    t.test( sample(High.Z.pool, 25, replace = FALSE),
            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
          )

bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
bunchOpvals.Z$stage <- "Z"

#could merge the lists of p-values
replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'

#ASSUMPTIONS; normality, equal variance

#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.

#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")

#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
#plot.pvals.L
#plot.pvals.Z

#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
#n use the var.equal = TRUE
#https://www.statmethods.net/stats/ttest.html
```

```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}

#these are histogram plots for testing Normality shape -- move to supplemenatry section

#- check that distributions are close to normal
#- check for equal variance.

DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")

DMC1.histogram
#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
#bartlet's test, Levine's test

#The counts look close to normal.
#Potentially figure 2. Using a function to #display the p-vales for mean comparison

```

<!--
interpertation notes, 
the number of DMC1 across stages (L and Z stages) are driven by different equations

Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.

Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells

*the repair decision can happen at both stages and remove DMC1. formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)

back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.

expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
OR

variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
-Baier et al 2014 supports this model (Chmr axis / DNA loop model)
-->

<!--
# Rapid male specific evolution

list out the points which support rapid male specific evolution for MLH1 rates.

number above bars are the raw counts for bivalent classes.

Potential figure 2 (eventually). Try to make the figure taller.
-- add more data to make sample sizes more equal.
-- add space to indicate where female data is missing

SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
What's up with Lew females? there's lots of 2COs, also WSB female.

This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells. If these data are random samples of bivalents, the X

ToDo:
- within strain across sex,
- males, across high and low groups
- look at mouse level proportions
-->

# Chromosome Class Proportions

```{r chrm_prop_data, echo=FALSE, warning=FALSE}
#shut off eval for now because of wd / path error
#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)

chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)

#try adding the BivData version of the proportions

#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)

#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)

chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
chrm.class.DF <- add_mouse(chrm.class.DF)
chrm.class.DF <- add_strain(chrm.class.DF)
chrm.class.DF <- add_sex(chrm.class.DF)
chrm.class.DF <- add_category(chrm.class.DF)
#chrm.class.DF <- add_euth_date(meta.data.file2)
#chrm.class.DF <- add_age(meta.data.file2)

#table(chrm.class.DF$category)
#str(chrm.class.DF)

#summarize by category
proportion_stats_category  <- ddply(.data=chrm.class.DF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),

                 n0CO = sum(n0CO, na.rm = TRUE),
                  n1CO = sum(n1CO,na.rm = TRUE),
                  n2CO = sum(n2CO,na.rm = TRUE),
                  n3CO = sum(n3CO,na.rm = TRUE),
                  n4CO = sum(n4CO,na.rm = TRUE),

                 per0CO = sum(n0CO) / tot_biv,                 
                 per1CO= sum(n1CO)/ tot_biv,
                 per2CO= sum(n2CO)/ tot_biv,
                 per3CO= sum(n3CO)/ tot_biv,
                 per4CO= sum(n4CO)/ tot_biv,

                 mean.MLH1.foci = mean(MLH1.sum.check),
                 mean.MLH1.foci2 = mean(nMLH1.foci),
                 meanQ = mean(quality)

)
#this is the table that needs melting
#but is also easy to redo by hand

#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
```

```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}

#add subspecies to these values
melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)

#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)

melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]

#add the subsps

melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )

melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])

chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))

#need a second DF / the first df
#+ annotate() text 
#annotate()text

#proportion_stats_category,  tot_biv, ncells
#add number of cells and bivalents in the corner

#chrm.class.barplot
```

```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}

#prop.test(x, y)  x number of successes, y total  
#strain sums of n
#look up continuity correction

#mouse.level.BivData_table <-- has more male strains
#Mouse.Table_BivData_4MM <- only sex matched strains

#perform test across strain 
#removed all lines with no hand.foci.count entry
Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,

                  nbiv = sum(tot.biv ),              
                 n0CO = sum(n0CO ),
                  n1CO = sum(n1CO ),
                 n2CO = sum(n2CO ),
                  n3CO = sum(n3CO ),
                 n4CO = sum(n4CO )
                          )

Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)

for(g in 1:length(Chrm.prop_strain_table$nbiv)){

  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
                                     Chrm.prop_strain_table$n1CO[g],
                                     Chrm.prop_strain_table$n2CO[g],
                                     Chrm.prop_strain_table$n3CO[g] ))  
}

Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
#there are still some wacky entries -- but much fewer than before

chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")

female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
#the pvalues for 0 are not sig dif
#the p values for 1 and 2CO are sig diff

#seperate by subsp
chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")

Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
#WSB has fewer 2COs than other 3 strains

Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
#KAZ has slightly more 1CO than the other two

#high.female_G.LEW.
chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")

high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )

#Low.female.G.LEW.
chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")

low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )

#all females

chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )

male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
#the pvals for these 3 class are sig dif

#seperate by subsp
chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")

Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
#non of the Dom proportions are different

Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
#all v different

Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )

#high vs low
chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))

chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked

chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)

Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
#the low strains do vary slightly in their 1CO and 2CO proportions

#includes SKIVE
Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
#all very sig p values

High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
#these proportions are v different

#no SKIVE
High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
#MSM and PWD only slightly sig dif 2CO

#MSM-MOLF
MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

#PWD-KAZ
PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )

PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )

PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )

```

```{r Chrm.props.plot2, echo=FALSE}

#barplots for proportions of chrm classes

#consider adding this the the setup block (could have done this to category level table)
mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv

mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv

mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv

mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv

mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv

#DF needs to be melted -- so they all in 1 col
#per code, I've done this by hand

#subset DF for ease
skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
col.names.list.99 <- colnames(skinny.mouse.level)

#all of the cols to use
MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",

               #leave out the Foci
                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],

  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))

MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]

#no filename for add mouse, strain subsp
MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)

#then this DF has to be summarized into categories
chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
                             mean.percent.type = mean(value)
                             #sum is the wrong!
                             #used average percentage, might consider a different summary
)

#can't apply function use this
chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",

          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",

                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",

             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))

chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
                                                                 "Musc",
                                                                 "Mol",
                                                                 "Spretus","Spic",
                                                                 "other") )
#ggplot

chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
#chrm.class.barplot
chrm.class.barplot
```

```{r Chrm.Prop.misc.calulations, echo=FALSE}

total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)

#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
```

<!--
- This is motivation to investigate more traits at the chromosome level
1.	Basic stats - bivalent data numbers 
2.	Rang of co per chrm, mostly 1co 2co
3.	High rec strains have evolved a higher proportion of 2CO
For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
-->

In order to decompose the cell wide rate, we decided to look at the proportion of chromosome classes (ie by the number crossovers on the same bivalent). A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).

The first attern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes. <these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>

< note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?) >

The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`). This relpicates what has been notice in the literature (cite stapley).

Most of the bivalents are 1CO and 2CO class bivalents -- and the proportion of these two classes distinguish high and low rec male strains. (distinguish the high rec males from minimal rec groups)

(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class. The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.  *(much smaller proportions and variation in the fringe classes*

**2. most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**

**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**

PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).

While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
 and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).

**3. these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**

<Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`). This significant p value indicates that at least one of the six strains had significant differences).>

<SUMMARY / TRANSITION to single bivalen measures>

# Single Bivalent Level Results

In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining features of the recombiantion landscape at the single bivalent level. recombination landscape refers to the i) total length of the SC-AE (potential area for COs), ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.

From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software. After the human curation step, `r length(new.Curate.FULL.org$Obj.ID)` single bivalent measures were left. (make sure these are the hand.foci bivalents).

how do these agree with other patterns of MLH1

```{r algorithm_performance_stats, echo=FALSE}

#sort by category,

#compare total boxes and rate passed
#Curated_BivData  -- pass rate?

#number of mice, number of cells
#mean number of isolate bivs per cell,

data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
```

Table X illustrates the break down of single bivalent observations by category 
Error compared to human measures and other details have been discribed in (Peterson_2019).

While the automated software doesn't isolate all bivalents/chromosomes from each cell (on average 17), we assume that the isolation process is not biased. Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.

We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?

These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.

The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.

# Q1 SC Lengths

<!-- breif background on longer female SC  (mouse, human) Gruhn-->

Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of PWD and MSM. A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.

EXPLAIN WHY USING TOTAL SC AREA --

turn out comparing the means of single bivalents is complicated by the fact we can't identify any of the bivalents. -- so we use total sc

```{r Q1.Short.Long_biv.setup, echo=FALSE}

Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean_SC = mean(chromosomeLength)
                         )

Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_species(Q1.real.short_mouse_table)

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
Q1.real.short_mouse_table$species <- factor(Q1.real.short_mouse_table$species, ordered = FALSE)

Q1.real.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean_SC = mean(chromosomeLength)
                         )

Q1.real.long_mouse_table <- add_strain(Q1.real.long_mouse_table)
Q1.real.long_mouse_table <- add_sex(Q1.real.long_mouse_table)
Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)

```

The plots below show the distributions of mouse average total SC area.

```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}

Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

#CLEAN UP THE FULL DATA SET

#adjust order of plot here
Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
  "WSB","G","LEW","PERC",  
"PWD", "SKIVE",  "KAZ",
     "MSM", "MOLF",   
                     "CAST",
                         "SPRET","SPIC" ), order=T )

#nested boxplots, pooled Bivalents

Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")

Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")

Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")

```

```{r Q1.tot.SC_show, echo=FALSE}

Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")

Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
   "WSB","G","LEW","PERC", 
  "PWD", "SKIVE",  "KAZ",
  "MSM", "MOLF"), order=T )

Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])

#order the strains
Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
  arrange(strain) %>%
  mutate(mouse = factor(mouse))

Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])

#add category mean
#totSC.strain.av.table %>% filter
totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )

totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )

Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")

Tot.SC_mouse.av_Dom

totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )

Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")

Tot.SC_mouse.av_Musc
```

these are the number of mice for each of the categories

`r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )`

```{r Q1.total.SC_t.test, echo=FALSE}

#run t-test for male female avergage for each strain

#WSB
Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
#G
Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
#LEW
Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)

#PWD
Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)

#KAZ
Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)

#SKIVE
#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")

#error not enought female observations

#MSM
Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)

```

The full data set of SC lengths illustrates the general pattern than females have longer SC-AE.

The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).

Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)

$$mouse \ average \ total \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE}

Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")

#str(Q1.lmer_totSC.mouse.av.table)

#unorder factors
Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)

#TOTAL SC AREA!!
lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)

lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)

lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)

lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]

lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]

lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]

lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value

#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
sex.results =  lmer.Q1_mus.av_totSC_sex, 
               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
lmer.Q1_mus.av_totSC_strain)

#Q1.lmer.totSC_results

#G
```

there are more significant effects for strain and subsp for the mixed model of mouse.av total sc

the 3 fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`)

These effects were followed up the glms

**M2.Linear Model** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$

**M3. Linear Model** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$

```{r Q1.glm_totalSC, echo=FALSE, }
#total-SC

#Q1.lmer_totSC.mouse.av.table

Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
#summary(Q1.glm_totSC_M2)
#subsp effect, many strain, #male

#G * male v.sig and male*LEW

Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
#summary(Q1.glm_totSC_M3)
#LOT OF SIGNIFICant strain effects in this model

```

The results from those models was that there were 
...lots of significant effects.  (G male, )

The summary for the total SC area comparison is...

<!-- transition to short-->

The short bivalent dataset was used -- because there are no confounding factors of the XX -- so makes hapier comparisons across male and females

```{r show.short.biv, echo=FALSE}
Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")

Q1.SC.short.mouse_scatter_boxplot

#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
#skiveQ1.SC.short.mouse_scatter_boxplot
#there are 2 lowish female means,

```

```{r Q1_SC.biv.short_ttest, echo=FALSE}

#run t.tests for the mouse av. for reduced single bivalent datasets
#WSB
Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)

#G
Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)

#LEW
Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)

#PWD
Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)

#KAZ
Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)

#SKIVE
Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)

#MSM
#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
#1 male, 3 females measures
#try to merge the manual - whole cell observations (at least 2 males)

```

The pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data

with the mouse average for 'short' bivalents

(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and `r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
for WSB, G and LEW respectively.

(t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for PWD, KAZ, SKIVE

for the short bivalents are not significantly differen, for this strain there are 4 female and 4 male

** there are more SKIVE female mouse observations -- but they are filtered out because I excluded them from the MLH1 dataset.  -- for low quality -- they might have good total SC spreads -- // but

* it think total sc only has 1 female observation but the short bivalent has 4 skive females

<!--
for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
-->

Mixed models for the sex differences in mouse averages for short bivalents (reduced dataset)

**M2.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$

```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE}

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)

#summary(Q1.glm_short.SC_M2)

summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)

#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)

#skive strin very slight sign
#male highly sig
#skive * male slightly sig
```

skive strin very slight sign, male highly sig,  skive * male slightly sig

**M3.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value

#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)

#Q1lmer.MM_SC_short_results

```

Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). interaction effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`)

**M3.** $$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$

```{r Q1.glm_short.SC_M3, echo=FALSE}

Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)

summary(Q1.glm_short.SC_M3)

summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)

round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
#sex male very sig -- only sig
```

male is the only significant p value from this model.

<!--
Because of confonding effects across chromsomes of different physical lenghts, we attempt to ... reduce the error from mixing chromosomes by constructing reduced bivalent sets, for the longest 5 and shortest 5 bivalents from the same cell. The mouse means for these two data sets were then calculated.
-->

Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.

Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.

The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.

In summary .... females (oocytes) have different chromatin compaction patterns compared to males.

```{r show.long.biv, echo=FALSE, include=FALSE}
#mouse av
Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")

```

```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}

#Q1.real.long_mouse_table

Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)

#G
Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)

#LEW
Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)

#PWD
Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)

#KAZ
Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)

#SKIVE
Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)

#MSM
#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)

#run t.tests for the mouse av. for reduced single bivalent datasets
```

```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}

#1CO  lmer.Q1_SCLength_M1
Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))

Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value

Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
sex.results =  Q1.lmer_SCLength.1CO_sex, 
               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
Q1.lmer_SCLength.1CO_strain)

#2CO
Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))

Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value

Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
sex.results =  Q1.lmer_SCLength.2CO_sex, 
               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
Q1.lmer_SCLength.2CO_strain)
#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
```

```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}

Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
```

```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}

#removing these plots for space

Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
#Musc.all
Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)

#mouse av
Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length")+facet_wrap(~sex)+ylim(c(75,125) )

#BivData Level
Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")

Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))

#BivData Level
Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")

Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))

```

```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE}
#removed these plots for space

male.long.biv <- real.long.bivData %>% filter(sex== "male")

Q1.long.biv_cell_table <- ddply(real.long.bivData, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.sc = mean(chromosomeLength)
                         )
Q1.long.biv_cell_table <- add_mouse(Q1.long.biv_cell_table)
Q1.long.biv_cell_table <- add_strain(Q1.long.biv_cell_table)
Q1.long.biv_cell_table <- add_sex(Q1.long.biv_cell_table)
Q1.long.biv_cell_table <- add_subsp(Q1.long.biv_cell_table)

Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")

#mouse av for long biv
cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.sc = mean(chromosomeLength)
                         )

cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)

cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")

#mouse av for long biv
mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,

                        n_obs = length(Obj.ID),
                         mean.sc = mean(chromosomeLength)
                         )

mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)

mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av. \nMale Long Biv SC Lengths")

```

```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE, include=FALSE}
###
lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)

Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value

Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
sex.results =  Q1lmer.MM_SC_long_sex, 
               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
Q1lmer.MM_SC_long_strain)
```

# Q1. General CO Positions

```{r Q1.F1pos_long.short_setup, echo=FALSE}

Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")

Q1.long.1COpos_mouse.av <-  ddply(Q1.long.biv_1CO, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean_SC = mean(chromosomeLength), 
                        mean_F1pos_PER =  mean(PER_Foci_1),
                         mean_F1pos_ABS =  mean(Foci1)
                         )
Q1.long.1COpos_mouse.av <- add_strain(Q1.long.1COpos_mouse.av)
Q1.long.1COpos_mouse.av <- add_sex(Q1.long.1COpos_mouse.av)
Q1.long.1COpos_mouse.av <- add_subsp(Q1.long.1COpos_mouse.av)

Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")

Q1.short.1COpos_mouse.av <-  ddply(Q1.short.biv_1CO, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean_SC = mean(chromosomeLength),
                        mean_F1pos_PER =  mean(PER_Foci_1),
                         mean_F1pos_ABS =  mean(Foci1)
                         )

Q1.short.1COpos_mouse.av <- add_strain(Q1.short.1COpos_mouse.av)
Q1.short.1COpos_mouse.av <- add_sex(Q1.short.1COpos_mouse.av)
Q1.short.1COpos_mouse.av <- add_subsp(Q1.short.1COpos_mouse.av)

```

<!-- motivating

To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
-->

We focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference. We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ). (all chromosomes in house mouse are acrocentric)

Chromosome size effects are a confonding factor for CO position, with shorter chromosomes having a more uniform landscape compared to larger chromosomes (cite). To account for this we again use the reduct bivalent sets (long.biv and short.biv)

Below are the general patterns / plots -- short and long bivalent groups are shown to try and account for the chromsomes size effect.

```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
#main figure of density plots,

MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain

#clean up all the non 1 to 4 hand.foci counts
MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]

MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]

# remove the na's
MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]

CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
CO.Pos.density_1CO

#category specific plots, don't run for now.
Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
#Musc male
Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]

CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
#Dom male
Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]

CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
```

```{r Q1show.F1pos_Full, echo=FALSE}

Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \n all bivalents")
Q1.mouse.av.biv_scatr_box

```

```{r Q1_1COpos.t.test, echo=FALSE}
#testing,
#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents

#WSB
Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
#G
Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
#LEW
Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)

#PWD
Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)

#KAZ
Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)

#SKIVE
Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )

#MSM
Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)

#Single Biv Level
#single.biv.1CO_pos_t.test <- t.test(

 # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],

#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
#  )

#t.tests for mous.av_nrm.1CO.pos 
```

In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (with female)
are highly significantly more medially placed (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively). With the musclus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM. Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).

<!-- lmer and glm here -->

$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1.CO1pos_lmer.M1, echo=FALSE}

#make sure this dataset has all the right strains
Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")

lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)

lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )

lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )

lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]

lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value

lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
sex.results = lmer.Q1.nrm1CO_M1.sex , 
               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)

lmer.Q1.nrm1CO_M1_results

#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)

```
For the Mixed model, sex and strain are significant effects for the normalized single foci position.

$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$

```{r Q1.CO1pos_glm.M2, echo=FALSE}

#in case factors aren't ordered
#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
summary(Q1.glm_F1pos_M2)

summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)

#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)

#DOM
Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1.glm_nrmF1pos_M2_Dom)

#MUSC
Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1.glm_nrmF1pos_M2_Musc)

#MOL
Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1.glm_nrmF1pos_M2_Mol)

```

$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$

```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=FALSE}

summry.Q1.glm_F1pos_M2

Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
summary(Q1.glm_1CO.pos_M3)

summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)

#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
```

Lew and KAZ had slightly significant strain speicific effects across the two glms (these results will be explored more fully in the Q2 section). The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).

<!--
-full 1CO tests

long 1CO tests

short 1CO tests

There are a few traits that fall within the CO positions  
- 1CO normalized position  
- sis.co.ten (sorta interference)  
- centromere and telomere distance  
**Heterochiasmy**

2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
  C) Centromere and telomere distances will be sexually dimorphic.
**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
-->

```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}

Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nShort Biv")
Q1.short.biv_scatr_box

Q1.long.biv_scatr_box   <- ggplot(Q1.long.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nLong Biv")
Q1.long.biv_scatr_box
```

```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}

#make the same plots as above but with the long biv data set

#mouse.av.long.biv.table
Q1.av.long.biv_scatr_box   <- ggplot(Q1.long.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nLong Biv")
Q1.av.long.biv_scatr_box

Q1.av.short.biv_scatr_box   <- ggplot(Q1.short.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nShort Biv")
Q1.av.short.biv_scatr_box

```

```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}

MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]

CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
CO.Pos.boxplots_1CO

```

```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}

MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]

CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
CO.Pos.boxplots_1CO

```

```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
#IGNORE THIS FOR KNOW, THIS

#filter all Curated biv data for 1 strain
#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs

# the stardell paper sums the number of COs for bins (10)

#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
#but then the model would be close to logistic

#summing the number of COs per bin, losses the chrom specific data....
#!! need to use a melted Data frame to capture all nrm.foci obs
#wsb.ttest<- mouse.avs_4MM %>% filter()
#start with same chrm classes
#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci

WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)

Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)

PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)

Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)

anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig

anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low

anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1

#or AIC(model1,model2)

```

```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}

#make a table for the sex specific values of the normalized F1 positions

#for strain, sex -- Curated_BivData
#or (Mouse.Table_BivData_4MM)
#norm.first.foci.pos

Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,

                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
                  n1CO = sum(n1CO )

                          )

```

Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs. This plot shows the sexual dimorphism in the density plots.

Consider adding annotate_text for the number of observations in each category.
think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.

```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}

#real.long.bivData
#MELT_Curated_BivData

#these isn't long.biv data
MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]

real.long.bivData
MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]

MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]

MELT_long_biv <- add_sex(MELT_long_biv)
MELT_long_biv <- add_strain(MELT_long_biv)
MELT_long_biv <- add_subsp(MELT_long_biv)
MELT_long_biv <- add_category(MELT_long_biv)

long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]

Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
Dom.male_MELT_long

long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]

Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
Musc_male_MELT_long

#there are v few observations in this dataset
```

<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->

These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males). They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains. MOLF males have much more medial positions than other strains.

the distribution of SC lengths and sis-coten seems very different across sexes

The mixed model data should only come from 1CO bivalent data.

the mouse average foci1 pos is more significant in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.

```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}

#LOGISTIC REGRESSION for FOCI POSITION

#mouse av, norm.first foci is from 1CO bivalents
log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_1CO_pos_mouse.av)

#single biv level
log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
summary(log.reg_1CO_pos_biv)

```

<!--
### Siscoten

The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.

The logic of how the sis-co-ten metric is outlined in the figure below. The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO. This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.

![from (Lee, J. (2019). Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?. Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)

```{r show.siscoten, echo=FALSE}
#think of a good way to display this

sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
sis.co.ten.plot_WSB_female

sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
sis.co.ten.plot_WSB_male

sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_LEW_female

sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_LEW_male

sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_PWD_female

sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_PWD_male

Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]

jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_PWDf

jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_PWDm

jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_WSBf

jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_WSBm
```

Males have much clearer separation of siscoten across chrm classes. This is emphasized when SC length is also plotted. It seems like musc males have higher amounts of this metric compared to Dom males.

To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.

BUT females have a greater range -- so maybe it's just a scale issue.

```{r show.norm.siscoten, echo=FALSE}

Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength

Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
Nrm.sis.co.ten.plot_WSB

Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
Nrm.sis.co.ten.plot_G

Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
Nrm.sis.co.ten.plot_LEW

Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
Nrm.sis.co.ten.plot_PWD

Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
Nrm.sis.co.ten.plot_KAZ

Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
Nrm.sis.co.ten.plot_SKIVE

Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
Nrm.sis.co.ten.plot_MSM

```

I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.

```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}

#glm for siscoten

#mouse.avs_4MM  - mouse averages with mean sis.co.ten

#MM
lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )

#sex and sex*subsp interaction effect are significant for mean sis.co.ten

lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(lmer_MM_siscoten)

#thing1  <- summary(lmer_MM_siscoten)
#drop1(lmer_MM_siscoten, test="Chisq")

rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
rand_siscoten_MM
#slightly sig
```

The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.

> Is the heterochiasmy prediction met?

Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.

```{r siscoten.test.poly, echo=FALSE, eval=FALSE}

#move this to the set up blocks

mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )

#test high vs low Musc strains
#mouse.avs_4MM

```

```{r log.reg_siscoten, echo=FALSE}

#replace the table

male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )

Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )

#SISCOTEN
#mouse av, norm.first foci is from 1CO bivalents

log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,

                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_SIS.co.ten_mouse.av)

#single biv level
log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_Sis.co.ten_biv)

```

```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}

#bivData --

t.test_mouse.av_siscoten <- t.test(

  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],

  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
  )

#1CO
t.test_mouse.av_siscoten_1CO <- t.test( 
  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],

   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
  )

#2CO
t.test_mouse.av_siscoten_2CO <- t.test( 
  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],

   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
  )

#3CO
t.test_mouse.av_siscoten_3CO <- t.test( 
  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],

   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
  )

####
#Single Biv Level -- from the main 
#pull the single BivData -- MUSC

#check that these have the extra male strains

t.test_single.biv_siscoten <- t.test(

  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],

   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )

#1CO
mouse.avs_4MM$mean.siscoten_2CO

#2CO

#3CO

```

All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.

### Telomere and centromere Distance

My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC). In the plots below each point is a single bivalent. I choose not to use the mark for centromere because it seems noisy and inconsistent...

```{r telo.dist, echo=FALSE}

sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)

Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]

telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))

telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
telo_jitter

nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
nrm.telo_jitter

#why do a different number of rows removed each time??

```

Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females. In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range. In females the ranges of telomere distances have much more overlap.

```{r MM.telo, echo=FALSE}
#HetC

lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )

drop1(lme_MM_telo.dist, test="Chisq")

as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )

#sex and sex*subsp interaction effect are significant for mean sis.co.ten

lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
#anova(lmer_MM_siscoten)

#thing1  <- summary(lmer_MM_siscoten)
drop1(lmer_MM_telo.dist, test="Chisq") 
rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
rand_telo.dist_MM

```

Mixed model result summary:

```{r log.reg_telo.dist, echo=FALSE}

Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )

#TELO.DIST
#mouse av, norm.first foci is from 1CO bivalents
log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,

                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_Telo.Dist_mouse.av)

#single biv level
log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_Telo.Dist_biv)

#PER versions
log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_Telo.Dist.PER_biv)
```

```{r t.test_Telo.Dist, echo=FALSE}

#mouse av
t.test_mouse.av_telo.dist <- t.test(

  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],

  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
  )

#Single Biv Level
single.biv_t.test_Telo.Dist <- t.test(

  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

#Telo Dist PER
single.biv_t.test_Telo.Dist_PER <- t.test(

  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],

   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

```

```{r cent.dist_show, echo=FALSE}
cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))

cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
cent.dist_jitter

norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
norm.cent.dist_jitter

```

```{r MM.cent.dist, echo=FALSE, eval=FALSE}

lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )

lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

exactRLRT(lmer_MM_cent.dist)
```

```{r log.reg_CENT.dist, echo=FALSE}

#CENT.DIST
#mouse av, norm.first foci is from 1CO bivalents
log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,

                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_CENT.Dist_mouse.av)

#single biv level
log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_CENT.Dist_biv)

#PER versions
log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_CENT.Dist.PER_biv)
```

```{r t.test_CENT.Dist, echo=FALSE}
#CENT.DIST
#mouse av
t.test_mouse.av_Cent.dist <- t.test(

  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],

  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
  )

#Single Biv Level
single.biv_t.test_Cent.Dist <- t.test(

  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

#Cent Dist PER
single.biv_t.test_Cent.Dist_PER <- t.test(

  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )

```

The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.

Females have more overlap in the distributions of centromere distances across chromosome class compared to males.

-->

#Q1. IFD

In order to test if there are sex differences between crossover interfernce, a major determinant of the positioning of crossovers along chromosomes, we examined ... ran models with the interfocal distance (IFD) of two foci on the same bivalent.

-We focus on observations from two-crossover bivalents for more comparable observations.
-comparisons of all IFDs of mulit crossover bivalents is also included
Interference is a major determinant of the positioning of chromosomes

<!-- 
Brief lit review for IFD / interference expectations  
**Petkov et al 2007,  CO interference underlies sex differences in RR**

"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb. However, the interference distance is the same in terms of bivalent length. We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes 
at the pachytene stage of meiosis."

Chrm1 genetic map, 
(other human and mice refs 
Tease n Hulten 2004: -- no difference between MLH1 foci in males and females

DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
<!--
ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols

explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
-->

<!--  set up -->

<!--  plots-->

```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}

#first draft of the IFD comparisons
#make the mouse boxplots -- instead of geom_points

#Only use 2CO bivalents
Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]

#boxplots
boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
#order is out of wack, but -- good for now

boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, PER")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))

first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)

first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))

#+theme(legend.position="none")
```

```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)

long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
```

<!--t.tests -->

```{r Q1.IFD_Ttests, echo=FALSE}

#below are calq just from 2COs, have no fear
#mean_IFD.2CO_ABS
#mean_IFD.2CO_PER

#FULL all strains
Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,

                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
#p-value = 0.07

Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,

                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
#p-value = 1e-12

```

The raw IFD measures for 2CO bivalents are slightly significant (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`). While the normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`)

(interpertation statement)  Males have stronger interference

<!-- lmer/ glms-->

### Mixed Model Tests, Fixed Effects

Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.

$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1.IFD.M1, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values

#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

#M1_IFD_2CO
lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value

Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)

#M1_IFD_2CO
lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value

Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
#don't understand why 2 rows?

#M1_IFD_2CO
lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)

#these 2 below are LR tests
#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
#exactRLRT(lmer.IFD.PER_MM.M1)

```

The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively. That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.

I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure. The tables below are from anova( for the lmer model ). Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER

$$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$

```{r IFD_HetC_M2, echo=FALSE}

glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)

summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)

#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)

#PER
glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)

summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)

#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)

```

$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$

```{r IFD_HetC_M3, echo=FALSE}

glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)

summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)

#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)

#PER
glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)

#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)

```

For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.

the interaction effects were slightly significant for both raw and nrm.IFD.

The most significant value was the sex effect for nrm.IFDs.

The the random strain effect was not significant for either model.

For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw. The other effects are for PWD ansD SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance. For the M3, only SKIVE*male effect is close to significant

For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males. SKIVE*male is the only other consistantly significant effect, which also increases the nrm.IFD measure.

Overall There's a low amount of significant effects across the 2CO IFD measures. This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.

Dive deeper into the sex specific pattern for each strain.  Below are code chunks which show the unusual sex specific results for IFD measures. The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD. The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.

For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?

```{r Comp_Ratios, echo=FALSE}

#calq ratios of belo 30

top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)

bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)

ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)

top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)

bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)

ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)

#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?

#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")

DF.strain.list <- list(
  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
  MSM.Bivalents.DF, MOLF.Bivalents.DF
  )

entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))

i = 1
for(x in (DF.strain.list) ){

  #print(head())

  #these filters won't apply to the indexed DF
  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)

  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)

  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)

  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)

  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)

  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)

  i = i+1

  }
#entr_DF

entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio

```

Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.

# Q2 SC Length

<!-- place set ups for total SC, short biv and long biv here -->

I made a sereies of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromsomes (possitive correlation of sc length and CO number).  Yet, given the complications from the chromsomes proportions outlined above, comparing the bivalents across chromosome classes is complecated

```{r juvimice.list.copy, echo=FALSE}
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )

ppmm <- OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$mouse  %in% juvi_mice, ]   
```

```{r Q2.SC.short.long.setup, echo=FALSE}
table(male.long.biv$strain)
#the Long biv dataset is quite unbalance (esp MSM)
male.long.biv <- real.long.bivData %>% filter(sex== "male")

#mouse av for long biv
cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)

cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)

Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                         #mean.first.1CO
                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
                         )

Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)

Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)

#divide musc males into groups
Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
                                           0)))

Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))

male.short.biv <- real.short.bivData %>% filter(sex== "male")

Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                        #mean.first.1CO
                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

                         )

Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)

Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)

Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
                                           0)))

Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
```

In order to test if sc lengths (sc area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.

The average sc lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confonded by the different proportions across strains.

First we examine the pattern of SC length variation across chromosome classes.

```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}

Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

#adjust order of plot here
Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
  "WSB","G","LEW","PERC",  
"PWD", "SKIVE",  "KAZ",
     "MSM", "MOLF",   
                     "CAST",
                         "SPRET","SPIC" ), order=T )

Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)

#nested boxplots, pooled Bivalents
#x axis for hand.focicount -- good for seperating the points --- but it pools

#prediction for Dom strains: low variation 
Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)

Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)

Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)

Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)

#+theme(legend.position="none") 
```

```{r Q2.SC.long.biv_plots, echo=FALSE}

shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")

cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")

mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av. \nMale Long Biv SC Lengths")

cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")
mouse.shoo.long.biv
```

```{r Q2.SCshort_plot, echo=FALSE}

```

```{r Q2_tot.sc_plot, echo=FALSE}

#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

#MLH1_data <- MLH1_data %>%
#  arrange(strain, sex, mouse) %>%
#  mutate(file.name = factor(file.name))

mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")

mouse.av_totsc_male

```

```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
#there might not be enough data 
#
ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
```

<!-- basic stats -->

total SC for males (which ar)
X additional cells were added to the male specifi total SC data set discirbed above

-reduced bivalent dataset (long and short) is made up of bined bivalents from X cells and X mice

<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
remember the pairwise t.test
Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
G-WSB is sig, WSB-LEW are slightly sig

-->

```{r Q2.totSC.ttest, echo=FALSE}

#totalSC high and low rec group

#
Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,

                               (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
#p-value = 0.007

```

```{r Q2.short.long.ttest, echo=FALSE}

#long and short 
#mouse.av.short.biv.table
#Q2mouse.av.short.biv.table
Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)

#p-value = 0.9

#mouse.av.long.biv.table
Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
#p-value = 0.2

```

**we ran t.tests on the SC based metrics to test is the high and low groups are significanly different**

**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`). (the summary of mouse averages for this subset of ) **

this might indicate -there's error in the total SC algorithm

- mouse effects

- low power / sampling bias in reduced data set

- total SC is capturing differences in the XY signal which is excluded from the bivalent dataset

glms

```{r Q2.totSC.glm, echo=FALSE}

#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel

Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))

sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients

#sum.Q2.glm_mus.av_totSC_M1.coef <-

#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)

Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))

sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
```

(the subspecis effect is tested in M1)
For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)

strain effects from M1; G, LEW, MSM, SKIVE and TOM

strain effects  (basicall all strain effects are significant) (tech effects / error?)
All strain effects were significant when tested in M2 (Supplemental table X)

For M2 all

```{r Q2.glm.SC.long, echo=FALSE}

#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
#table(mouse.av.long.biv.table$mouse)
#Q2mouse.av.long.biv.table

glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)

sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)

glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)

sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)

#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
```

**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`). The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**

**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**

```{r Q2.glm.SC.short, echo=FALSE}
#Q2mouse.av.short.biv.table
#mouse.av.short.biv.table

glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)

#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)

glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)

```

In the short bivalent data set -- analyzed by M1, both subspeces are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`). And the strain effec for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).

For the M2 modeanalyzing the mouse means short bivalents, all strains have significant effect but LEW
G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.

```{r Q2.SC.long.glm.bivglm, echo=FALSE}

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)

#summary(glm_SC.male_Musc_long.biv)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)

## logistic regression for Rec group
#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
#              data=Long.biv_mouse_table,  family=binomial(link="logit"))

#summary(log.SC.male_long.SC)

log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))

```

**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.**

We purpose -- we think that running the models for comparing SC means by chromosome class and devided by rec group. (since the chromsome porportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )

The null hypothesis is that these won't be highly significant strain effects.
If there are still strain effects fo these models, that would be evidence that there has been eevolution for SC length compaction across these strains

`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images. X observations removed for having extra signal from other cells. `r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)

The skeletonized measures show that the general pattern holds for total SC and the high rec strains. However there are a few exceptions -- However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel

-the SKIVE are within dom and kaz level -- so why does SKIVE have inntermediate CO proportions and gwRR

24sep14_PWD_m1 is lower than the others
5jul17_MSM_m1 is the low MSM

The CZECH and AST are higher than KAZ

The number of cells and mice is difference across the low group strains, 
-- permut the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)

abova are glm tables for mouse.av SC lengths for each strain.

Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.

The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).

The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class. Mainly that 1CO are shorter in the high rec strains then low rec strains. This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up. Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.  This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group. They have a lower probability of 2COs below a certain SC length threshold.

<!--
blocks of other male Q2 - SC lengths

what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.

```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
#I don't think the Mixed models fit well for the high vs low polymorphism tests.

#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
#MSM_SC vs MOLF_SC

#2. logistic regression for sc length and chrm class?
#if the prediction / significance

#isolate data
#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]

mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)

#build and test MM
lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )

lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)

anova(lmer_MM_SC_length_male.poly)

rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm

```

-->

> **Summary**

When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression. I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length. (IE the scatter plots of SC length, binned by chrm class on the X).

Each point is a mouse average SC. When all bivalents are pooled, the logistic regression SC lengths predict the groups. When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.

 and the mean 2COs SC are higher in the high strains
(this must have something to do with the physical size effect)

If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)

Ideas for the above 3 tests,

1. no difference in DOM
- no sig logistic regression test
- glm, strain effect?

When all male mice are used, the predictive power is greater, than when just the Musc strains are used. When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)

> Is the prediction, high rec musc male strains have long SC met?

In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.  I couldn't get the Mixed models working for the male polymorphism predictions...

<!-- junk chunks below -->

```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}

log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2,  family=binomial(link="logit"))

summary(log.SC_male_chrm.class_all)

#use the 1CO or 2CO DF
Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")

log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))

summary(log.SC_male_chrm.class.WSB)

Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_G,  family=binomial(link="logit"))
summary(log.SC_chrm.class.G)

Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
summary(log.SC_chrm.class.LEW)

Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
summary(log.SC_chrm.class.PWD)

Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
summary(log.SC_chrm.class.SKIVE)

Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
summary(log.SC_chrm.class.KAZ)

Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
summary(log.SC_chrm.class.CZECH)

Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
summary(log.SC_chrm.class.MSM)

Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
summary(log.SC_chrm.class.MOLF)

```

```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}

log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
log.reg.chrm.class_all.strains

WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)

MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)

PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)

```

```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
#REMOVE COMPARES ACROSS FROM FULL BIVDAAT

#1CO
log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_1CO.poly)

log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_1CO.Musc)
#double check these directions

log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))

#summary(log.SC_male_1CO.Mol)

#2CO
log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_2CO.poly)

log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_2CO.Musc)

log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_2CO.Mol)

#3CO
log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_3CO.poly)

log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Musc)

#there's no rec group for this table
log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Mol)

#reverse logistic regression (single bivalent level)
#how well does SC length predict 1 or 2 COs?

Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)

#need to switch the co counts to 1 - 0
Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))

Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)

log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
              data=Bivalent_1o2,  family=binomial(link="logit"))

summary(log.SC_male_chrm.class)
#not sure if this is the equivilent of running each strain seperatly
#the strain effect is probably also pulling in the different proportion of chrm class
#useing the chrm.length*strain gives fewer significant factors.

#i'd be more convinced by running all those log regression plots for each strain
#come back later
```

These are the plots of the logistic regression.

```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}

plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)

plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)

plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)

```

the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group

```{r Q2.SClength.ttest, echo=FALSE, include=FALSE}
#COMPARISONS ACROSS ALL CLASSES

#MOUSE AVE, all
Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238

#mouse.av 1CO
Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993

#MOUSE.av 2CO
Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428

#single biv level
#male.biv?

#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])

#Rec.group_BivData_F1post_Ttest$p.value

```

```{r log.reg_SC.length, echo=FALSE, include=FALSE}
#REMOVE THIS COMPARES FROM FULL BIV DATA

#male_mouse.av
#Male.poly.Mouse.Table_BivData_4MM

Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))

#male poly shouldn't be here
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))

#All chrm
log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av)

#considering seperating out by hand foci counts

#1CO
log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_1CO)

#2CO
log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))

#summary(log.reg_SC.length_mouse.av_2CO)

#3CO
log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_3CO)

# bivalent level log regression

log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

#summary(log.reg_SC.length_BIV)

#1CO  (BIVDATA)
log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,

                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_1CO)

#2CO
log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_2CO)

#3CO
log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))

#summary(log.reg_SC.length_BIV_3CO)
```

```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
#REMOVE THIS, IT uses total bivalent data

#rerun the models for datasets -- deviding by chrm class and rec group
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )

Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )

#calq_mouse av
mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                        n_cells = length(unique(fileName) ),
                        mean.SC.total = mean(chromosomeLength),

                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

                         )

mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)

mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)

#calq_mouse av
mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)

mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
```

```{r Q2.SC.Length.stuff, echo=FALSE, include=FALSE}
#REMOVE THIS, COMPARES FULL BIVDATA ACROSS CHRM CLASS

#all pooled SC lengths
#WSB     G   LEW  MOLF   KAZ CZECH 
#    5     9     5     4     3     3

#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14

#G sample 5 mice

#break down of mice in the low rec dataset

#glm.1CO
Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)

#glm.2CO
Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
#fewer significant differences

Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
#all but KAZ are slightly significant, G and CZECH more so

#glm.3CO
Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
#not sig differences
summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)

#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)

#all pooled SC lengths
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value

#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)
Q1lmer.MM_SC_short_results
```

```{r Q2.SC.ploy.male, echo=FALSE, include=FALSE}
#REMOVE THIS< COMARES FROM FULL BIVDATA set

# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect

#1. determine which strain effects are sig
#use mouse biv average
#Male.poly.Mouse.Table_BivData_4MM

glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm1.SC.male.poly) 
#SC many strains are different (more than the gwRR)
#subsps are sig

glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm2.SC.male.poly)

#Dom
glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(glm_SC.male_Dom)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(glm_SC.male_Musc)

#Mol
glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(glm_SC.male_Mol)

#Logistic regressions

log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))

                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],

#summary(log.SC.male.poly)

log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))

#summary(log.SC.male.poly_musc)
#not sig -- when just looking at Musc

#significant in predicting group
log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))

#summary(log.SC.male.poly_mol)
#not sig -- when just looking at Mol
```

```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)

fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)

table( droplevels(fomo.male_mouse.av$strain) )
#there are cast strains too

Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
musc.male.plot

# + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +
#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +

#make the Mol and
Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")

Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
Mol.male.plot

#Dom version 
Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")

Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
Dom.male.plot
```

```{r SC.length.male_tests, echo=FALSE, include=FALSE}

## run the tests to check following expectations

#1.Dom strains not different from ech other
#2. PWD and SKIVE are different from the other strains
#3.MSM and MOLF are different from each other

#Male.poly.Mouse.Table_BivData_4MM
#Long.biv_mouse_table_male.poly

#decide between lm or glm..?
#mouse avs, males

#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]

male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
#the subsp labels get messed up, unless they are designated chr first
Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )

Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
#total sc sig, 1CO sig, 2CO sig

Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
#0.000369 ***
Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
#0.000816 ***
Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
#0.0148 *

show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()

show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
#making plots above show that mouse average SC lengths are lower could there be microscope issue?

#make sure to remove some of the juvi mice!! from the AP_mouse table
show.Dom.male.CO <- ggplot(
male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()

Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
#P not sig for CO mean

#seems like there are more mice in the MLH1 data set than bivData
#

#WSB all dissected and imaged in '15
#G imaged 15 - 16
#LEW imaged 15 to 17

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "", xlab = "")

show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "", xlab = "")

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "1CO SC Length", xlab = "")

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "2CO SC Length", xlab = "")
```

```{r SC.length.pairwise.T, echo=FALSE, include=FALSE}

#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av

male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]

Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
#G-WSB is sig, WSB-LEW are slightly sig

Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
#PWD and MSM are most diff from SKIVE?

Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
#the pooled SC value tests are sig

#the mol comparisons are different due to adj p values

Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
#G-WSB is sig, WSB-LEW are slightly sig

Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
#no sig (only 3) -- does include CZECCH

#non-sig, with mol pooled with musc

Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)

Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)

Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
#non of these sig with the mulitple adjust

Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)

#pairwise.t.test(Ozone, Month, p.adj = "bonf")

```

# Q2 Normalized CO positions

In order to test /

We tested another aspect of the rec landscape -- focuing on differences between the high and the low rec strains.

To detect if there are significant differences between the high and the low rec strains we performed t.tests across the two groups

(and logistic regressions?)

<!-- set up chunks here

-->

```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

##long biv
male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
#table( droplevels( male.long.biv$strain)  )

mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,

                        n_obs = length(Obj.ID),
                         mean_F1 = mean(Foci1, na.rm=TRUE),
                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),

                        mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)

                                               )
#add subsp

mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)

mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)

#add rec group
mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
```

<!-- plots /figures below-->

```{r Q2.show.nrmF1, echo=FALSE}
#Curated_BivData_male

#boxplots of nrm F1
male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)

Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")

Q2.show.nrmF1

nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")

nrmF1_mouse_level
```

Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.

```{r Q2.F1.pos.LONG.BIV, echo=FALSE}

#make the same plots as above but with the long biv data set
male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)

#mouse.av.long.biv.table

Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")

Q2.show.nrmF1_long.biv

Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Nbiv  = length(Obj.ID),
                       mean.1CO_PER = mean(PER_Foci_1)
                      # mean_CO = mean()
                       )

Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)

Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")

Q2.shoooo
```

A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.

or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.

<!-- 
Main DFs for this section:
Male.poly.Mouse.Table_BivData_4MM
male.bivdata.1CO
-->

```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}

#mouse.avs_4MM
#31may19_SKIVE_m2  has NA for norm.first.foci.pos

#High = MSM, PWD, SKIVE
Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
#p-value = 1

#high = PWD and MSM
Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
#p-value = 1

#single biv level
Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])

#Rec.group_BivData_F1post_Ttest$p.value 
#p-value = 0.238

```

The t.tests are not significant for normalized foci position comparing all bivalets

The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).

```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
#these don't have mean.first.foci

#short
Q2.1COpos_short_ttest_HL <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
#p-value = 0.06

#remove SKIVE
Q2.1COpos_short_ttest_HL.ex <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
            (Q2mouse.av.short.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
#p-value = 0.4

#long
Q2mouse.av.long.biv.table_HL  <-  t.test( (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (Q2mouse.av.long.biv.table %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
#p-value = 0.5

#long exclude skive  Q2mouse.av.long.biv.table
Q2mouse.av.long.biv.table_HL.ex  <-  t.test( (Q2mouse.av.long.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
            (Q2mouse.av.long.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
#p-value = 0.08

#they kinda have ooposite effects

```

The t-tests for the high and low rec strains for the reduced bivalent datasets are mostly NS

mouse mean
short bivalent

So try using the bivalent sample levels

```{r Q2.1COpos_BIV_ttest, echo=FALSE}

whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group == "1"))$PER_Foci_1   )
#p-value = 0.2

whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )

#p-value = 0.4

```

The t.test at the bivalent level is also NS (test; p = )

```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}

#compare within subsp

#Musc
Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )

# p-value = 0.4
# exclusive rec group not sig
```

**Most all of the t.tests I've done indecate that there are no sig diff between the High and the low rec groups**

The normalized 1CO position is not a good metric to distinguish the high and low rec groups.

GLMs

glms are constructed to test if there has been evolution between strains ... ect.
I will use the full data set the reduced datasets have such low power

```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}

#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos

Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))

sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients

#M2
Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))

sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients

```

for M1
Mol subsp is significant, strain G (LEW slight sig), strain MSM -- strain MOLF is NA

for M2,

G, LEW, PWD, MSM, MOLF, SKIVE, KAZ have significant effects

put together, these glm's indeceate that MOLF is the most different in it's 1CO landscape (compared to other males). Is this a sampling issue

For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice

there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
(motivation for MOLF female observations and more male MSM biv measures)

```{r Q2.1CO.landscape, echo=FALSE}
#Male.poly.Mouse.Table_BivData_4MM
#Curated_BivData_male

#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm1.SC.male.poly) 
#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm2.SC.male.poly)

#M1
glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm.M1_Q2.1COnrm)

#M2
glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm.M2_Q2.1COnrm)

#Dom
glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(glmQ2_M2.1COnrm_Dom)

#Musc
glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(glmQ2_M2.1COnrm_Musc)

#Mol
glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(glmQ2_M2.1COnrm_Mol)

#m3 (log)
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))

Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm)

#Musc
Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
              family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm_Musc)

#Mol
Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
              family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm_Mol)

#biv level logistic regression
#Curated_BivData_male
#male.bivdata.1CO

male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))

Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO,  family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm)

#Musc
Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
              family=binomial(link="logit"))
#summary(Log.Q2.F1_PER_Musc)

#Mol
Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
              family=binomial(link="logit"))

#summary(Log.glmQ2_M2.1COnrm_Mol)

```

## JUNK Section

# Q2 IFD

<!-- set up chunks here-->

B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.

1. glm across subsp

2. anova within subsp

3. logistic regression for musc (or all males)

```{r Q2.show.IFD, echo=FALSE}

#subset
male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)

male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))

biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))

biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
biv.IFDs_PER_by.group
```

The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
 The normalized plot displays the pattern of higher rec group having slightly higher longer IFDs.

```{r Q2.IFD.glm.test, echo=FALSE}
#glm --
#mouse aver

#M1.ABS
glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
summary(glm.M1.Q2_IFD.raw) 
#M1.PER
glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
summary(glm.M1.Q2_IFD.PER)

#M2.raw
glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
summary(glm.M2_Q2.1COnrm)

#Dom
Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
summary(Q2.glm.M2_IFD_Dom)

#Musc
Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
summary(Q2.glm.M2_IFD_Musc)

#Mol
Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
summary(Q2.glm.M2_IFD_Mol)

#M2.PER
glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
summary(glm.M2_Q2.IFD.PER)

#Dom
Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
summary(Q2.glm.M2_IFD.PER_Dom)

#Musc
Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
summary(Q2.glm.M2_IFD.PER_Musc)

#Mol
Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
summary(Q2.glm.M2_IFD.PER_Mol)
```

```{r Q2IFD_logreg, echo=FALSE}

Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))

#IFD.raw
log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.raw)

#MUSC
log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.raw.Musc)

#Mol
log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.raw.Mol)

#SINGLE BIV LEVELs
male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)

male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))

Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
#summary(Log.Q2.IFD.raw.PER_bivdata)

#Musc
Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))

#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)

#Mol
Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))

#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
```

```{r Q2.IFD.PER, echo=FALSE}

#IFD.PER
log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
summary(log.reg.Q2.IFD.PER)

#MUSC
log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
summary(log.reg.Q2.IFD.PER.Musc)

#Mol
log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
summary(log.reg.Q2.IFD.PER.Mol)

#SINGLE BIVALENT LEVEL

Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
summary(Log.Q2.IFD.per_bivdata)

#Musc
Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))

summary(Log.Q2.IFD.per_bivdata.Musc)

#Mol
Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))

summary(Log.Q2.IFD.per_bivdata.Mol)

```

```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}

#t.test for IFD
#mouse.av - the mouse average IFD is from just 2CO
t.test_mouse.av.IFD_ABS <- t.test(
  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

t.test_mouse.av.IFD_PER <- t.test(

  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

```

```{r Q2.t.test_IFD,echo=FALSE}

#already made
#subset
male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
#t.tests for the Rec group
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))

#Curated_BivData <- add_IFD(Curated_BivData)
#THIS SHOULD ALREADY BE ADDED

#single biv level
Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])

Rec.group.BivData_IFD.PER_Ttest$p.value
#p-value < 2.2e-16

#MOUSE AVE
Rec.group_IFD.PER_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group_IFD.PER_Ttest$p.value 
#p-value = 6.872e-07

```

The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).

### Q2 Long Bivalent Dataset

Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern. In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5). The section on IFD only looks at the 2CO bivalents.

```{r Q2.long_biv.IFD, echo=FALSE}
#long BivData
male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)

male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))

#show
Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))

Long.biv.IFDs_PER_by.group

```

table(droplevels(male.long.biv.2CO_IFD$strain) )

The plot above shows the normalized IFD measures from 2CO bivalents

There are many fewer long bivalent data observations, but the same positive correlation with

-table of long bivalents

- glms

- logistic regression

```{r Q2.IFD.longbiv_glm, echo=FALSE}

#Full dataset

male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)

#single bivalent
Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
#summary(Q2.glm_long.biv_M2.IFD.PER)
#The PER.IFDs don't have significant strain effects

```

There are not enough observations for sub setting the data into strains or use the mouse averages.

Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
The above t.tests are breaking the knitr

The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.

None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.

Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent. This also indicated that interference distance may evolved in the house mouse complex.

#Caveats

put all of the code chunks/analysis for caveats here

## Chromosome Size Effect

I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
--> 
Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plots excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correctly mapped)

```{r SC.distb_manual_Auto, echo=FALSE}

#calq the mouse means
manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,

                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
                        median_SC = median(SC.length, na.rm = TRUE),
                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)

)

manual_summary_stats <- add_mouse(manual_summary_stats)
manual_summary_stats <- add_strain(manual_summary_stats)
manual_summary_stats <- add_sex(manual_summary_stats)
manual_summary_stats <- add_subsp(manual_summary_stats)

###
#calq the mouse means
Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        median_SC = median(chromosomeLength),
                        Qar.4 = quantile(chromosomeLength)[[4]],

                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
        #add Q4

)
Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)

#colors_of_strains
#scale_color_manual(values=colors_of_strains)

#un-order these factors
```

```{r long.biv.plots, echo=FALSE}

show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  scale_color_manual(values=colors_of_strains)+
  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")

show.the.cells_females_dom
#PWD seems to have the biggest variance across cells

almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
  scale_color_manual(values=colors_of_strains)+
  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
almost_cells_DOM_f

#### Musc female

show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")

almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+

  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")

### Dom Male
show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")

almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+

  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")

#MUSC male
show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")

almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+

  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")

#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+

#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+

#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

 # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
#show.the.cells_male

#almost_whole_cell_BivData
#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]

#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
```

All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data. It shows the amount of within cell variance across strains. There is a bit of variance across the SC length distributions in the PWD females.

This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).

The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. The full data set is `r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category for this long dataset are

**- Try to merge this DF with the whole.cell manual measures.**

- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.

in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
10202 bivalents from 86 mice.

```{r show.long.SC, echo=FALSE}

female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
female_long_biv_plot

male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
male_long_biv_plot

```

These plots show the SC lengths for the 'long SC data set'. They are supposed to be the longest 4-5 SC from cells where I could get good measures. These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position). Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).

## Adjusting for XX

(rethink where this section should go)
new outline

1. illustrate problem(affects mostly SC length)

2. Expected impact on sex comparisons, estimated effect size of the X  
3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons

4. 19 female, 19male, 20female 20 male

The female mouse averages should have adjustments for the XX. working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains. Subtract this amount from the female mouse averages ... This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)

- Of all female single bivalents observations, 5% are XX (1 of 20).

- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).

- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.

I think a formula something like this can be applied to adjust for XX

- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*

#### Whole Cell Manual Comparison

```{r XX.adj_manual.whole.cell, echo=FALSE}

#load and examine the the whole cell hand measures here

female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
data.frame(table(droplevels(female.whole.cell$category)))

# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv

#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC

data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
#missing some strains, most have 3 cells (min). MSM has 5 full cells

#remove the incomplete LEW female cell
female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]

female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]

#calq mean SC lengths by rank

mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,

                       mean.sc =   mean(SC.length, na.rm = TRUE),
                         nbiv = length(Obj.ID),
                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
                         se.SC_length   = sd.SC_length / sqrt(nbiv)
)
         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =

top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]

show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
                    position=position_dodge(width=0.7), 
                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")

show.SC_by.rank_top

bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]

show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
                    position=position_dodge(width=0.7), 
                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")

#CALQ PERCENT TOTAL SC

CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,

                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
                         nbiv = length(Obj.ID),
                       #rank.by.per

                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
)

#MELT ABOVE
MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
              #leave out the Foci
               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],

            id=c( "fileName","total.sc.cell","nbiv"))

MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]

MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)

show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")

show_long.sc.proportions

#original plot code 
#ggplot(data = df1, aes(x, y, color = Group)) +
#    geom_pointrange(aes(ymin = lb, ymax = ub), 
#                    position=position_jitter(width=0.5), 
#                    linetype='dotted') +
#    theme_bw()
```

The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.

The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).

The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank. For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).

(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).

The other figure shows of each single bivalent contributes to the total SC area. Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell. on average, each of the top longest bivalents make up ~10% of the cell's total SC area. So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,

- Is the difference between cell averages for males and females less that 10%?

- also interesting, the pwd and msm don't have longer SC, compared to other strains.

#### Automated BivData Comparison

```{r XX.Automated.biv.stuff, echo=FALSE}

#make automated female data to compare to
female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]

female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]

table(female.Bivdata_comp$strain)

#female.Bivdata_comp

#1. calq the rate of chrm segmentation in female categories in automateted data

box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,

                   Nmice = length(unique(mouse)),
                   mx.box.num = max(boxNumber), # need to run this at the cell level
                    N.biv  = length(Obj.ID),
                   pass.biv = N.biv / mx.box.num
)
box.rates <- add_mouse(box.rates)
box.rates <- add_strain(box.rates)
box.rates <- add_category(box.rates)

cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

cell.pass.rates

category.box.rates <- ddply(box.rates, c("category"), summarise,

                   mean.pass.rate = mean(pass.biv)
)
#category.box.rates

#2. pull the 3rd rank biv's calculate the SC means
#calq means by rank (focus on the top 5)

```

For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.

The table above shows the number of bivalents from the same strains as in the manual whole cell data. The plot shows the bivalent passing rate across all of the individual cells from this female data set. For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).

(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level. (it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)

strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*

** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX. But I am not using 'whole cell' summaries to compare female and males.

**What is the effect of an extra XX-autosome on single bivalent means?**

use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents. Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females. Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly. Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.  The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.

# Note on Heterochiasmy Definition

<!-- add outline for new XX adjust section 
adjust the colors for the plot below (standardize)
-->

I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison. In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape. Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count). (I note if the and Y are paired, which they are at a high rate). The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y). To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.

1. Compile full-cell data from females (all 20 bivalents measured)
2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
3. check how variable the REC is across the 1st,2nd,4th, and 5th are.

According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).

(Put the XX adjustment section here)

```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells

#outgroups
#SPIC
HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
#unique(HetC.SPIC$mouse)
#non-passing mice
SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
#calq sex mean
SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]

#SPRET
HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
#unique(HetC.SPRET$mouse)
#non-passing mice
#SPRET.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]

#CAROLI
HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
#unique(HetC.CAROLI$mouse)
#non-passing mice
#SPRET.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

#cast
HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
#unique(HetC.CAST$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]

#HMI
HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
#unique(HetC.HMI$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

#dom

#WSB
HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
#unique(HetC.WSB$mouse)
#non-passing mice
WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
#calq sex mean
WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]

#LEW
HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
unique(HetC.LEW$mouse)
#non-passing mice
LEW.remove.mice <- c("4jan17_LEW_f2")
HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]

#calq sex mean
LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]

#G
HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
#unique(HetC.G$mouse)
#non-passing mice
G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]

#calq sex mean
G.mean.MLH1.obs <- ddply(.data=HetC.G, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]

#PERC
HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
#unique(HetC.PERC$mouse)
#non-passing mice
#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]

#calq sex mean
PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]

#MUSC

#PWD
HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
#unique(HetC.PWD$mouse)
#non-passing mice
PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]

# CZECH
HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
#unique(HetC.CZECH$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]

# SKIVE
HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
#unique(HetC.SKIVE$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]

# KAZ
HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
#unique(HetC.KAZ$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]

#TOM
HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
#unique(HetC.TOM$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]

#AST
HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
#unique(HetC.AST$mouse)

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]

#Molosinus

#MSM
HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
unique(HetC.MSM$mouse)

#non-passing mice
MSM.remove.mice <- c("12sep16_MSM_f3")
HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]

#calq sex mean
MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]

#MOLF
HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
unique(HetC.MOLF$mouse)

#non-passing mice
MOLF.remove.mice <- c("27oct18_MOLF_m3")
HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]

#calq sex mean
MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]

```

```{r HetC.plot, echo=FALSE, warning=FALSE}
#HETC plot

#set up data

HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#
#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]

HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]

#remove the 'other'

HetC_strain_table.M$sex <- "male"
HetC_strain_table.F$sex <- "female"

HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")

HetC_strain_table <- add_subsp(HetC_strain_table)

#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")

HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
#str(HetC_strain_table)

Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )

Het.C.plot
#this plot is missing several

```

There is now MOLF, which has female biased hetC
3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
1 of the musc strains has female biased heterochiasmy, KAZ.

```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}

#REMOVE This BLOCK?

#### Distributions by mouse
#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.

#remove this block since it has it's own src/script

### to integrate 'pass'/fail
### this should be in setup data

#mice that don't have female data will break the chain-loop
#MOLF female, HMI, CZECH female, CAROLI female, Tom male

#instead of adding fake data, remove the categories from list

#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing

pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")

#this loop makes a quant_status col for a list of mice
for(i in 1:length(MLH1_data$mouse)){
  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
}

df_list <- split(MLH1_data, (MLH1_data$category))
#remove sections which are in category but have no data
#this is where you remove the non-data data!
df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]

#7 PERC, 12 MOLF female, 
#

#df list, is a list of df's which were split() from BIG df by category
df_list 
o=0
i=0

MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)

#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
for(i in 1:(length(df_list))){
 # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
                      mouse_nMLH1 = mean(adj_nMLH1.foci)
  )
# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
  for(o in 1:length(df_list[[i]]$Original.Name)){
    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
        }
}

plot_list = list()
for (h in 1:length(df_list)) {
  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
    facet_wrap(~ mouse, drop = TRUE) + 
    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
  #make an empty ggplot? for missing data

  plot_list[[h]] =   pf
}

#orig (I think this is the same as above)
#plot_list = list()
#for (h in 1:length(df_list)) {
#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)

#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +

 #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +

    #theme for indiviual mice, 
#theme(plot.background = element_rect(fill = "green"))
#  plot_list[[h]] =   pf
#}
#can't change the alpha/transparency of the red rec

#aes must be same length ---   #geom_point(stat = "identity") +
# the last one is correct....?

#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
```

```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
geom_hline(aes(yintercept = med, group = gr), colour = 'red')
#no plot is made that goes in the list -- so it can't be shown
for (gg in 1:length(plot_list) ){
  show(plot_list[gg])
 # invisible(lapply(obj, function(x) plot(x,main="some plot")))
}

#this stops at MSM
```

The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.

Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse. The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.

```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}

no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")

MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]

MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]

MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]

MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]

kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)

bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)

#add titles
#use multiplot
#remove X axis

multiplot(kuku,bubu, cols=2)
```

```{r F1.by.F2, eval=FALSE}

#try remaking the plot Megan suggested
# for 2CO positions, Foci1, Position  on x and Foci 2 position on y

CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]

CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]

#isolate 2COs
#facet by sex and subsp

F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
F1.x.F2

```

# Deleted notes

```{r variance.hist, echo=FALSE, eval=FALSE}
#figures of histograms,

#make series for total dataset?
Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +
  ggtitle("Histograms of within mouse variance, all mice")
Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
Var.Hist.total.mice

#DF.HetC.MixedModel.HQ
Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  + 
  ggtitle("Histograms of within mouse variance, 10 cell min")
Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")

Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")

Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")

Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
Q12.var.hist.plot

#some LEW female still have v high within mouse variance (2 mice, )

```

# References