### 2020-03-14 doc/ThesisIntrOutline.rmd
notes for outline draft

---
title: "Intro Outline"
output: html_notebook
---

# themes across chapters

## 1. recombination rate variation in wild population

natural variation for recombiantion rate,

how Rec evolves, parasites, stress het Environments

genetic variation and environmental effects,

## 2. Heterochiasmy, sexspecific evolution of recombination rates

## 3. between cell variation in recombination (meiotic features)

## Figure / Image ideas

- chrom axis / DNA loop

## aim of the thesis

the aim of this thesis --- is to investigate variation and evolution of genome with recombination rates

within species variation (natural populaiton) and sex

genome wide rates -- broken down (to investigate the chromosome - level landscape / broad scale patterns / meiotic features)