### 2020-04-21 doc/MatMeth.Rmd
changing the Rdata files across scripts

---
title: ''
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

#Material Methods Draft v1

```{r set.up, echo=FALSE, include=FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)
#library(patchwork)

#set wd
#load datafile
load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

#common bugs
#html won't compile -- one of my citations in bib was breaking things
```

```{r data.sets, echo=FALSE}

#Q1.real.short_mouse_table

#

```

### Mouse Husbandry

To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used. The _Mus musculus_ strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA). The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).  (https://en.brc.riken.jp) and rederived by the UW-Biotech centre (Madison UW).
The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken. We use abriviated superscript notation for the remainder of the paper (e.g. _musculus^WSB^_). All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.  A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (G) is maintained at UW Veterinary school facilities.

<Mice were fed on dry standard breeder chow. Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival. Adult mice were euthanized by CO asphyxiation. Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison. Over the course of data collection some breeding colonies  mice were moved from facilities. Additionally the GI strain was kept at a separate facility.>

### Tissue Collection and Immunohistochemistry

The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes. Spermatocyte spreads were collected and prepared as described in [@peterson2019]. The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification. Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.

Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively. Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution. Cells were fixed in this solution and dried in a humid chamber at room temperature overnight. The following morning, slides were treated with a photoflow wash (Kodak, cite). 
**slides were stored at -20*C if not stained imeediantly**

<Because some strains have breeding issues, all mice some strains were kept in breeding pairs (before collecting, and aged out of the desired age range. We performed tests for effects on MLH1 counts and found no effects in the older male mice ages. (Also we tested other effects, including maternal age...)>

The **immunohistochemistry**

 To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of of the SC (SYCP3). Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively. The staining protocol was based on that in [@anderson1999] and [@koehler2002]. Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).

Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C. The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB. Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees. Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200)  were suspended in ADB. The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees. Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.

### Image Processing

```{r Image.profile, echo=FALSE}

# microscope
# cell certeria

```

Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK). Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0). Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.

### Analysis

We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).

### DMC1

```{r DMC1.section.outline, echo=FALSE, include=FALSE}
#cell certeria, number, t.test
#DMC1 -- leptotene and sygotene cells were identified by SC morphology (by degree of synapsis) (cite another paper?). foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
```

Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.

SC morphology and CREST/centromere number was used to stage spermatocytes **Certeria for early and late zygotene** - (bright staining - )
early zygotene defined as seperate centromere pairing -- mostly complete axis signal. late zygotene judged as -- low degree of centromere pairing -- and more homologous pairing (but not fully synapsis). Differences in DMC1 counts per cell were tested by ttests.
<numbers of cell?>

### MLH1 (Analytical Framework)

```{r MLH1.analysis, echo=FALSE}

#1.certeria
#2. model framework 
#2A (logic of fixed and random effects) -- 
#  2b. packages used

```

**certeria**  MLH1 foci were quantified from cells with intact and compelete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci. A quality score, between 1 to 5, was asigned to each cell based on staining quality and general spread of bivalents. <cell with small defects were included -- and coded as a score of 5 - but these cell weren't included in the final analysis.>  The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).

**the XY foci is not counted / analyzed in this study**

-Background for setting up/organizing models -- (heiarchal nature of multiple strains within subspecies -- ) - hierarchical nature of categories for inviual mouse averages ()

-(primary focus is mean MLH1 count per cell - at the mouse observation level - apply to more meiocyte traits)

**partioning variation in recombination rate**

-Our goal -- design an utilize a mixed model for our analytical framework - which leverages / accounts for the hierarchical/nested nature of categories across all the individual mice (e.g. sex, strain subspecies).

<The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).>

The use of wild derived inbred strains enabled us to design an utilize a mixed model for our analytical framework. We built our model to predict the strain average MLH1 count per cell.

**X package used to test the fixed effects, Z package used to test the random effects.**

The main effects and their interaction of subspecies and sex were coded as fixed.

strain (or genetic background) was coded as a **random** effect to reflect a random sample of natural genetic variation from each subspecies.

In addition to mouse mean of MLH1 count per cell
**to expand the analysis of meiotic features** We applied the following models to several dependant variables (mouse averages) <consider a figure/diagram>

 the mouse average of variance in crossover count per cell,

 mouse mean total SC per cell, mouse mean single SC length, mouse mean IFD^norm^ and IFD^raw^, and position of crossover for single crossover bivalents.

**M1. Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$

**M2. Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$

**M3. Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$

**M4. Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$

This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect. We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner). From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.

We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].

We used x package for testing significance of fixed effects and V package for testing significance of (strain) random variables (strain).

### Single bivalent charaecterization

**combine any X and XY issue with reduced dataset** An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).

```{r single.biv.chr, echo=FALSE, include=FALSE}

# image analysis programs
# Richard total sc

# DNA.CrossOver (site, efficiency, curation, chromosome size effects-reduced dataset,

# chromosome proportions, number of CO per chrm, proportions perpared by chi-sq tests

#chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
#these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups

# sc bivalent used in models (aldready above), the length of single SC compared by t-tests -- and fit into models
# Foci Position -- t-test  and put into models
# IFD / interference
```

To charaecterize the SC length of individual chromosomes which is a marker of chromatin compaction <from MLH1 stained cells>, two image analysis algorithms were used to specifically (measure indiviual cells) [@wang_2019_sc] and indiviaul chromosomes [@peterson2019]. Both apply a 'skeletonizing' transformation to the synapsed chromosomes which transforms the chromosome space into a single pixel wide 'trace' of the originial chromosom shape. Hand measures of SC/pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].

The **total SC** per cell (all bivalents) was quantified across single cell images. Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X). the mouse average total sc was applied to the model framework.  <how else was this analyzed?!>

**walk through how outliers are removed**

##### DNACrossover

The **DNA CrossOver** algorithm isolates single straightened bivalents/chromosomes shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively. This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents. In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).

A curation step is applied to the total single bivalent dataset to remove poor isolation/segmentation instances. From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software. After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.

#main traits

**This dataset is used to charaecterize single bivalent traits**

The number of MLH1 foci on single chromosomes was quantified from the curated bivalent dataset to better understand the total recombination landscape. The **proportions** of single and double crossover bivalents were tested across mouse categories by chi-sqaure tests with the prop.test function.

**other traits** were also charaecterized -- SC length (a marker for chromatin compaction), foci position (raw and normalized), and interfocal distance a marker for interference, these traits were fit into the main models (as mentioned above).
In addition to fiting the mouse means into the models discribed above, t.tests were used to test the means between sexes and groups.

**IFD, Interference** In order to examine interference, Interfocal distance was measured from the single bivalent data sets. - This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers. Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).  **IFD^norm^**

**SC Length model**

- **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$

- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$

**REDUCED DATASET**

To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell. This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes. A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes. Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.

- The long bivalent and short bivalent data sets were compiled by calculating the 1st and 4th quartille of SC lengths for each cell with at least 16 bivalent measures. Bivalents from each cell were assigned into the long and short bins based on these SC Length cut offs.

Using the single bivalent dataset, the within cell distribution of bivalent SC lengths from cells with at least 17 isolated chromosome measures. The longest and shortest 4 to 5 chromosomes within each cell were identified by calculating the fourth and second quartille.

# References