### 2020-05-14 src/SetupData.Rmd

---
title: "Setting up Data"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(xlsx)
library(reshape2)
#knitr::opts_chunk$set(echo = FALSE)

opts_knit$set(root.dir = "~./MLH1repo/")
setwd("~./MLH1repo/")

meta.data = read.csv( "~./MLH1repo/data/clean.Meta.Data.txt", sep = "\t", header=TRUE)
#remove all MAD (peromyscus) from meta.data
#meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]

source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )

#MLH1_data = read.csv(here("data", "/AnonData.csv"), sep="\t", header=TRUE)
#read.csv was breaking things
MLH1_data = read.xlsx2("~./MLH1repo/data/AnonData_formated.xlsx", header=TRUE, 1)

DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)

#total SC -- the one file should be all I need?
Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)

#Curated Bivdata - this excel file is too big to work with read.xlsx2
#Curated_BivData = read.xlsx2("~./MLH1repo/data/Curated_BivData.xlsx", header=TRUE, 1)

Curated_BivData <-  read.csv("~./MLH1repo/data/Curated_BivData.csv", header=TRUE)

```

```{r MLH1.DF.setup, warning=FALSE, echo=FALSE}

original_DF = MLH1_data

#isolate the duplicates (fileName) from main MLH1 DFs
n_occur <- data.frame(table(original_DF$fileName))
#n_occur[n_occur$Freq > 1,] displays the dups
original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,

#use the above DF for quantifying repeatability in the Main_report.Rmd
#duplicates are cells counted twice

#are there duplicated filenames+batch
original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")

anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch and batch number --

#remove duplicates from original as well
original_DF <- original_DF %>% distinct(Original.Name, .keep_all = TRUE)#5216 to 4983

#calculate the correlation across duplicates. First issue is to match up and organize the matching 
#for each fileName - make cols, for MLH1-1 and MLH1-2
#remove Xs, for fileName with 
original_DF_duplicates <- original_DF_duplicates[!grepl("X", original_DF_duplicates$X),]
#make a new DF -- with one fileName and the MLH1-1 and MLH1-2
#  (duplicate / dostinct fileName )

#keep just 1 of duplicates from MLH1_data
#MLH1_data <- MLH1_data[duplicated(MLH1_data$fileName),]
MLH1_data <- MLH1_data %>% distinct(fileName, .keep_all = TRUE)

```

`r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.

```{r clean_MLH1_data, echo=FALSE}
#Clean up main DF, MLH1_data
#remove duplicated data from MLH1 -- remove 1 or the two rows
#https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/
#use distinct() in dlpyr()
MLH1_data <- MLH1_data %>% distinct(Original.Name, .keep_all = TRUE)#5216 to 4983

MLH1_data$Original.Name <- as.character(MLH1_data$Original.Name)
MLH1_data$fileName <- (MLH1_data$Original.Name)

#these introduce NAs
MLH1_data$n <- as.character(MLH1_data$n)
MLH1_data$nMLH1.foci<- as.character(MLH1_data$nMLH1.foci)
MLH1_data$quality<- as.character(MLH1_data$quality)

MLH1_data$n <- as.numeric(MLH1_data$n)
MLH1_data$nMLH1.foci<- as.numeric(MLH1_data$nMLH1.foci)
MLH1_data$quality<- as.numeric(MLH1_data$quality)

#remove X's and NAs
MLH1_data <- MLH1_data[ !grepl("X", MLH1_data$X) , ]
MLH1_data <- MLH1_data[!(is.na(MLH1_data$nMLH1.foci) | MLH1_data$nMLH1.foci==""), ]
#this should also have the empty column for removing X
MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "X",]
MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]

```

```{r misc.setup, echo=FALSE}

#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4',

                       'TOM'= 'indianred1','AST' = 'tan2',
                       'CZECH'= 'brown1',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
#4!!

```

```{r exclude.mice, echo=FALSE}
#remove mice -- 18nov17_WSB_f5 (only 1 cell)
exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                 "3sep18_AST_m1",
                  "3sep18_TOM_m1",
                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]

#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
#juvi.mice_DF <- juvi.mice_DF[!(is.na(juvi.mice_DF$mouse) | juvi.mice_DF$mouse==""), ]

#juvi mice ~ 2 weeks old
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )

#make juvi.mice DF
juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice),]

#remove juvi mice from big MLH1 -- but Then i need to remake the mouse_table
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice, ]

# > 25 weeks
# I might need to keep some -- espcially outgroups
#old mice -- This list includes mice which don't have MLH1 observations, but I have records of dissecting them

old.mice <- c(
  "12oct15_WSB_m1","9oct17_WSB_m1","1mar18_WSB_m2","21dec14_WSB_m2","21dec14_WSB_m3", "1mar18_WSB_m1",
  "14mar15_G_m1","14mar15_G_m2","24mar15_G_m1","19oct15_G_m1",
  "8may17_LEW_m1","18sep17_LEW_m1","8may17_LEW_m2","20dec16_LEW_m2",
  "5jul17_PERC_m1",

  "27oct18_SKIVE_m2","4feb19_SKIVE_m1",
  "1feb18_KAZ_m1","11jan18_KAZ_m1","19may18_KAZ_m2","9apr18_KAZ_m1","9apr18_KAZ_m2",
  "20dec18_CZECH_m1","3dec18_CZECH_m1",
  "28jun18_TOM_m1","3jul18_TOM_m1",
  "27aug18_AST_m1","7aug18_AST_m1",

  "13nov16_MSM_m1","13nov16_MSM_m2","4feb19_MSM_m3","4feb19_MSM_m4","9oct17_MSM_m2","4feb19_MSM_m2","30may17_MSM_m1","5jul17_MSM_m1",
  "4feb19_MOLF_m1",

  "24aug17_CAST_m1","23sep17_CAST_m1","23sep17_CAST_m2","30jun16_CAST_m3","5jul17_CAST_m1",
  "27aug18_HMI_m2","28jun18_HMI_m1","27aug18_HMI_m1","4dec17_HMI_m1",

  "11jan18_SPRET_m1","27aug18_SPRET_m1","7aug18_SPRET_m2","7aug18_SPRET_m1","13jul17_SPRET_m1","2aug18_SPRET_m1",
  "19may18_SPIC_m1","19may18_SPIC_m2","30oct17_SPIC_m1","11jan18_SPIC_m1",
  "1mar18_CAROLI_m1","9apr18_CAROLI_m1"
)

```

I have made a list of the old mice, but from the effect document the older mice might not be that different.

the excluded TOM, had low quality cells, only 4. The exclude AST, had low poor quality cells, there are 3 other mice.  The excluded CZECH males are poor quality (but near the mean). There are 3 other males.

```{r adj_nMLH1, echo=FALSE, eval=FALSE}
#add a column with male adjusted MLH1 values (+1 to all males)
MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)

#adjusted MLH1 no longer used
```

```{r reorder.DF, echo=FALSE}
#this is most important for plotting

#reorder dataframe
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name) )

#order the batches in MLH1 and original
#apply this ordering to all other DF?

```

```{r Batch.Stats, echo=FALSE, eval=TRUE}
#plot / table comparing number of images in quantification batches
#basic visualization for number of images in batches

#ToDo:  correct the case of all batches, better names
#set an order to batches?
original_batches <- as.data.frame(table(original_DF$Batch) ) #these look bad
colnames(original_batches) <- c("BatchNum", "orginal")

final_batches <- as.data.frame(table(MLH1_data$Batch) ) #these look bad
colnames(final_batches) <- c("BatchNum", "main")

batch.stats <- merge(original_batches,final_batches, by="BatchNum")## mm is the table of batch stats
##batch18 has the wrong number, ~390
```

All images were manually quantified after the file name was anonimized. The original (full) dataset contains the quantification batches with the following number of images. Each batch csv file was merged into Anon.csv by hand.

`r kable(final_batches, caption = "Number of Images across all the anonimized Batches") %>% kable_styling()`

# Quality Cleaning steps

The original dataset was cleaned up in the following ways before analysis. Settting up main data file with explanation and examination of data. Make the raw original DF and main working DF with bad cells exlcuede. The original dataset had `r length(original_DF_duplicates$fileName)` duplicated file names. These were removed and can be used to quantify repeatability or repair. Using the distinct() function for the MLH1_data, the first duplicated filename is kept and the second removed.

Take out mice from the exclude list based on the output from Mouse level quality plots.

1. The quality of MLH1 counts for individual mice is / was assessed in the MouseLevelPlots Rmd (can I link this file?). A list, mice.to.exclude was made.

2. The list of excluded mice is applied in the MLH1.setup.data script, to remove these mice from the main dataset/dataframe. This script saves .RData file of the dataset which is loaded into Rmd files for analysis

Mouse specific plots of CO count and quality were examined for general quality from single mice. Mice which produced few cells of low quality were excluded to the 'HQ' (high quality) dataset. (The full dataset was used to look at some patterns)

```{r Date.imaged, echo=FALSE}

#for issues with the total SC metrics, code in date imaged
#use the add_mouse function

#mini <- MLH1_data[1:500,]

count=1
for(i in MLH1_data$Original.Name){
   # print(i)
    templist= strsplit(i, split="_")[[1]]
    c = (templist[1])
    MLH1_data$image.string[count] <-  c
    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
    count= count +1
}
#works better if dat is converted out of image
MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')

#MLH1_data <- add_mouse2(MLH1_data)

#mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
#                        Nmice = length(unique(mouse)),
#                        Ncells  = length(Random.Name),
#                        #most recent date
#                        oldest = min(date.imaged),
#                        newest = max(date.imaged)
#)

```

# Mouse Data

Make mouse level tables  (re think this name, AP to distinguish BD -- but this isn't really an issue)

```{r all.data.table, echo=FALSE, eval=FALSE}

#make the table with ALL MLH1 obs
FULL_mouse_table_MLH1 <- ddply(MLH1_data, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

#NA cause by CAST female
#this table isn't used much
```

```{r HQ.mouse.table, echo=FALSE}
#Make a mouse level table

#change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
#

Mouse.Table_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm=TRUE),

                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm=TRUE),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        oldest.image.date = min(date.imaged),
                        newest.image.date = max(date.imaged)
                        #quality?
)

Mouse.Table_HQ <- add_strain(Mouse.Table_HQ)
Mouse.Table_HQ <- add_subsp(Mouse.Table_HQ)
Mouse.Table_HQ <- add_sex(Mouse.Table_HQ)
Mouse.Table_HQ <- add_category(Mouse.Table_HQ)
Mouse.Table_HQ <- add_species(Mouse.Table_HQ)

batch.mouse.count <- table(original_DF$mouse,original_DF$Batch) #mouse count by batch
batch.mouse.count.DF <- as.data.frame(batch.mouse.count)#this has a dif str, 3 cols -- like it was melted
batch.mouse.count.DF <- batch.mouse.count.DF[!(is.na(batch.mouse.count.DF$Freq)|batch.mouse.count.DF$Freq==0),]

#subsp in metadata not fully correct
meta.data <- add_strain(meta.data)
meta.data <- add_subsp(meta.data)
meta.data <- add_sex(meta.data)
meta.data <- add_category(meta.data)
meta.data <- add_species(meta.data)

#add DOB to this table?
mice.frm.meta <- Mouse.Table_HQ[as.character(Mouse.Table_HQ$mouse) %in% as.character(meta.data$mouse),]

Mouse.Table_HQ.OG <- Mouse.Table_HQ

Mouse.Table_HQ <- Mouse.Table_HQ.OG

#during merging -- try to just merge a few cols: 
Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
#this removes ~28 mice

old.mice.df <- Mouse.Table_HQ[(Mouse.Table_HQ$age.weeks > 25),]
#remove the NAs
old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]

Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]

#remove juvi mice via the list
#remove bad ages
#this is removing female

#this removes everything but males in these categories
#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]

#i don't understand why so many WSB female are removed..
# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)

```

Made up the main mouse level table:

- (Q1-Q4)

- no juvi mice

Subset DFs from here of other analysis.

```{r Q12_mouse.table, echo=FALSE}

#This DF is used for some MM, apply the HetC MM certia...
#this is a DF with a different set of quality certeria

#Make a mouse level table
Q12_mouse_table_MLH1 <- ddply(MLH1_data[MLH1_data$quality < 3,], c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)

#REMOVE mice with fewer than 3 cells
Q12_mouse_table_MLH1 <- Q12_mouse_table_MLH1[Q12_mouse_table_MLH1$Ncells > 3,]

Q12_mouse_table_MLH1 <- add_strain(Q12_mouse_table_MLH1)
Q12_mouse_table_MLH1 <- add_subsp(Q12_mouse_table_MLH1)
Q12_mouse_table_MLH1 <- add_sex(Q12_mouse_table_MLH1)
Q12_mouse_table_MLH1 <- add_category(Q12_mouse_table_MLH1)

#use things from prv block
#mice.frm.meta 
#meta.data
#mouse.table_w.Ages

#MERGE HERE
Q12_mouse_table_w.Ages <- merge(Q12_mouse_table_MLH1, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
#what are the differences between these two

#remove the NAs
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[!(is.na(Q12_mouse_table_w.Ages$category.x) | Q12_mouse_table_w.Ages$category.x==""), ]

#CLEAN up and format for mixed models
Q12_mouse_table_w.Ages <- subset(Q12_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

Q12_mouse_table_w.Ages <- add_strain(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_sex(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_category(Q12_mouse_table_w.Ages)

Q12_mouse_table_w.Ages <- add_species(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$species == "M.musculus",]

#REMOVE EXTRA STRAINS -- why?
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
#remove musc strains
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]

#REMOVE WRONG AGES
#afriad this is removing NA's
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[( (Q12_mouse_table_w.Ages$age.weeks < 25) &  (Q12_mouse_table_w.Ages$age.days != 13) ) | is.na(Q12_mouse_table_w.Ages$age.weeks), ] #keeps mice without ages, and removes juvinilles

#unclass order factors --(applies to the models)
#main df with variables coded as chr
Q12_mouse_table_w.Ages$subsp <- factor(Q12_mouse_table_w.Ages$subsp, ordered = FALSE)
Q12_mouse_table_w.Ages$sex <- factor(Q12_mouse_table_w.Ages$sex, ordered = FALSE)
Q12_mouse_table_w.Ages$strain <- factor(Q12_mouse_table_w.Ages$strain, ordered = FALSE)
Q12_mouse_table_w.Ages$species <- factor(Q12_mouse_table_w.Ages$species, ordered = FALSE)

```

```{r Q123_mouse.table, echo=FALSE, eval=FALSE}

Q123_mouse_table <- ddply(MLH1_data[MLH1_data$quality < 4,], c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)

#REMOVE mice with fewer than 3 cells
Q123_mouse_table <- Q123_mouse_table[Q123_mouse_table$Ncells > 3,]

Q123_mouse_table <- add_strain(Q123_mouse_table)
Q123_mouse_table <- add_subsp(Q123_mouse_table)
Q123_mouse_table <- add_sex(Q123_mouse_table)
Q123_mouse_table <- add_category(Q123_mouse_table)

```

### DataFrame for Mixed Models

Below I format the DF's for the models;

i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus. start with **AP_mouse_table**

```{r HetC.DF.4MM, echo=FALSE}

#variance is still chr
Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)

DF.HetC.MixedModel <- Mouse.Table_HQ %>% filter(species == "M.musculus")

#remove the extra merged cols
#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
#                            subsp.y,subsp.x,
#                    category.y,category.x, strain.x,
#                    strain.y, mouse.1  )  )

#REMOVE STRAINS WITHOUT FEMALES
#DF.HetC.MixedModel <- DF.HetC.MixedModel %>% filter( subsp != "Cast") %>% filter(strain != "AST")

# These are removed in the analysis
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
#F.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#Only us the righ age window
#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles

#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)

#remove blank rows (which are these?)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
#151 mice, no rows with nas for values

#main df with variables coded as chr
#DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
#DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
#DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
#only female male reps

```

```{r male.specific.table,echo=FALSE}

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels
mlh1male_mouse.av <-  Mouse.Table_HQ %>% filter(sex== "male")

#add the Rec.group markers
#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
#                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
#                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )

#formate the table for pluggin into models
```

update the dissection document (that makes the meta.data DF)

## Number of mice per category

 number of mice are in the main DF (discribe the results for missing mice)

Number of cells per mouse, (were any mice excluded for low numbers of cells?) (were all cells pooled?)

## Mouse Ages

Testes were collected from mice at various ages. For some poor breeding mice, I couldn't collect males at younger ages.

Is there a detectable age effect?

```{r Age.Hists, echo=FALSE, warning=FALSE}
#age.weeks col is missing

male_mouse_table_w.Ages <- Mouse.Table_HQ[(Mouse.Table_HQ$sex == "male") &(Mouse.Table_HQ$category != "other"),]

all.ages_plot <- ggplot(Mouse.Table_HQ[Mouse.Table_HQ$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
all.ages_plot

#male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[male_mouse_table_w.Ages$age.weeks < 20,]

male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[( (male_mouse_table_w.Ages$age.weeks < 25) &  (male_mouse_table_w.Ages$age.days != 13) ) | is.na(male_mouse_table_w.Ages$age.weeks), ]

#remove juvi mice!!
#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ]

#droplevels on DF above

good.ages_plot <- ggplot(male_mouse_table_w.good.ages, aes(age.weeks))+geom_histogram(binwidth = 1)+facet_wrap(~category)+ggtitle("Good Male Age Distributions")

#male_mouse_table_w.good.ages
```

# DMC1 Section

move the chunk below to setup

```{r DMC1.setup, echo=FALSE}
#load DMC1 csv
#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)

DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_category(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
                 .(category),
                 summarize,

                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.L$stage <- "early Z"

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "late Z"

DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)

DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
```

# BivData

for these full curated, add all the labels and save them as excel files in the BivData setup file.

full curated, reduced datasets

Are the strain levels plots needed?

Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.

```{r no.Foci.Pos, echo=FALSE}
#no posittion
#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]

#table(just.SC.length$category)
```

Main files from this Script:

  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures

  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information

  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)

  4. **MELT_ABS**: make melt version of foci positions for abs measuers

  5. CELL level table (full all mice) (use this to make long biv dataset)

  6. Long Biv Dataset

  7. mouse table (full all mice)

  8. **Strain DFs** for Het models (summary strain DFs and tables?)

  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models

  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**

# Male specific BivData set

not sure if SC.Length.male_table is used for anything.

```{r male.sc.length.strain, echo=FALSE}

#these measure ect should be in the male - mouse 
#isolate all musc and male
Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "M.musculus"),]

#make strain average table
SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)

```

# Cell Level Table

```{r cell.level.table, echo=FALSE}

cell.level.BivData_table <- ddply(.data=Curated_BivData, 
                       .(fileName),
                       summarize, 
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                       ncells = length(unique(fileName)),
                       nbivs = length(unique(Obj.ID)),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number

                       cv.SC = cv(chromosomeLength),

#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),

                #this below might not be correct
                     mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),

                       var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)

#cent_dis
                      mean_cent_dist = mean(dis.cent, na.rm=TRUE),
                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]], #calq threshold for all cells, pull out 'almost.whole.cells' later
                      Qar.3 = quantile(chromosomeLength, na.rm = TRUE)[[3]],
                      Qar.2 = quantile(chromosomeLength, na.rm = TRUE)[[2]],
                      Qar.1 = quantile(chromosomeLength, na.rm = TRUE)[[1]]

                        )

cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
cell.level.BivData_table <- add_sex(cell.level.BivData_table)
cell.level.BivData_table <- add_strain(cell.level.BivData_table)
cell.level.BivData_table <- add_category(cell.level.BivData_table)
cell.level.BivData_table <- add_species(cell.level.BivData_table)

#as.data.frame( head(cell.level.BivData_table) )
#use quartille 2 as a min, on average it grabs 4 
```

```{r show.pass.rate, echo=FALSE}

pass.rate.hist <- ggplot(cell.level.BivData_table, aes(x = nboxes_passed) )+geom_histogram(binwidth = 1, alpha=.6) + facet_wrap(~category)+ggtitle("number of objects ID'd per cell")
pass.rate.hist

```

make histogram of the pass rate per cell.

3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.  
8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev

# Reduced Bivalent Dataset (Long and Short)

I want to make a dataset of 'longest' 25% chromosomes from cells with at least 16 segmented bivalents from a single cell.

```{r Reduced.Biv.DataSet.chrm.dataset, echo=FALSE}

#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

#check for blank obj.id
BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")

#check for dup
#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
#dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
#no duplicates

#remove duplicate obj.ids
#BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
#BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]

#make sure obj.id and fileName are charcter for below
BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)

missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ] 
#all these files stem from

#these are bivalents which don't appear the in cell level table, which can break the loop
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
#final cleaned BivData table, 2428

#2. calq the Q4 threshold
#loop for greping the mouse from file name ect
#each Q is a quartille

#mini.now  <- BivData_from_15biv_cells[1:300,]

for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_" )[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q3 <-cell.level.BivData_table$Qar.3[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q2 <-cell.level.BivData_table$Qar.2[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q1 <-cell.level.BivData_table$Qar.1[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
    BivData_from_15biv_cells$cell.Q3[i] <- the.Q3
  BivData_from_15biv_cells$cell.Q2[i] <- the.Q2
  BivData_from_15biv_cells$cell.Q1[i] <- the.Q1

  the.FN <- as.character(cell.level.BivData_table$fileName[which(grepl(new.egg, cell.level.BivData_table$fileName ) )] )

  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)
  BivData_from_15biv_cells$Q1.FN[i] <- as.character(the.FN)
}

#3.mark which bivalents are above the threshold

BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= as.numeric(BivData_from_15biv_cells$cell.Q4)  ), "yes", "no" )

real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]#713 observations

#isolate the short bivdata
BivData_from_15biv_cells$short.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength <= as.numeric(BivData_from_15biv_cells$cell.Q2)  ), "yes", "no" )

real.short.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$short.Biv == "yes",]# 708 much few than

```

```{r Look.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}

#showing SC lengths per cell, with lines for the quartille values.

mini_G.15_15biv_FN <- BivData_from_15biv_cells %>% filter(category == "G male")
#ploting short biv data
G.15_15biv_male <- BivData_from_15biv_cells %>% filter(category == "G male")
mini_G.15_15biv_FN <- sample( G.15_15biv_male$fileName, 10)
mini_G.15_15biv  <- BivData_from_15biv_cells[BivData_from_15biv_cells$fileName %in% mini_G.15_15biv_FN, ]

ppll <- ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
  geom_hline(aes(yintercept= cell.Q2), color='red') +
 geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
 geom_hline(aes(yintercept= cell.Q4), color='black')

  #this plot shows that the Q1 is capturing only single values -- I should try Q2
#use Q2 as a 
```

```{r table.long.biv, echo=FALSE}

# make a table for the Mixed Models -- from the long.BivData
# Calq mouse averages

#add in all the other stats there are in the other tables

Long.biv_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        #SC lengths by class
                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

        #          mean_siscoten = mean(siscoten)
        #
)
#mouse averages for long.biv

Long.biv_mouse_table <- add_sex(Long.biv_mouse_table)
Long.biv_mouse_table <- add_strain(Long.biv_mouse_table)
Long.biv_mouse_table <- add_subsp(Long.biv_mouse_table)
Long.biv_mouse_table <- add_category(Long.biv_mouse_table)

#unordered for the mixed models (move this to the MM blocks...?)
Long.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
Long.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)

#rec group
Long.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Long.biv_mouse_table$category), 1, 
                            ifelse(grepl("MSM", Long.biv_mouse_table$category), 1,
                             ifelse(grepl("SKIVE", Long.biv_mouse_table$category), 1, 0)) )

#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)

#1) Long Biv for 
#remove all the non-used strains!!
Long.biv_mouse_table_HetC <- Long.biv_mouse_table

Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)

#remove na's
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]

#table(droplevels( Long.biv_mouse_table_HetC$category) )  #for double checking

#2) Long Biv for male.poly
#just males, just Musculus
Long.biv_mouse_table_male.poly <- Long.biv_mouse_table

Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]

#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
```

The above code chunk makes 2 mouse level tables; **Long.biv_mouse_table_HetC** and **Long.biv_mouse_table_male.poly**. Run these through the model systems.

# Mouse Level Single BivData Table

Mouse level tables are the main meat of the models. There will be several for use in different models. Format these tables for run the mixed models (unordered factors, ect).

The block below creates clean.BivData, which removes ~600 rows if just SC measures
(This might be important for the proportion code)

```{r FULL.mouse.BivData.table, echo=FALSE}

#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36

table(Curated_BivData$hand.foci.count, exclude = NULL)
#there are ~600 rows with SC and no hand foci count (blah)

#remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )

mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
                       .(mouse),
                       summarize,

                       ncells = length(unique(fileName)),

                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks

                      tot.biv = length( !is.na(hand.foci.count) ),

                       n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(Nbiv), 3),

                       #sis.co.ten
                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
)

mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
mouse.level.BivData_table <- add_category(mouse.level.BivData_table)

#unordered for the mixed models (move this to the MM blocks...?)
mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)

mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
```

```{r mouse.table.HetC.MM, echo=FALSE}

#this mouse level table -- should not have strains without female obs
#only Mmusculus

Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(  unique(fileName) ),
                      tot.biv = length(Obj.ID),

                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#
Mouse.Table_BivData_4MM <- add_sex(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_strain(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_subsp(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_category(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_species(Mouse.Table_BivData_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
Mouse.Table_BivData_4MM$subsp <- factor(Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
Mouse.Table_BivData_4MM$sex <- factor(Mouse.Table_BivData_4MM$sex, ordered = FALSE)
Mouse.Table_BivData_4MM$strain <- factor(Mouse.Table_BivData_4MM$strain, ordered = FALSE)

#remove all the non-used strains!!

Mouse.Table_BivData_4MM <-  Mouse.Table_BivData_4MM %>%  filter(strain != "AST")  %>% filter(strain != "TOM")  %>% filter(strain != "CZECH")

#Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
#Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),] #so many NAs still

Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
                              ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)))

```

```{r mouse.table.Male.Poly, echo=FALSE}

#this table will have male BivData
#it will include Musc strains

Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

Male.poly.Mouse.Table_BivData_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

Male.poly.Mouse.Table_BivData_4MM <- add_sex(Male.poly.Mouse.Table_BivData_4MM)
Male.poly.Mouse.Table_BivData_4MM <- add_strain(Male.poly.Mouse.Table_BivData_4MM)
Male.poly.Mouse.Table_BivData_4MM <- add_subsp(Male.poly.Mouse.Table_BivData_4MM)
Male.poly.Mouse.Table_BivData_4MM <- add_category(Male.poly.Mouse.Table_BivData_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
Male.poly.Mouse.Table_BivData_4MM$subsp <- factor(Male.poly.Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
Male.poly.Mouse.Table_BivData_4MM$sex <- factor(Male.poly.Mouse.Table_BivData_4MM$sex, ordered = FALSE)
Male.poly.Mouse.Table_BivData_4MM$strain <- factor(Male.poly.Mouse.Table_BivData_4MM$strain, ordered = FALSE)

#remove females
Male.poly.Mouse.Table_BivData_4MM$sex <- droplevels(Male.poly.Mouse.Table_BivData_4MM$sex)

```

analysis using the mouse level table -- to justify that using mouse averages is a good idea.

# Melt Position Table

The BivData MELT DF is primarily for the nrmF1.pos plots/analysis. It is a Melted DF for raw and normalized foci positions from the Curated_BivData.

```{r MELT.block1, echo=FALSE}

#not sure what the melt table is needed for
col.names.list <- colnames(Curated_BivData)

#all of the cols to use
MELT_Curated_BivData <-   melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse", "category",      "strain",        "sex",           "subsp",         "species", "IFD3_ABS", "IFD1_ABS","IFD2_ABS","IFD1_PER",
              #remoevd PER.Foci positions
"dis.cent", "dis.cent.PER",  "telo_dist", "telo_dist_PER", "SisCoTen", "Rec.group",
              #leave out the Foci
                "Foci1","Foci2","Foci3", "Foci4")],

            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
       "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count",

       "notes", "mouse", 
      "category","strain","sex","subsp","species", "IFD3_ABS","IFD1_ABS","IFD2_ABS","IFD1_PER",
              #remoevd PER.Foci positions
"dis.cent", "dis.cent.PER",  "telo_dist", "telo_dist_PER", "SisCoTen", "Rec.group"))

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]

#this might not be right
colnames(MELT_Curated_BivData) <- c( col.names.list[c(1:11, 16:26, 31:36)], "Foci_Type", "Pos")

#
MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)

#calq normalized melt pos
MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]

```

```{r SHORT.biv_setup, echo=FALSE, eval=TRUE}

#move to Data.setup

#FROM q1
#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
#from Q2 IFD
#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
##long biv
#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
#table( droplevels( male.long.biv$strain)  )

#FROM q1 
REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                        ncells = length(unique(fileName)),
                         mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                      #make sure 1CO  
                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)

                         )

REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
REAL.short_mouse_table <- add_category(REAL.short_mouse_table)

REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)

#divide musc males into groups
REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
                                           0)))

###LONG BIV

REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
                       n_obs = length(Obj.ID),
                        ncells = length(unique(fileName)),
                         mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
                         )

REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
REAL.long_mouse_table <- add_category(REAL.long_mouse_table)

REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)

#divide musc males into groups
REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
                                           0)))

#find which one

#REAL.short_mouse_table
```

# Total SC

total sc is processed in the results setup.

# Saving Data file

Now that the dataframes have been process and formated, they will be saved as RData. the main files are:
  1. MLH1_data and original_DF
  2. AP_mouse_table and AP_mouse_table_w.Ages
  3. meta.data

work on cleaning up all other df with bad names

```{r save.RData, echo=FALSE}

#remove non relevant files -- not sure where they load from
#BD data and Peromyscus data
#rm(c(BD_MLH1_data, BDmouse_level_table, MLH1_data.hq) )

save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")

#  OutPut: big large MLH1_data (AP's) df, big DF of BD with just the mice I want.
#  MLH1_data_table, means and variance of AP and BD MLH1 values. made from seperate tables from AP and BD data.
#  make sure decimal places are consistant

```

# Deciding on Mice for Final Dataset

Use the above information to make the final dataset for analysis

- age (for males)
- enough cells
- both male and female