### 2020-05-16 Chap3_MLH1var.rmd
notes and updates of the chapter 3 variances

---
title: "Between Cell Variance"
author: "April Peterson"
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(tidyverse)
library(knitr)
library(ggplot2)
#library(pwr)
library(purrr) #for replicate
library(dplyr)
library(plyr)
library(lattice)

library(raster)
library(lme4)
library(nlme)
library(RLRsim)
#library(ggpubr)
library(reshape2)
library(patchwork)
library(png)
library(grid)
library(gridBase)
library(gridExtra)
library(dplyr)
library(cowplot)
library(kableExtra)
library(xlsx)

#set wd
#load datafile

options(digits=2)

#setwd("~.")#switch over to here()

load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")

load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

```{r female.strain.table, echo=FALSE}

female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )

```

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

#variance plots
musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse \nMLH1 variance")+scale_color_manual(values=colors_of_strains)

Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female\nwithin mouse MLH1 variance")+scale_color_manual(values=colors_of_strains)

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse \nMLH1 variance") + scale_color_manual(values=colors_of_strains)

```

```{r male.strain.table, echo=FALSE}

M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        totNcells  = sum(Ncells, na.rm=TRUE),

                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm=TRUE),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#M.strain.table
#same issue as the female table .. ugh

```

```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

#variance plots
OG.male.VAR.plot <- ggplot(M.strain.table, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)

```

> How much do CO counts vary among cells from the same individual?

> What factors shape between-cell variance in CO count?

Answering these will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution

Below is the table for the number of cells per category by quality score. For these between cell variance will just use those in quality 1-2.

```{r OG.table.qual.num, echo=FALSE}

table(MLH1_data$quality, MLH1_data$category, exclude = NULL)

```

# Q12 Varince Plots

The plots below show each point as a mouse-level CO count variance, calculated from Quality 1-2 cells. The number of cells per mouse is indicated by size.

```{r Q12.table.qual.num, echo=FALSE, warning=FALSE}

#high quality cells
Q12.MLH1_data <- MLH1_data %>% filter(quality < 3 )

Q12.MLH1_data$fileName <- Q12.MLH1_data$Original.Name

#table(Q12.MLH1_data$quality, Q12.MLH1_data$category, exclude = NULL)
#all this information is in the top table

#make mouse level table
FULL_mouse_table_Q12.MLH1 <- ddply(Q12.MLH1_data, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

FULL_mouse_table_Q12.MLH1 <- add_strain(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_subsp(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_sex(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_category(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_species(FULL_mouse_table_Q12.MLH1)

```

```{r Q12.mouse.var.plots.male, echo=FALSE, fig.width = 6, fig.height=6, warning=FALSE}

#variance plots
#musc.male.VAR.plot  #%>% filter(subsp == "Musc")

FULL.male.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))

Dom.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Dom")  ),
                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

mol.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Mol")  ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

FULL.male.Var.plot
```

```{r Q12.mouse.var.plots.female, echo=FALSE, fig.width = 5, fig.height=5, warning=FALSE}

FULL.female.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))

#variance plots
musc.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Musc")  ),
                             aes(color= strain, size = Ncells))+
  geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

Dom.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Dom")  ),
                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

mol.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Mol")  ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

FULL.female.Var.plot
```

The figure above nominates MOLF should be removed from the full dataset. I think all cells can be included for the sex specific data.

The female LEW strains have a larger distribution of variance. (or it could just be the 2 mice). Some of the low cell number mouse.  (note on the dot size, small dots could be mice with few high Q cells. Med size cells have a higher number of cells.)

# Models (var)

The same glm models are being applied to variance and cV (M1 to M3) and sex specific models (M4 and M5). Out groups and strains with just one sex have been removed for M1-M3. Al factors are unordered.

```{r model.set.up, echo=FALSE}

#FULL_mouse_table_Q12.MLH1
#sex-specific

FULL_mouse_table_Q12.MLH1$subsp <- factor(FULL_mouse_table_Q12.MLH1$subsp, ordered = FALSE)
FULL_mouse_table_Q12.MLH1$sex <- factor(FULL_mouse_table_Q12.MLH1$sex, ordered = FALSE)
FULL_mouse_table_Q12.MLH1$strain <- factor(FULL_mouse_table_Q12.MLH1$strain, ordered = FALSE)

#remove all the mice with fewer than 3cells
One.cell.mice <- FULL_mouse_table_Q12.MLH1 %>% filter(Ncells == 1)
oo <- FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$mouse %in% One.cell.mice$mouse, ]
```

```{r varMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}

#FULL_mouse_table_Q12.MLH1, remove molf not enough female

#DF.for.model <- ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )

#these are all mouse level values

varMLH1.M1 <- lmer(var ~ subsp * sex + (1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )  )

varMLH1.MM.M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

varMLH1.MM.M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

#summary(varMLH1.M1)$coefficients

varMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
varMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
varMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
varMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value

varMLH1.MMlmer_results <- data.frame(sub.sp = varMLH1.MM_subp, 
sex.results =  varMLH1.MM_sex, 
               interaction.pval = varMLH1.MM_interaction,  rand.pval = 
varMLH1.MM_strain)

```

**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain `r varMLH1.MMlmer_results`. (Sex is the most significant).

**M2**
```{r varMLH1.glm.M2, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed, 
varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

summ.varMLH1.M2.glm.strain.fixed.full <- summary(varMLH1.M2.glm.strain.fixed.full)
round(summ.varMLH1.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
#adding cell number has some significant parts
```

**M3**
```{r varMLH1.glm.M3, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

varMLH1.M3.strain.fixed <- glm(var ~ sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ] )

summ.MLH1.M3.strain.fixed <- summary(varMLH1.M3.strain.fixed)

round(summ.MLH1.M3.strain.fixed$coefficients[,c(1,4)],5)
#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
```

LEW and G are slightly significant strain p values. Sex is also slightly significant. Intercept is also significant.

**M4-M5 female**
```{r Female.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

Var.female.strain.subsp <- glm(var ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

Var.summry.female.strain.effects <- summary(Var.female.strain.subsp)
round(Var.summry.female.strain.effects$coefficients[,c(1,4)],5)

varMLH1.female.glm.M2 <- glm(var ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

sum_varMLH1.female.spc.glm.M2 <- summary(varMLH1.female.glm.M2)
round(sum_varMLH1.female.spc.glm.M2$coefficients[,c(1,4)],5)
```

Only the intercept's are significant in the female specific models. None of the subsp or strain effects are significant.

**M4-M5 male**
```{r Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

#M4
varMLH1.male.glm.M4 <- glm(var ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
sum_varMLH1.male.glm.M4 <- summary(varMLH1.male.glm.M4)
round(sum_varMLH1.male.glm.M4$coefficients[,c(1,4)],5)

#M4.2
varMLH1.male.glm.M4.2 <- glm(var ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )

sum_varMLH1.male.glm.M4.2 <- summary(varMLH1.male.glm.M4.2)
round(sum_varMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)
```

There are no signifcant subspecies + strain effects for the male specific data.

# Models (cV)
```{r cVMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}

#FULL_mouse_table_Q12.MLH1, remove molf not enough female

#DF.for.model <- ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )

#these are all mouse level values

cVMLH1.M1 <- lmer(cV ~ subsp * sex + (1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )  )

cVMLH1.MM.M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

cVMLH1.MM.M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

#summary(varMLH1.M1)$coefficients

cVMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
cVMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
cVMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value

cVMLH1.MMlmer_results <- data.frame(sub.sp = cVMLH1.MM_subp, 
sex.results =  varMLH1.MM_sex, 
               interaction.pval = cVMLH1.MM_interaction,  rand.pval = 
cVMLH1.MM_strain)

```

**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain `r cVMLH1.MMlmer_results`. (Sex is the most significant).

**M2**
```{r cVMLH1.glm.M2, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed, 
cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

summ.cvMLH1.M2.glm.strain.fixed.full <- summary(cVMLH1.M2.glm.strain.fixed.full)
round(summ.cvMLH1.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)

#adding cell number has some significant parts
```

**M3**
```{r cVMLH1.glm.M3, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

cVMLH1.M3.strain.fixed <- glm(cV ~ sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ] )

summ.cV.MLH1.M3.strain.fixed <- summary(cVMLH1.M3.strain.fixed)

round(summ.cV.MLH1.M3.strain.fixed$coefficients[,c(1,4)],5)

```

LEW and G are slightly significant strain p values. Sex is also slightly significant. Intercept is also significant.

**M4-M5 female**
```{r Female.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

cV.female.strain.subsp <- glm(cV ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

cV.summry.female.strain.effects <- summary(cV.female.strain.subsp)
round(cV.summry.female.strain.effects$coefficients[,c(1,4)],5)

cVMLH1.female.glm.M2 <- glm(cV ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

sum_cVMLH1.female.spc.glm.M2 <- summary(cVMLH1.female.glm.M2)
round(sum_cVMLH1.female.spc.glm.M2$coefficients[,c(1,4)],5)

```

Only the intercept's are significant in the female specific models. None of the subsp or strain effects are significant.

**M4-M5 male**
```{r Male.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

#M4
cVMLH1.male.glm.M4 <- glm(cV ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
sum_varMLH1.male.glm.M4 <- summary(varMLH1.male.glm.M4)
round(sum_varMLH1.male.glm.M4$coefficients[,c(1,4)],5)

#M4.2
cVMLH1.male.glm.M4.2 <- glm(cV ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )

sum_cVMLH1.male.glm.M4.2 <- summary(cVMLH1.male.glm.M4.2)
round(sum_cVMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)

```

<!-- Bret's suggestions / Permutations

On the question of variance in MLH1 count, I agree that it is crucial to reduce experimental noise to find a signal.

I would suggest the following approach:
1. Consider only high-quality cells.
2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.

3. Randomly permute cells among strains, keeping the structure of the dataset otherwise the same.

4. For each permutation, compute the among-strain variance in the MLH1 count variance (separately for each sex).
Compute the among-strain variance in the MLH1 count variance (separately for each sex) in your original data.
Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).
-->

# Permutations

Calculate the category level between cell variance (and cV).

```{r pool.cell.calq, echo=FALSE}

OG.vals_Q12.MLH1 <- ddply(Q12.MLH1_data, c("strain", "sex"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

OG.vals_Q12.MLH1 <- add_subsp(OG.vals_Q12.MLH1)
OG.vals_Q12.MLH1 <- add_species(OG.vals_Q12.MLH1)

```

Made a table with the cell-pooled calculations of var and cV.
Next step is to i) record the data structure ii) then permute within sex, then calculate the cell-pooled values.

```{r permutations, echo=FALSE, eval=FALSE}

Q12.permute.DF <- Q12.MLH1_data %>% filter()

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]

#Data structure
len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )

len.G.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
len.G.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )

len.LEW.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
len.LEW.male <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "male") )$fileName )

len.PWD.female <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "female") )$fileName )
len.PWD.male <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "male") )$fileName )

len.KAZ.female <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "female") )$fileName )
len.KAZ.male <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "male") )$fileName )

len.SKIVE.female <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$fileName )
len.SKIVE.male <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$fileName )

len.MSM.female <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "female") )$fileName )
len.MSM.male <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "male") )$fileName )

#these are the bones of the structure -- for permuting a new DF -- then calculating

#this DF is kinda hard, I'm not sure how to make a full randomized DF

#replicate(n = 3, rnorm(5, 0, 1), simplify = FALSE )

nrepeat = 50

permute.var <- data.frame( "wsb.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.WSB.female, replace = FALSE) ) ),

                            "wsb.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
  "lew.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.LEW.female, replace = FALSE) ) ),

  "lew.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.LEW.male, replace = FALSE) ) ),

  "g.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.G.female, replace = FALSE) ) ), 
  "g.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.G.male, replace = FALSE) ) ),

  "pwd.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
  "pwd.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.PWD.male, replace = FALSE))),

  "kaz.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.KAZ.female,replace = FALSE) ) ),

  "kaz.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.KAZ.male, replace = FALSE)) ),

  "skive.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),

  "skive.m.var" =  replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),
  "msm.f.var" = replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.MSM.male, replace = FALSE) )   )  )

#now that these is a dataframe of new/permuted variance (specific to each sex) ... compare them to the origincal values?
#my prediction is that there would be many significant differences from the true values

#to visualize, I can plot the variances .. I guess with a histogram

#now automate making these plots -- or calculate the propotion of permutations which are below the observed

pp <- ggplot(permute.var, aes(x = wsb.f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )

#randomized data frame, sample mouse means from all P.leucopus mice

#calculate pvalue as proportion of differences greater than observed difference
p.val_permut_mean.lab.wild <- length(permute.dif.meanOmeans$permuted.diff[permute.dif.meanOmeans$permuted.diff >= obs.dif.mean.lab.wild]) / permut.x
var.pval <- paste( "p-value ", p.val_permut_mean.lab.wild)

#visualize the permuted distribution
permute_hist_means  <-  ggplot(permute.dif.meanOmeans,aes(x=permuted.diff))+
  geom_histogram()+geom_vline(xintercept=obs.dif.mean.lab.wild,color='red') +
  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of mouse means, wild to lab")

#answer: The difference in CO means between wild and LL Pleucopus, is greater than if mice are randomized.

#Same permutations, at the cell level
Pl.cells <- MLH1_data.hq[MLH1_data.hq$species == "Pleuc",]

obs.status.cell.diff <- mean(Pl.cells$nMLH1.foci[Pl.cells$status == "wild"])-
    mean(Pl.cells$nMLH1.foci[Pl.cells$status == "lab"])

permute.dif.cell.status  <- as.data.frame( replicate(permut.x, 
                                  (  mean(sample(Pl.cells$nMLH1.foci,  243, replace = FALSE) ) - 
                                   mean(sample(Pl.cells$nMLH1.foci, 221, replace = FALSE) ) )  
) )

colnames(permute.dif.cell.status) <- "permuted.diff"
p.val_permut_cell.diff.lab.wild <- length(permute.dif.cell.status$permuted.diff[permute.dif.cell.status$permuted.diff >= obs.status.cell.diff]) / permut.x
var.pval <- paste( "p-value ", p.val_permut_cell.diff.lab.wild)

permute_hist_means.cell_level  <-  ggplot(permute.dif.cell.status, aes(x=permuted.diff))+
  geom_histogram()+geom_vline(xintercept=obs.status.cell.diff,color='red') +
  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of cell level means, wild to lab")

```