### 2020-05-21 Chap3_MLH1var.rmd
notes and preliminary results for Chap3

---
title: "Between Cell Variance"
author: "April Peterson"
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(tidyverse)
library(knitr)
library(ggplot2)
#library(pwr)
library(purrr) #for replicate
library(dplyr)
library(plyr)
library(lattice)
library(tidyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
#library(ggpubr)
library(reshape2)
library(patchwork)
library(png)
library(grid)
library(gridBase)
library(gridExtra)
library(dplyr)
library(cowplot)
library(kableExtra)
library(xlsx)

library(ggpubr) #for arranging ggplots

#set wd
#load datafile

options(digits=2)

#setwd("~.")#switch over to here()

load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")

#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize
#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

```{r female.strain.table, echo=FALSE}

female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )

```

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

#variance plots
musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse \nMLH1 variance")+scale_color_manual(values=colors_of_strains)

Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female\nwithin mouse MLH1 variance")+scale_color_manual(values=colors_of_strains)

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse \nMLH1 variance") + scale_color_manual(values=colors_of_strains)

```

```{r male.strain.table, echo=FALSE}

M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        totNcells  = sum(Ncells, na.rm=TRUE),

                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm=TRUE),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#M.strain.table
#same issue as the female table .. ugh

```

```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

#variance plots
OG.male.VAR.plot <- ggplot(M.strain.table, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)

```

#Writing Outline

I find these results encouraging. The higher variance in females is visible across a variety of genetic backgrounds, including those in which females have lower means than males.

I feel like this result solidifies a plan for Chapter 3.

It could take the following basic form:
Lower stringency of checkpoints during oogenesis predicts higher variance in crossover number within females
Despite acceptance of this prediction, it has rarely been tested
We tested the prediction across a wide array of genetic backgrounds
Females consistently have higher variance than males in the same inbred line, with little evidence for line effects
[Perhaps something about variance in SC length]
In terms of recombination mechanisms, our results mean ...
In terms of evolutionary mechanisms, our results imply ...

> How much do CO counts vary among cells from the same individual?

> What factors shape between-cell variance in CO count?

The approach will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution.

Below is the table for the number of cells per category by quality score. For this document just the cells with quality 1-2 will be used.

```{r OG.table.qual.num, echo=FALSE}

table(MLH1_data$quality, MLH1_data$category, exclude = NULL)

```

# Q12 Variance Plots

The plots below show each point as a mouse-level CO count variance, calculated from Quality 1-2 cells. The number of cells per mouse is indicated by size.

```{r Q12.table.qual.num, echo=FALSE, warning=FALSE}

#high quality cells
Q12.MLH1_data <- MLH1_data %>% filter(quality < 3 )

Q12.MLH1_data$fileName <- Q12.MLH1_data$Original.Name

#table(Q12.MLH1_data$quality, Q12.MLH1_data$category, exclude = NULL)
#all this information is in the top table

#make mouse level table
FULL_mouse_table_Q12.MLH1 <- ddply(Q12.MLH1_data, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

FULL_mouse_table_Q12.MLH1 <- add_strain(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_subsp(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_sex(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_category(FULL_mouse_table_Q12.MLH1)
FULL_mouse_table_Q12.MLH1 <- add_species(FULL_mouse_table_Q12.MLH1)

```

```{r Q12.mouse.var.plots.male, echo=FALSE, fig.width = 6, fig.height=6, warning=FALSE}

#variance plots
#musc.male.VAR.plot  #%>% filter(subsp == "Musc")

FULL.male.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain), width = .2, alpha=0.5 )+ ggtitle("Q12 VAR male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))

FULL.male.Var.plot
```

```{r Q12.mouse.var.plots.female, echo=FALSE, fig.width = 5, fig.height=5, warning=FALSE}

FULL.female.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain), width = .2, alpha=0.5 )+ ggtitle("Q12 VAR female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free", ncol = 2)+ylim(c(0,55))

FULL.female.Var.plot

#variance plots
musc.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Musc")  ),
                             aes(color= strain, size = Ncells))+
  geom_jitter(aes(x = as.factor(strain), y =var, color= strain), alpha=0.5, width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

```

# observed values

```{r pool.cell.calq, echo=FALSE}

OG.vals_Q12.MLH1 <- ddply(Q12.MLH1_data, c("strain", "sex"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(mean(nMLH1.foci), 3), nsmall=3)),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round( var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

OG.vals_Q12.MLH1 <- add_subsp(OG.vals_Q12.MLH1)
OG.vals_Q12.MLH1 <- add_species(OG.vals_Q12.MLH1)

#remove extra strains
remove.strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
vals_Q12.MLH1 <- OG.vals_Q12.MLH1[!OG.vals_Q12.MLH1$strain %in% remove.strains, ]

#for males and females, calq the total variance across the strain-specific var
Obs.f.VAS <- var( ( vals_Q12.MLH1 %>% filter(sex == "female") )$var )
Obs.m.VAS <- var( ( vals_Q12.MLH1 %>%  filter(sex == "male") )$var )

```

These are the female strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$var`, the variance across strains (VAS) is `r Obs.f.VAS`.

These are the male strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "male") )$var`, the variance across strains (VAS) is `r Obs.m.VAS`. (For `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$strain` respectively)

```{r str.of.DF, echo=FALSE, eval=TRUE}
#redundant
#remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
#Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]

#Data structure (from the cell level data)
len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )

len.G.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
len.G.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )

len.LEW.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
len.LEW.male <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "male") )$fileName )

len.PWD.female <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "female") )$fileName )
len.PWD.male <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "male") )$fileName )

len.KAZ.female <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "female") )$fileName )
len.KAZ.male <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "male") )$fileName )

len.SKIVE.female <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$fileName )
len.SKIVE.male <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$fileName )

len.MSM.female <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "female") )$fileName )
len.MSM.male <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "male") )$fileName )

```

# Mean and Variance correlations

```{r var.x.mean, echo=FALSE, warning=FALSE, message=FALSE}

#FULL_mouse_table_Q12.MLH1
Fcor.test <- cor.test( (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_co,
  (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female")%>% filter(species == "M.musculus") )$var)

Mcor.test <- cor.test( (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus")%>% filter(species == "M.musculus"))$mean_co,
  (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus"))$var)

small.DF <- data.frame( corr = paste("corr =", c(Fcor.test$estimate, Mcor.test$estimate) ), 
                     p.val = paste("p.val =", c(Fcor.test$p.value, Mcor.test$p.value) ), sex = c("female", "male") )

Cor.plot <- ggplot(( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus") ), aes(x=mean_co, y=var, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= small.DF, aes(x=27, y=80, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= small.DF, aes(x=27, y=75, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by CO variance per Mouse")

Cor.plot
```

# Models MLH1(var)

The same glm models are being applied to variance and cV (M1 to M3) and sex specific models (M4 and M5). Out groups and strains with just one sex have been removed for M1-M3. Al factors are unordered.

```{r model.set.up, echo=FALSE}

#FULL_mouse_table_Q12.MLH1
#sex-specific

FULL_mouse_table_Q12.MLH1$subsp <- factor(FULL_mouse_table_Q12.MLH1$subsp, ordered = FALSE)
FULL_mouse_table_Q12.MLH1$sex <- factor(FULL_mouse_table_Q12.MLH1$sex, ordered = FALSE)
FULL_mouse_table_Q12.MLH1$strain <- factor(FULL_mouse_table_Q12.MLH1$strain, ordered = FALSE)

#remove all the mice with fewer than 3cells
One.cell.mice <- FULL_mouse_table_Q12.MLH1 %>% filter(Ncells == 1)
oo <- FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$mouse %in% One.cell.mice$mouse, ]
```

```{r varMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}

#FULL_mouse_table_Q12.MLH1, remove molf not enough female

#DF.for.model <- ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )

#these are all mouse level values

varMLH1.M1 <- lmer(var ~ subsp * sex + (1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )  )

varMLH1.MM.M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

varMLH1.MM.M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

#summary(varMLH1.M1)$coefficients

varMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
varMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
varMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
varMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value

varMLH1.MMlmer_results <- data.frame(sub.sp = varMLH1.MM_subp, 
sex.results =  varMLH1.MM_sex, 
               interaction.pval = varMLH1.MM_interaction,  rand.pval = 
varMLH1.MM_strain)

```

**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain `r varMLH1.MMlmer_results`. (Sex is the most significant).

**M2**
```{r varMLH1.glm.M2, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed, 
varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

summ.varMLH1.M2.glm.strain.fixed.full <- summary(varMLH1.M2.glm.strain.fixed.full)
round(summ.varMLH1.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
#adding cell number has some significant parts
```

**M3**
```{r varMLH1.glm.M3, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

varMLH1.M3.strain.fixed <- glm(var ~ sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ] )

summ.MLH1.M3.strain.fixed <- summary(varMLH1.M3.strain.fixed)

round(summ.MLH1.M3.strain.fixed$coefficients[,c(1,4)],5)
#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
```

LEW and G are slightly significant strain p values. Sex is also slightly significant. Intercept is also significant.

**M4-M5 female**
```{r Female.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

Var.female.strain.subsp <- glm(var ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

Var.summry.female.strain.effects <- summary(Var.female.strain.subsp)
round(Var.summry.female.strain.effects$coefficients[,c(1,4)],5)

varMLH1.female.glm.M2 <- glm(var ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

sum_varMLH1.female.spc.glm.M2 <- summary(varMLH1.female.glm.M2)
round(sum_varMLH1.female.spc.glm.M2$coefficients[,c(1,4)],5)
```

Only the intercept's are significant in the female specific models. None of the subsp or strain effects are significant.

**M4-M5 male**

```{r Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

#M4
varMLH1.male.glm.M4 <- glm(var ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
sum_varMLH1.male.glm.M4 <- summary(varMLH1.male.glm.M4)
round(sum_varMLH1.male.glm.M4$coefficients[,c(1,4)],5)

#M4.2
varMLH1.male.glm.M4.2 <- glm(var ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )

sum_varMLH1.male.glm.M4.2 <- summary(varMLH1.male.glm.M4.2)
round(sum_varMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)
```

There are no signifcant subspecies + strain effects for the male specific data.

# Models MLH1(cV)

```{r cVMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
#these are all mouse level values

cVMLH1.M1 <- lmer(cV ~ subsp * sex + (1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )  )

cVMLH1.MM.M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

cVMLH1.MM.M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )

#summary(varMLH1.M1)$coefficients

cVMLH1.MM_interaction <- drop1(cVMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
cVMLH1.MM_sex <- anova(cVMLH1.M1, cVMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
cVMLH1.MM_subp <- anova(cVMLH1.M1, cVMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(cVMLH1.M1)$p.value

cVMLH1.MMlmer_results <- data.frame(sub.sp = cVMLH1.MM_subp, 
sex.results =  cVMLH1.MM_sex, 
               interaction.pval = cVMLH1.MM_interaction,  rand.pval = 
cVMLH1.MM_strain)

```

**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain `r cVMLH1.MMlmer_results`. (Sex is the most significant).

**M2**

```{r cVMLH1.glm.M2, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed, 
cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

summ.cvMLH1.M2.glm.strain.fixed.full <- summary(cVMLH1.M2.glm.strain.fixed.full)
round(summ.cvMLH1.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)

#adding cell number has some significant parts
```

**M3**

```{r cVMLH1.glm.M3, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

cVMLH1.M3.strain.fixed <- glm(cV ~ sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ] )

summ.cV.MLH1.M3.strain.fixed <- summary(cVMLH1.M3.strain.fixed)

round(summ.cV.MLH1.M3.strain.fixed$coefficients[,c(1,4)],5)

```

LEW and G are slightly significant strain p values. Sex is also slightly significant. Intercept is also significant.

**M4-M5 female**

```{r Female.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

cV.female.strain.subsp <- glm(cV ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

cV.summry.female.strain.effects <- summary(cV.female.strain.subsp)
round(cV.summry.female.strain.effects$coefficients[,c(1,4)],5)

cVMLH1.female.glm.M2 <- glm(cV ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

sum_cVMLH1.female.spc.glm.M2 <- summary(cVMLH1.female.glm.M2)
round(sum_cVMLH1.female.spc.glm.M2$coefficients[,c(1,4)],5)

```

Only the intercept's are significant in the female specific models. None of the subsp or strain effects are significant.

**M4-M5 male**

```{r Male.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

#M4
cVMLH1.male.glm.M4 <- glm(cV ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
sum_varMLH1.male.glm.M4 <- summary(varMLH1.male.glm.M4)
round(sum_varMLH1.male.glm.M4$coefficients[,c(1,4)],5)

#M4.2
cVMLH1.male.glm.M4.2 <- glm(cV ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )

sum_cVMLH1.male.glm.M4.2 <- summary(cVMLH1.male.glm.M4.2)

round(sum_cVMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)

```

<!-- Bret's suggestions / Permutations

On the question of variance in MLH1 count, I agree that it is crucial to reduce experimental noise to find a signal.

I would suggest the following approach:
1. Consider only high-quality cells.

2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.

3. Randomly permute cells among strains, keeping the structure of the dataset otherwise the same.

4. For each permutation, compute the among-strain variance in the MLH1 count variance (separately for each sex).
Compute the among-strain variance in the MLH1 count variance (separately for each sex) in your original data.
Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).

-->

# Permutations; MLH1

We'd like to know if the set of strains has bigger differences in between-cell variance than expected by chance. Your liner models provide one test, but they could suffer from small sample sizes per mouse. We can conduct a permutation test that enables you to pool all cells in each strain to increase the sample size.

<!--
Here's the permutation test I was imagining:
1. Separate the sexes.

2. Pool all cells within a strain and compute the **variance for each strain**.

3. Compute the **variance across strains** (VAS) of the variance within strains. This will leave you with a single estimate for the set of strains.

4. Randomly permute cells across strains. (pool all male cells, pool all female cells)

5. For the permuted data set, pool all cells within a strain and compute the variance for each strain. (compute a variance for a strain, var( sample( all.female.cells, N.WSB.f ) ) ).

6. For the permuted data set, compute the variance across strains of the variance within strains. This will leave you with a single estimate for the set of strains.

7. Repeat 4-6 10,000 times to establish a null distribution for the variance in variances.

8. Compare observed variance in variances to null distribution.  
-->

```{r permutations, echo=FALSE}
#I made a DF of strain X sex specific variance

nrepeat = 10000

#I need to sample within strain? -- then I feel like I need to change the sample size -- otherwise I'm just sampling all the same episodes

permuted.var <- data.frame( "wsb_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$sex == "female")], len.WSB.female, replace = FALSE) ) ),
                            "wsb_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
  "lew_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.LEW.female, replace = FALSE) ) ),

  "lew_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.LEW.male, replace = FALSE) ) ),

  "g_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.G.female, replace = FALSE) ) ), 
  "g_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.G.male, replace = FALSE) ) ),

  "pwd_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
  "pwd_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.PWD.male, replace = FALSE))),

  "kaz_f.var" = replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"],len.KAZ.female,replace = FALSE) ) ),

  "kaz_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.KAZ.male, replace = FALSE)) ),

  "skive_f.var" = replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),

  "skive_m.var" =  replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),

  "msm_f.var" = replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.MSM.male, replace = FALSE) )   )  
  )

```

```{r functions.Chap3, echo=FALSE}
#formate the permuted DF, adding strain and sex to the DF

add_sex_chp3 <- function(dat){
  dframe <- dat
  dframe$sex <- "other"
  dframe$sex <- ifelse(grepl("f.", dframe$names), "female", 
                       ifelse(grepl("m.",dframe$names), "male", "other"))
  dframe$sex<- factor(dframe$sex, ordered = TRUE, levels =c( "female", "male"))
  return(dframe)
}

add_strain_Chap3 <- function(dat){
  dframe <- dat
  dframe$strain <- "other"
  dframe$strain <- ifelse(grepl("wsb", dframe$names), "WSB",
                    ifelse(grepl("g",dframe$names), "G",
                   ifelse(grepl("lew", dframe$names), "LEW",
                     ifelse(grepl("perc", dframe$names), "PERC",
                     ifelse(grepl("msm", dframe$names), "MSM",
                 ifelse(grepl("molf", dframe$names), "MOLF",
                     ifelse(grepl("skive", dframe$names), "SKIVE",      
                      ifelse(grepl("pwd", dframe$names), "PWD",
                  ifelse(grepl("kaz", dframe$names), "KAZ",
                   ifelse(grepl("tom", dframe$names), "TOM",
                  ifelse(grepl("ast", dframe$names), "AST",
                   ifelse(grepl("czech", dframe$names), "CZECH",
                    ifelse(grepl("cast", dframe$names), "CAST",
                   ifelse(grepl("hmi", dframe$names), "HMI",
                       ifelse(grepl("_SPRET_", dframe$names), "SPRET",
                       ifelse(grepl("_SPI_", dframe$names), "SPIC",
                                                                                                                                          ifelse(grepl("spic", dframe$names), "SPIC",
                                                                                                                                                 ifelse(grepl("_CAROLI_", dframe$names), "CAROLI",
                                 ifelse(grepl("F1_", dframe$names), "F1",         
                                   "other")))))))))))))))))))

  #the ordering factor below deletes all strain entries
  dframe$strain<- factor(dframe$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW", "PERC",
                                                                   "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                                   "CAST", "HMI",
                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
  return(dframe)
}

```

```{r pvals.VAS, echo=FALSE, warning=F, message=FALSE,  fig.width = 5, fig.height=5}

#need to merge all the pvales into a DF
#for(u in unique( OG.vals_Q12.MLH1$strain[ ( (OG.vals_Q12.MLH1$sex == "female") & (OG.vals_Q12.MLH1$species == "M.musculus") )] ) ){
#  print(u)
#  }

transposed.permuted <- t(permuted.var)

permuted.DF <- data.frame(names = row.names(transposed.permuted), transposed.permuted); rownames(transposed.permuted) <- NULL

permuted.DF <- add_sex_chp3(permuted.DF)
permuted.DF <- add_strain_Chap3(permuted.DF)

# 
VAS.F.permuted.DF <- permuted.DF %>% filter(sex == "female")
VAS.M.permuted.DF <- permuted.DF %>% filter(sex == "male")

#careful aboout matching the size of the dataframe!!
#calculate the variance across permutations
list.VAS_F <- c()
for(ff in 1:length(VAS.F.permuted.DF[,2:9997]) ){
  list.VAS_F[ff]  <- var( VAS.F.permuted.DF[,ff] )
}
#list of VAS
list.VAS_F.DF <- data.frame( "female.VAS" =  list.VAS_F)

list.VAS_M <- c()
for(mm in 1:length(VAS.M.permuted.DF[,2:9997]) ){

  perm.VAS.mm <- var( VAS.M.permuted.DF[,mm] )
  list.VAS_M[mm]  <- perm.VAS.mm

}
#p value is calculated for the distribution.

list.VAS_M.DF <- data.frame("male.VAS" = list.VAS_M)
#cbind the 2 VAS DF
Full.VAS <- cbind( list.VAS_F.DF, list.VAS_M.DF)

male.plot <- ggplot(Full.VAS, aes(x = male.VAS)) + geom_histogram() + geom_vline(xintercept = Obs.m.VAS ) + ggtitle("Male Variance Across Species")

female.plot <- ggplot(Full.VAS, aes(x = female.VAS)) + geom_histogram() + geom_vline(xintercept = Obs.f.VAS ) + ggtitle("Female Variance Across Species")

male.plot
female.plot
```

The plots above show the distributions of permuted VAS for 10,000 permutations, the vertical line shows the observed VAS.

The proportion of permutations which are less than the oberserved VAS are `r sum(  Full.VAS$male.VAS <= Obs.m.VAS ) / nrepeat` and `r sum(  Full.VAS$female.VAS <= Obs.f.VAS ) / nrepeat`  for male and female respectively

```{r old.pval, echo=FALSE, eval=F}

F.strain.var <- c(0)
counter = 1
for(u in unique(permuted.DF$strain[permuted.DF$sex == "female"])){
  # print(u)

  F.strain.var[counter] <-  sum( ( permuted.DF %>% filter(sex == "female")  %>% filter(strain == u )  )[,2:1000]  <= ( OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "female") )$var  ) / nrepeat
  counter = counter + 1
  #print(counter)
}
names(F.strain.var) <- unique( permuted.DF$strain[permuted.DF$sex == "female"]  )

#male loop
M.strain.var <- c(0)
counter = 1
for(u in unique(permuted.DF$strain[permuted.DF$sex == "male"])){
  # print(u)
  # print( sum( ( permuted.DF %>% filter(sex == "male")  %>% filter(strain == u )  )[,2:1000]  <= (    OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "male") )$var  ) / nrepeat  )

  M.strain.var[counter] <-  sum( ( permuted.DF %>% filter(sex == "male")  %>% filter(strain == u )  )[,2:1000]  <= ( OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "male") )$var  ) / nrepeat

  counter = counter + 1
  #print(counter)
}
names(M.strain.var) <- unique( permuted.DF$strain[permuted.DF$sex == "male"]  )

df.M.strain.var <- data.frame( vals = M.strain.var, strain = names(M.strain.var) )
df.F.strain.var <- data.frame( vals = F.strain.var, strain = names(F.strain.var) )

df.M.strain.var$sex <- "male"
df.F.strain.var$sex <- "female"

row.names(df.M.strain.var) <- c()
row.names(df.F.strain.var) <- c()

Strain.var.full <- rbind(df.M.strain.var, df.F.strain.var)

```

```{r p.vals, echo=FALSE, eval=FALSE}

p.val.WSB.f <- sum( (permuted.var$wsb.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )) / nrepeat

p.val.WSB.m <- sum( (permuted.var$wsb.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "male") )$var )) / nrepeat

p.val.G.f <- sum( (permuted.var$g.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "female") )$var )) / nrepeat
p.val.G.m <- sum( (permuted.var$g.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "male") )$var )) / nrepeat

p.val.LEW.f <- sum( (permuted.var$lew.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "female") )$var )) / nrepeat
p.val.LEW.m <- sum( (permuted.var$lew.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "male") )$var )) / nrepeat

p.val.KAZ.f <- sum( (permuted.var$kaz.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "female") )$var )) / nrepeat
p.val.KAZ.m <- sum( (permuted.var$kaz.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "male") )$var )) / nrepeat

p.val.SKIVE.f <- sum( (permuted.var$skive.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$var )) / nrepeat
p.val.SKIVE.m <- sum( (permuted.var$skive.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$var )) / nrepeat

p.val.PWD.f <- sum( (permuted.var$pwd.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "female") )$var )) / nrepeat
p.val.PWD.m <- sum( (permuted.var$pwd.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "male") )$var )) / nrepeat

#calculate p values
p.val.MSM.f <- sum( (permuted.var$msm.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "female") )$var )) / nrepeat

p.val.MSM.m <- sum( (permuted.var$msm.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "male") )$var )) / nrepeat

```

```{r plots.var.dist, echo=FALSE, eval=F, warning=FALSE, message=FALSE}

#good ref
#http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/

wsb.fff <- ggplot(permuted.var, aes(x = wsb_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var ) +  
  geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "WSB") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

wsb.mmm <- ggplot(permuted.var, aes(x = wsb_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "WSB") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

lew.fff <- ggplot(permuted.var, aes(x = lew_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "female") )$var ) + geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "LEW") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#lew female observed var is greater than randomized
lew.mmm <- ggplot(permuted.var, aes(x = lew_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "male") )$var ) + geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "LEW") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

g.fff <- ggplot(permuted.var, aes(x = g_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "G") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#lew female observed var is greater than randomized
g.mmm <- ggplot(permuted.var, aes(x = g_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "G") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

kaz.fff <- ggplot(permuted.var, aes(x = kaz_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "KAZ") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

kaz.mmm <- ggplot(permuted.var, aes(x = kaz_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "KAZ") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

pwd.fff <- ggplot(permuted.var, aes(x = pwd_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "PWD") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#pwd female oberserved variance is lower than the permutations.
pwd.mmm <- ggplot(permuted.var, aes(x = pwd_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "PWD") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

skive.fff <- ggplot(permuted.var, aes(x = skive_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "SKIVE") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

skive.mmm <- ggplot(permuted.var, aes(x = skive_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "SKIVE") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

msm.mmm <- ggplot(permuted.var, aes(x = msm_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "MSM") ) , aes(x=18, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

msm.fff <- ggplot(permuted.var, aes(x = msm_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "MSM") ) , aes(x=18, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#answer: The difference in CO means between wild and LL Pleucopus, is greater than if mice are randomized.

permutation.results <- ggarrange(wsb.fff, wsb.mmm,
          lew.fff, lew.mmm,
          g.fff, g.mmm, 
          kaz.fff, kaz.mmm,
          pwd.fff, pwd.mmm,
          skive.fff, skive.mmm, 
           msm.fff, msm.mmm + rremove("x.text"), 
          ncol = 4, nrow = 4)

annotate_figure(

  permutation.results, top = text_grob("permutations", color = "Black", face = "bold", size = 14) )

```

# Permutations: var(F) - var(M)

1. Choose 1 strain.
2. Compute femaleVar-maleVar (pooling all high-quality cells within sexes for that strain).
3. Permute cells between sexes.
4. Compute femaleVar-maleVar for the permuted dataset.
5. Repeat 3-4 10,000 times.
6. Compare the observed value to the distribution of permuted values to compute a p-value.
7. Repeat 1-6 separately for each strain.

```{r Obs.SDvar, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}

#observed sex difference in varince, by strain
#Q12.MLH1_data by strain

Diff.Var  <- ddply(Q12.MLH1_data, c("strain"), summarise,

                  N.F.mice = length( unique( mouse[which(sex == "female")])  ),
                  N.F.cells  = length( unique(fileName[which(sex == "female")]) ),

                  N.M.mice = length(unique(mouse[which(sex == "male")]) ),
                  N.M.cells = length(unique(fileName[which(sex == "male")]) ),

                F.mean.MLH1 = mean( nMLH1.foci[which(sex == "female")]),
                F.var.MLH1 = var(nMLH1.foci[which(sex == "female")]),

                M.mean.MLH1 = mean( nMLH1.foci[which(sex == "male")]),
                M.var.MLH1 = var(nMLH1.foci[which(sex == "male")]),

               Fvar.Mvar.dif = F.var.MLH1 - M.var.MLH1

)

Diff.Var <- add_subsp(Diff.Var)
Diff.Var <- add_species(Diff.Var)

#plot the values of difference

Var.axis.plot <- ggplot( data = (Diff.Var %>% filter(species == "M.musculus") ), aes(x=F.var.MLH1, y=M.var.MLH1, color=strain))+geom_point()+scale_color_manual(values=colors_of_strains) + ylim(c(0,30))+xlim(c(0,30))+ ggtitle("MLH1 variance by Sex")
Var.axis.plot

Dif.Var.plot <- ggplot( data = (Diff.Var %>% filter(species == "M.musculus") ), aes(x=strain, y=Fvar.Mvar.dif, color=strain))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE) + ggtitle("Degree of Sex Difference in MLH1 Variance")

Dif.Var.plot

```

Above are the plots for observed sex differences for pooled cell variance. The first plot shows that all female values are higher than male, but PWD and KAZ have the lowest female variance and close to the highest male variance.

```{r permute.by.strain, echo=FALSE}

#pool.cells within strain, calc a Fvar and Mvar, calc the difference, plug into a lis / DF

nrepeat = 10000

FM.var.diff_permuted <- data.frame( "pwd.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "PWD")], Diff.Var$N.F.cells[Diff.Var$strain == "PWD"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "PWD")], Diff.Var$N.M.cells[Diff.Var$strain == "PWD"], replace = FALSE)  )
                        )  ),

             "msm.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MSM")], Diff.Var$N.F.cells[Diff.Var$strain == "MSM"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MSM")], Diff.Var$N.M.cells[Diff.Var$strain == "MSM"], replace = FALSE)  )
                        ) ),

             "skive.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "SKIVE")], Diff.Var$N.F.cells[Diff.Var$strain == "SKIVE"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "SKIVE")], Diff.Var$N.M.cells[Diff.Var$strain == "SKIVE"], replace = FALSE)  )
                        ) ),
             "wsb.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "WSB")], Diff.Var$N.F.cells[Diff.Var$strain == "WSB"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "WSB")], Diff.Var$N.M.cells[Diff.Var$strain == "WSB"], replace = FALSE)  )
                        ) ),
             "lew.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "LEW")], Diff.Var$N.F.cells[Diff.Var$strain == "LEW"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "LEW")], Diff.Var$N.M.cells[Diff.Var$strain == "LEW"], replace = FALSE)  )
                        ) ),

             "g.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "G")], Diff.Var$N.F.cells[Diff.Var$strain == "G"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "G")], Diff.Var$N.M.cells[Diff.Var$strain == "G"], replace = FALSE)  )
                        ) ),

              "kaz.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "KAZ")], Diff.Var$N.F.cells[Diff.Var$strain == "KAZ"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "KAZ")], Diff.Var$N.M.cells[Diff.Var$strain == "KAZ"], replace = FALSE)  )
                        ) ),

             "molf.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MOLF")], Diff.Var$N.F.cells[Diff.Var$strain == "MOLF"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MOLF")], Diff.Var$N.M.cells[Diff.Var$strain == "MOLF"], replace = FALSE)  )
                        ) )

)#dataframe, FM.var.diff_permuted

#transpose the perrmuted DF, add a strain col  
#t_FM.var.diff_permuted <- t(FM.var.diff_permuted)
#Perm2.DF <- data.frame(names = row.names(t_FM.var.diff_permuted), #t_FM.var.diff_permuted); rownames(t_FM.var.diff_permuted) <- NULL
#Perm2.DF <- add_sex_chp3(Perm2.DF) #sex doesn't make sense for this
#Perm2.DF <- add_strain_Chap3(Perm2.DF)

#calc p - values  (permuted > observed )
#FM.var.diff_permuted

```

```{r plot.perm2, echo=FALSE, fig.width = 4, fig.height=2.5, message=FALSE}

#plot the first permuted DF-- don't need the transposed
pwd.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = pwd.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) + ggtitle("PWD difference in variance across sexes")  + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$pwd.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) / nrepeat )  
pwd.perm2.plot

skive.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = skive.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) + ggtitle("SKIVE difference in variance across sexes") + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$skive.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) / nrepeat )
skive.perm2.plot

msm.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = msm.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) + ggtitle("MSM difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$msm.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) / nrepeat )
msm.perm2.plot

g.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = g.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) + ggtitle("G difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$g.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) / nrepeat )
g.perm2.plot

wsb.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = wsb.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) + ggtitle("WSB difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$wsb.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) / nrepeat )
wsb.perm2.plot

lew.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = lew.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) + ggtitle("LEW difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$lew.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) / nrepeat )
lew.perm2.plot

kaz.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = kaz.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "KAZ"] ) + ggtitle("KAZ difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=5, y=750, label= sum(FM.var.diff_permuted$kaz.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "KAZ"] ) / nrepeat )
kaz.perm2.plot

molf.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = molf.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MOLF"] ) + ggtitle("MOLF difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=9, y=750, label= sum(FM.var.diff_permuted$molf.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MOLF"] ) / nrepeat )
molf.perm2.plot

```

The above plots show the difference of Female.var - Male.var for n=10,000 permutations.

For all strains the female cell variance is larger than if cells are randomized across sex. But the degree of the difference varies.  PWD and KAZ females have smaller between cell variance.

I should assess how these strain differences might be affected by the number of mice and cells, (espically MOLF).

# Total SC

I isolated cells with Q12 with MLH1 count and Skel measure. I removed rows that had a 'yes' for redo crop.

```{r mouse.total.SC, echo=FALSE}

# factor -> chr
Total.SC_DF$n <- as.character(Total.SC_DF$n)
Total.SC_DF$nMLH1.foci<- as.character(Total.SC_DF$nMLH1.foci)
Total.SC_DF$quality<- as.character(Total.SC_DF$quality)

Total.SC_DF$bin_size<- as.character(Total.SC_DF$bin_size)
Total.SC_DF$skel_size<- as.character(Total.SC_DF$skel_size)

#chr -> num
Total.SC_DF$n <- as.numeric(Total.SC_DF$n)
Total.SC_DF$nMLH1.foci<- as.numeric(Total.SC_DF$nMLH1.foci)
Total.SC_DF$quality<- as.numeric(Total.SC_DF$quality)

Total.SC_DF$bin_size<- as.numeric(Total.SC_DF$bin_size)
Total.SC_DF$skel_size<- as.numeric(Total.SC_DF$skel_size)

Q12.mouse.level.skel <- ddply(.data= (Total.SC_DF %>% filter(quality < 3) ),
                          .(mouse),
                          summarize, 
                          ncells = length(unique(Original.Name)),
                          nmice = length(unique(mouse)),
                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                          var.MLH1 = var(nMLH1.foci),

                          mean.skel = mean(skel_size),
                          var.skel = var(skel_size),

                          mean.bin = mean(bin_size)
)

Q12.mouse.level.skel <- add_strain(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_sex(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_category(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_subsp(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_species(Q12.mouse.level.skel)

mouse.level.skel <- ddply(.data= (Total.SC_DF %>% filter(quality < 5) ),
                          .(mouse),
                          summarize, 
                          ncells = length(unique(Original.Name)),
                          nmice = length(unique(mouse)),

                          Q12MLH1.mean = mean( nMLH1.foci[which(quality < 3)], na.rm = TRUE ),
                          Q12MLH1.var = var( nMLH1.foci[which(quality < 3)], na.rm = TRUE ),
            #  mean(chromosomeLength[which(hand.foci.count == 1)]
                          Q12skel.mean = mean( skel_size[which(quality < 3)], na.rm = TRUE ),
                          Q12skel.var = var( skel_size[which(quality < 3)], na.rm = TRUE ),

                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                          var.MLH1 = var(nMLH1.foci),

                          mean.skel = mean(skel_size),
                          var.skel = var(skel_size),

                          mean.bin = mean(bin_size)
)

mouse.level.skel <- add_strain(mouse.level.skel)
mouse.level.skel <- add_sex(mouse.level.skel)
mouse.level.skel <- add_category(mouse.level.skel)
mouse.level.skel <- add_subsp(mouse.level.skel)
mouse.level.skel <- add_species(mouse.level.skel)

```

```{r matching, echo=FALSE}

#plot the differences between the larger dataset and Q12

ii <- ggplot(mouse.level.skel, aes(x=Q12MLH1.mean, y=mean.MLH1, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)

skel.skel <- ggplot(mouse.level.skel, aes(x=Q12skel.mean, y=mean.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)

skel.skel.var.F <- ggplot( data = (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(sex == "female") ), aes(x=Q12skel.var, y=var.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)+facet_wrap(~ subsp, scales = "free", ncol=2)+ylim(c(0, 310000))+xlim(c(0, 310000))
#for some reason, Q12 has a greater range -- but might just be a outlier

#woah, sex differences are large. there is one female PWD with really large variance, 17apr15_PWD_f4 (11 cells)

skel.skel.var.all <- ggplot( data = (mouse.level.skel %>% filter(species == "M.musculus") ), aes(x=Q12skel.var, y=var.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)+xlim(c(0, 500000))+ylim(c(0, 500000))

```

```{r tot.sc.plot, echo=FALSE, warning=FALSE, message=FALSE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#mouse.level.skel  #mouse level mlh1 and skel
#table(mouse.level.skel$sex)

#
total.SC.plots.F <- ggplot( ( mouse.level.skel %>% filter(sex == "female")%>% filter(species == "M.musculus") %>% filter(strain != "CAST" )  ),  aes(color= strain, size = ncells))+geom_jitter(aes(x = as.factor(strain), y = var.skel, color= strain, alpha=0.5), width = .2 )+ggtitle("Variance of Total Skel Female by Mouse")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free", ncol=2)+ ylim(c(0, 350000 ) )

total.SC.plots.M <- ggplot( ( mouse.level.skel %>% filter(sex == "male")%>% filter(species == "M.musculus") ),  aes(color= strain, size = ncells))+geom_jitter(aes(x = as.factor(strain), y = var.skel, color= strain, alpha=0.5), width = .2 )+ggtitle("Variance of Total Skel Male by Mouse")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ ylim(c(0, 180000 ) )

total.SC.plots.F
total.SC.plots.M

```

This is the number of cells per category

`r kable(table((Total.SC_DF %>% filter(quality < 3) )$sex, (Total.SC_DF %>% filter(quality < 3) )$strain ))`

These plots are for the number of mice `r kable(table(Q12.mouse.level.skel$sex, Q12.mouse.level.skel$strain))`

```{r corr.total.skel, echo=FALSE, warning=FALSE, message=FALSE}

#FULL_mouse_table_Q12.MLH1
tot.SC.Fcor.test <- cor.test( (Q12.mouse.level.skel %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean.skel,
  (Q12.mouse.level.skel %>% filter(sex == "female")%>% filter(species == "M.musculus") )$var.skel)

tot.SC.Mcor.test <- cor.test( (Q12.mouse.level.skel %>% filter(sex == "male") %>% filter(species == "M.musculus")%>% filter(species == "M.musculus"))$mean.skel,
  (Q12.mouse.level.skel %>% filter(sex == "male") %>% filter(species == "M.musculus"))$var.skel)

total.SC.small.DF <- data.frame( corr = paste("corr =", c(tot.SC.Fcor.test$estimate, tot.SC.Mcor.test$estimate) ), 
                     p.val = paste("p.val =", c(tot.SC.Fcor.test$p.value, tot.SC.Mcor.test$p.value) ), sex = c("female", "male") )

total.SC.Cor.plot <- ggplot(( Q12.mouse.level.skel %>% filter(species == "M.musculus") ), aes(x=mean.skel, y=var.skel, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= total.SC.small.DF, aes(x=2000, y=300000, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= total.SC.small.DF, aes(x=2000, y=250000, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by total SC variance per Mouse")

```

# Total SC Models of mouse-level variance

The same glm models are being applied to variance and cV (M1 to M3) and sex specific models (M4 and M5). Out groups and strains with just one sex have been removed for M1-M3. Al factors are unordered.

Changing the quality cut off I apply for the total SC dataset changes the significance of the sex coefficient.

```{r total.SC.model.set.up, echo=FALSE, eval=T, warning=FALSE}

#Total.SC_DF - cells
#sex-specific   Q12.mouse.level.skel

Q12.mouse.level.skel$subsp <- factor(Q12.mouse.level.skel$subsp, ordered = FALSE)

Q12.mouse.level.skel$sex <- factor(Q12.mouse.level.skel$sex, ordered = FALSE)

Q12.mouse.level.skel$strain <- factor(Q12.mouse.level.skel$strain, ordered = FALSE)

mouse.level.skel$subsp <- factor(mouse.level.skel$subsp, ordered = FALSE)
mouse.level.skel$sex <- factor(mouse.level.skel$sex, ordered = FALSE)
mouse.level.skel$strain <- factor(mouse.level.skel$strain, ordered = FALSE)

```

```{r total.SC.var.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=T}

#Q12.mouse.level.skel
varTot.SC.M1 <- lmer(var.skel ~ subsp * sex + (1|strain), data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")  ) )

varTot.SC.M1.reduce.sex <- lmer( var.skel ~ subsp  +(1|strain), data= 
                                   ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")  ) 
                                 )

varTot.SC.M1.reduce.subp <- lmer( var.skel ~ sex  +(1|strain), data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC") ) )

#summary(varTot.SC.M1)$coefficients

varTot.SC.MM_interaction <- drop1(varTot.SC.M1, test="Chisq")$`Pr(Chi)`[2]

varTot.SC.MM_sex <- anova(varTot.SC.M1, varTot.SC.M1.reduce.sex)$`Pr(>Chisq)`[2]

varTot.SC.MM_subp <- anova(varTot.SC.M1, varTot.SC.M1.reduce.subp)$`Pr(>Chisq)`[2]

varTot.SC.MM_strain <- exactRLRT(varTot.SC.M1)$p.value

varTot.SC.MMlmer_results <- data.frame(  sub.sp = varTot.SC.MM_subp, 
        sex.results =  varTot.SC.MM_sex, 
               interaction.pval = varTot.SC.MM_interaction,  rand.pval = 
varTot.SC.MM_strain)

```

**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain. The p values for the total sc variance are as follow, `r varTot.SC.MMlmer_results`. There are significant subspecies, sex and interaction effects, (but no significant strain effects). This uses the larger dataset for total SC.

**M2**

```{r total.SC.varMLH1.glm.M2, echo=FALSE, eval=T}

#strain = fixed, 
totalSC.M2.glm.strain.fixed.full <- glm(var.skel ~ subsp * sex * strain, data= ( Q12.mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )

summ.totalSC.M2.glm.strain.fixed.full <- summary(totalSC.M2.glm.strain.fixed.full)

#round(summ.totalSC.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
#adding cell number has some significant parts

totalSC.M2.glm.strain.fixed.full <- glm(var.skel ~ subsp * sex * strain, data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )

summ.totalSC.M2.glm.strain.fixed.full <- summary(totalSC.M2.glm.strain.fixed.full)

round(summ.totalSC.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
```

The sex effect for total SC variance isn't significant, only in interactions.

**M3**

```{r total.SC.varMLH1.glm.M3, echo=FALSE, eval=T}

totalSC.M3.strain.fixed <- glm(var.skel ~ sex * strain, data= (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )

summ.totalSC.M3.strain.fixed <- summary(totalSC.M3.strain.fixed)
round(summ.totalSC.M3.strain.fixed$coefficients[,c(1,4)],5)

```

Again most strain effects than sex effects.

**M4-M5 female**

```{r total.SC.Female.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

totalSC.female.strain.subsp <- glm(var.skel ~ subsp * strain, data= (mouse.level.skel %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

summry.totalSC.female.strain.subsp<- summary(totalSC.female.strain.subsp)
round(summry.totalSC.female.strain.subsp$coefficients[,c(1,4)],5)

totalSC.female.glm.M2 <- glm(var.skel ~ strain, data= (mouse.level.skel %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )

sum_totalSC.female.glm.M2 <- summary(totalSC.female.glm.M2)
round(sum_totalSC.female.glm.M2$coefficients[,c(1,4)],5)
```

**M4-M5 male**

```{r total.SC.Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

#M4
totalSC.male.glm.M4 <- glm(var.skel ~ subsp * strain, data= (mouse.level.skel %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
sum_totalSC.male.glm.M4 <- summary(totalSC.male.glm.M4)
round(sum_totalSC.male.glm.M4$coefficients[,c(1,4)],5)

#M4.2
totalSC.male.glm.M4.2 <- glm(var.skel ~ strain, data= (mouse.level.skel %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )

sum_totalSC.male.glm.M4.2 <- summary(totalSC.male.glm.M4.2)
round(sum_totalSC.male.glm.M4.2$coefficients[,c(1,4)],5)

```

<!-- Permutations; Total SC

1. structure of data, 2. permute all the data

```{r}

```
-->

# Direct tests of MLH1 and SC compaction

The Q12 dataset for MLH1 and the larger dataset for total SC are used in these models.

**MLH1variance = SCvariance * sex**

```{r M.SC.models, echo=FALSE}
#glm
#Q12.mouse.level.skel[  ! Q12.mouse.level.skel$strain %in% remove_strains, ]

the.two.sex.glm1 <- glm(Q12MLH1.var ~ var.skel + sex, data= (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   )  )

summ.the.two.sex.glm1 <- summary(the.two.sex.glm1)
round(summ.the.two.sex.glm1$coefficients[,c(1,4)],5)

```

**MLH1variance = SCmean * sex**

```{r M2SC.skel, echo=FALSE}

the.two.sex.glm <- glm(Q12MLH1.var ~ mean.skel + sex, data= (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )
summ.the.two.sex.glm <- summary(the.two.sex.glm)
round(summ.the.two.sex.glm$coefficients[,c(1,4)],5)

```

**MLH1variance = SCvariance * SCmean * sex**

```{r M2SC.skel3, echo=FALSE}

the.three.sex.glm <- glm(Q12MLH1.var ~ mean.skel + var.skel + sex, data= 
                           (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )

summ.the.three.sex.glm <- summary(the.three.sex.glm)
round(summ.the.three.sex.glm$coefficients[,c(1,4)],5)

```