### 2020-05-21 src/SetupData.Rmd

---
title: "Setting up Data"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(xlsx)
library(reshape2)
#knitr::opts_chunk$set(echo = FALSE)

opts_knit$set(root.dir = "~./MLH1repo/")
setwd("~./MLH1repo/")

meta.data = read.csv( "~./MLH1repo/data/clean.Meta.Data.txt", sep = "\t", header=TRUE)

#MLH1_data = read.csv(here("data", "/AnonData.csv"), sep="\t", header=TRUE)
#read.csv was breaking things
MLH1_data = read.xlsx2("~./MLH1repo/data/AnonData_formated.xlsx", header=TRUE, 1)

DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)

#total SC 
Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)

Curated_BivData <-  read.csv("~./MLH1repo/data/Curated_BivData.csv", header=TRUE)

source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )
```

```{r MLH1.DF.setup, warning=FALSE, echo=FALSE}

original_DF = MLH1_data

#isolate the duplicates (fileName) from main MLH1 DFs
n_occur <- data.frame(table(original_DF$fileName))
#n_occur[n_occur$Freq > 1,] displays the dups
original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,

#use the above DF for quantifying repeatability in the Main_report.Rmd
#duplicates are cells counted twice

#are there duplicated filenames+batch
original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")

anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch and batch number --

#remove duplicates from original as well
original_DF <- original_DF %>% distinct(Original.Name, .keep_all = TRUE)#5216 to 4983

#calculate the correlation across duplicates. First issue is to match up and organize the matching 
#for each fileName - make cols, for MLH1-1 and MLH1-2
#remove Xs, for fileName with 
original_DF_duplicates <- original_DF_duplicates[!grepl("X", original_DF_duplicates$X),]
#these are images counted across mulitple batches

#keep just 1 of duplicates from MLH1_data
#MLH1_data <- MLH1_data[duplicated(MLH1_data$fileName),]
MLH1_data <- MLH1_data %>% distinct(fileName, .keep_all = TRUE)

```

`r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.  `r length(original_DF_duplicates$fileName)` rows (images) have been counted across mulitple batches. These are removed in the MLH1_data DF.

```{r clean_MLH1_data, echo=FALSE}
#Clean up main DF, MLH1_data
#remove duplicated data from MLH1 -- remove 1 or the two rows
#https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/
#use distinct() in dlpyr()
MLH1_data <- MLH1_data %>% distinct(Original.Name, .keep_all = TRUE)#5216 to 4983

MLH1_data$Original.Name <- as.character(MLH1_data$Original.Name)
MLH1_data$fileName <- (MLH1_data$Original.Name)

#these introduce NAs
MLH1_data$n <- as.character(MLH1_data$n)
MLH1_data$nMLH1.foci<- as.character(MLH1_data$nMLH1.foci)
MLH1_data$quality<- as.character(MLH1_data$quality)

MLH1_data$n <- as.numeric(MLH1_data$n)
MLH1_data$nMLH1.foci<- as.numeric(MLH1_data$nMLH1.foci)
MLH1_data$quality<- as.numeric(MLH1_data$quality)

#remove X's and NAs
MLH1_data <- MLH1_data[ !grepl("X", MLH1_data$X) , ]
MLH1_data <- MLH1_data[!(is.na(MLH1_data$nMLH1.foci) | MLH1_data$nMLH1.foci==""), ]
#this should also have the empty column for removing X
MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "X",]
MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]

```

`r length(original_DF$fileName[ grepl("X", original_DF$X)] )` images (rows) have X's, they are removed from the main dataset.

```{r misc.setup, echo=FALSE}

#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4',

                       'TOM'= 'indianred1','AST' = 'tan2',
                       'CZECH'= 'brown1',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
#4!!

```

# Mouse Data MLH1

```{r exclude.mice, echo=FALSE}
#remove mice -- 18nov17_WSB_f5 (only 1 cell)
exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                 "3sep18_AST_m1",
                  "3sep18_TOM_m1",
                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]

#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
#juvi.mice_DF <- juvi.mice_DF[!(is.na(juvi.mice_DF$mouse) | juvi.mice_DF$mouse==""), ]

#juvi mice ~ 2 weeks old
juvi_mice.list <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )

#make juvi.mice DF
juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice.list),]

#remove juvi mice from big MLH1 -- but Then i need to remake the mouse_table
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice.list, ]

# > 25 weeks
# I might need to keep some -- espcially outgroups
#old mice -- This list includes mice which don't have MLH1 observations, but I have records of dissecting them

old.mice.list <- c(
  "12oct15_WSB_m1","9oct17_WSB_m1","1mar18_WSB_m2","21dec14_WSB_m2","21dec14_WSB_m3", "1mar18_WSB_m1",
  "14mar15_G_m1","14mar15_G_m2","24mar15_G_m1","19oct15_G_m1",
  "8may17_LEW_m1","18sep17_LEW_m1","8may17_LEW_m2","20dec16_LEW_m2",
  "5jul17_PERC_m1",

  "27oct18_SKIVE_m2","4feb19_SKIVE_m1",
  "1feb18_KAZ_m1","11jan18_KAZ_m1","19may18_KAZ_m2","9apr18_KAZ_m1","9apr18_KAZ_m2",
  "20dec18_CZECH_m1","3dec18_CZECH_m1",
  "28jun18_TOM_m1","3jul18_TOM_m1",
  "27aug18_AST_m1","7aug18_AST_m1",

  "13nov16_MSM_m1","13nov16_MSM_m2","4feb19_MSM_m3","4feb19_MSM_m4","9oct17_MSM_m2","4feb19_MSM_m2","30may17_MSM_m1","5jul17_MSM_m1",
  "4feb19_MOLF_m1",

  "24aug17_CAST_m1","23sep17_CAST_m1","23sep17_CAST_m2","30jun16_CAST_m3","5jul17_CAST_m1",
  "27aug18_HMI_m2","28jun18_HMI_m1","27aug18_HMI_m1","4dec17_HMI_m1",

  "11jan18_SPRET_m1","27aug18_SPRET_m1","7aug18_SPRET_m2","7aug18_SPRET_m1","13jul17_SPRET_m1","2aug18_SPRET_m1",
  "19may18_SPIC_m1","19may18_SPIC_m2","30oct17_SPIC_m1","11jan18_SPIC_m1",
  "1mar18_CAROLI_m1","9apr18_CAROLI_m1"
)

```

`r length(old.mice.list)` old mice (> 25 weeks) were dissected, some of the RIKEN musc strains measures only have old mouse observations.

`r length(juvi_mice.list)` juvinile mice were dissected, mostly only for DMC1 counts. Few had good MLH1 cells so they were removed from the main MLH1 DF.

```{r adj_nMLH1, echo=FALSE, eval=FALSE}
#add a column with male adjusted MLH1 values (+1 to all males)
MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)

#adjusted MLH1 no longer used
```

```{r reorder.DF, echo=FALSE}
#this is most important for plotting
#reorder dataframe
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name) )

#order the batches in MLH1 and original
#apply this ordering to all other DF?

```

```{r Batch.Stats, echo=FALSE, eval=TRUE}
#plot / table comparing number of images in quantification batches
#basic visualization for number of images in batches

#ToDo:  correct the case of all batches, better names
#set an order to batches?
original_batches <- as.data.frame(table(original_DF$Batch) ) #these look bad
colnames(original_batches) <- c("BatchNum", "orginal")

final_batches <- as.data.frame(table(MLH1_data$Batch) ) #these look bad
colnames(final_batches) <- c("BatchNum", "main")

batch.stats <- merge(original_batches,final_batches, by="BatchNum")## mm is the table of batch stats
##batch18 has the wrong number, ~390
```

All images were manually quantified after the file name was anonimized. The original (full) dataset contains the quantification batches with the following number of images. Each batch csv file was merged into Anon.csv by hand.

`r kable(final_batches, caption = "Number of Images across all the anonimized Batches") %>% kable_styling()`

# Quality Cleaning steps

The original dataset was cleaned up in the following ways before analysis. Settting up main data file with explanation and examination of data. Make the raw original DF and main working DF with bad cells exlcuede. The original dataset had `r length(original_DF_duplicates$fileName)` duplicated file names. These were removed and can be used to quantify repeatability or repair. Using the distinct() function for the MLH1_data, the first duplicated filename is kept and the second removed.

Take out mice from the exclude list based on the output from Mouse level quality plots.

1. The quality of MLH1 counts for individual mice is / was assessed in the MouseLevelPlots Rmd (can I link this file?). A list, mice.to.exclude was made.

2. The list of excluded mice is applied in the MLH1.setup.data script, to remove these mice from the main dataset/dataframe. This script saves .RData file of the dataset which is loaded into Rmd files for analysis

Mouse specific plots of CO count and quality were examined for general quality from single mice. Mice which produced few cells of low quality were excluded to the 'HQ' (high quality) dataset. (The full dataset was used to look at some patterns)

## MLH1 mouse level tables**

```{r HQ.mouse.table, echo=FALSE}
#Make a mouse level table

#add date image
count=1
for(i in MLH1_data$Original.Name){
   # print(i)
    templist= strsplit(i, split="_")[[1]]
    c = (templist[1])
    MLH1_data$image.string[count] <-  c
    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
    count= count +1
}
#works better if dat is converted out of image
MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')

#change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
#

Mouse.Table_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm=TRUE),

                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm=TRUE),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        #dates Imaged not in the DF
                        oldest.image.date = min(date.imaged),
                        newest.image.date = max(date.imaged)
                        #quality?
)

Mouse.Table_HQ <- add_strain(Mouse.Table_HQ)
Mouse.Table_HQ <- add_subsp(Mouse.Table_HQ)
Mouse.Table_HQ <- add_sex(Mouse.Table_HQ)
Mouse.Table_HQ <- add_category(Mouse.Table_HQ)
Mouse.Table_HQ <- add_species(Mouse.Table_HQ)

#remove juvi mice via the list
#remove bad ages
#this is removing female

#this removes everything but males in these categories
#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]

#i don't understand why so many WSB female are removed..
# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)

```

```{r merge.meta_MLH1, echo=FALSE}

#merge - MLH1 with metdata

batch.mouse.count <- table(original_DF$mouse,original_DF$Batch) #mouse count by batch
batch.mouse.count.DF <- as.data.frame(batch.mouse.count)#this has a dif str, 3 cols -- like it was melted
batch.mouse.count.DF <- batch.mouse.count.DF[!(is.na(batch.mouse.count.DF$Freq)|batch.mouse.count.DF$Freq==0),]

#subsp in metadata not fully correct
meta.data <- add_strain(meta.data)
meta.data <- add_subsp(meta.data)
meta.data <- add_sex(meta.data)
meta.data <- add_category(meta.data)
meta.data <- add_species(meta.data)

#add DOB to this table?
mice.frm.meta <- Mouse.Table_HQ[as.character(Mouse.Table_HQ$mouse) %in% as.character(meta.data$mouse),]

Mouse.Table_HQ.OG <- Mouse.Table_HQ

Mouse.Table_HQ <- Mouse.Table_HQ.OG

#during merging -- try to just merge a few cols: 
Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
#this removes ~28 mice

old.mice.df <- Mouse.Table_HQ[(Mouse.Table_HQ$age.weeks > 25),]
#remove the NAs
old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]

Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]

```

Made up the main mouse level table:

- (Q1-Q4)

- no juvi mice

Subset DFs from here of other analysis.

```{r Q12_mouse.table, echo=FALSE}

#This DF is used for some MM, apply the HetC MM certia...
#this is a DF with a different set of quality certeria

#Make a mouse level table
Q12_mouse_table_MLH1 <- ddply(MLH1_data[MLH1_data$quality < 3,], c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)

#REMOVE mice with fewer than 3 cells
Q12_mouse_table_MLH1 <- Q12_mouse_table_MLH1[Q12_mouse_table_MLH1$Ncells > 3,]

Q12_mouse_table_MLH1 <- add_strain(Q12_mouse_table_MLH1)
Q12_mouse_table_MLH1 <- add_subsp(Q12_mouse_table_MLH1)
Q12_mouse_table_MLH1 <- add_sex(Q12_mouse_table_MLH1)
Q12_mouse_table_MLH1 <- add_category(Q12_mouse_table_MLH1)

#use things from prv block
#mice.frm.meta 
#meta.data
#mouse.table_w.Ages

#MERGE HERE
Q12_mouse_table_w.Ages <- merge(Q12_mouse_table_MLH1, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
#what are the differences between these two

#remove the NAs
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[!(is.na(Q12_mouse_table_w.Ages$category.x) | Q12_mouse_table_w.Ages$category.x==""), ]

#CLEAN up and format for mixed models
Q12_mouse_table_w.Ages <- subset(Q12_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

Q12_mouse_table_w.Ages <- add_strain(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_sex(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- add_category(Q12_mouse_table_w.Ages)

Q12_mouse_table_w.Ages <- add_species(Q12_mouse_table_w.Ages)
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$species == "M.musculus",]

#REMOVE EXTRA STRAINS -- why?
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
#remove musc strains
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]

#REMOVE WRONG AGES
#afriad this is removing NA's
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[( (Q12_mouse_table_w.Ages$age.weeks < 25) &  (Q12_mouse_table_w.Ages$age.days != 13) ) | is.na(Q12_mouse_table_w.Ages$age.weeks), ] #keeps mice without ages, and removes juvinilles

#unclass order factors --(applies to the models)
#main df with variables coded as chr
Q12_mouse_table_w.Ages$subsp <- factor(Q12_mouse_table_w.Ages$subsp, ordered = FALSE)
Q12_mouse_table_w.Ages$sex <- factor(Q12_mouse_table_w.Ages$sex, ordered = FALSE)
Q12_mouse_table_w.Ages$strain <- factor(Q12_mouse_table_w.Ages$strain, ordered = FALSE)
Q12_mouse_table_w.Ages$species <- factor(Q12_mouse_table_w.Ages$species, ordered = FALSE)

```

```{r Q123_mouse.table, echo=FALSE, eval=FALSE}

Q123_mouse_table <- ddply(MLH1_data[MLH1_data$quality < 4,], c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)

#REMOVE mice with fewer than 3 cells
Q123_mouse_table <- Q123_mouse_table[Q123_mouse_table$Ncells > 3,]

Q123_mouse_table <- add_strain(Q123_mouse_table)
Q123_mouse_table <- add_subsp(Q123_mouse_table)
Q123_mouse_table <- add_sex(Q123_mouse_table)
Q123_mouse_table <- add_category(Q123_mouse_table)

```

### DataFrame for Mixed Models

Set up the dataframes in results or MS.

Below I format the DF's for the models;

i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus.

update the dissection document (that makes the meta.data DF)

### Number of mice per category

 number of mice are in the main DF (discribe the results for missing mice)

Number of cells per mouse, (were any mice excluded for low numbers of cells?) (were all cells pooled?)

## Mouse Ages

Testes were collected from mice at various ages. For some poor breeding mice, I couldn't collect males at younger ages.

Is there a detectable age effect?

```{r Age.Hists, echo=FALSE, warning=FALSE}
#age.weeks col is missing

male_mouse_table_w.Ages <- Mouse.Table_HQ[(Mouse.Table_HQ$sex == "male") &(Mouse.Table_HQ$category != "other"),]

all.ages_plot <- ggplot(Mouse.Table_HQ[Mouse.Table_HQ$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
all.ages_plot

#male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[male_mouse_table_w.Ages$age.weeks < 20,]

male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[( (male_mouse_table_w.Ages$age.weeks < 25) &  (male_mouse_table_w.Ages$age.days != 13) ) | is.na(male_mouse_table_w.Ages$age.weeks), ]

#remove juvi mice!!
#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ]

#droplevels on DF above

good.ages_plot <- ggplot(male_mouse_table_w.good.ages, aes(age.weeks))+geom_histogram(binwidth = 1)+facet_wrap(~category)+ggtitle("Good Male Age Distributions")

#male_mouse_table_w.good.ages
```

# DMC1 Section

move the chunk below to setup

```{r DMC1.setup, echo=FALSE}
#load DMC1 csv
#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)

DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_category(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
                 .(category),
                 summarize,

                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.L$stage <- "early Z"

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "late Z"

DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)

DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
```

# BivData

for these full curated, add all the labels and save them as excel files in the BivData setup file.

full curated, reduced datasets

Are the strain levels plots needed?

Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.

```{r FULL.mouse.BivData.table, echo=FALSE}

#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36

table(Curated_BivData$hand.foci.count, exclude = NULL)
#there are ~600 rows with SC and no hand foci count (blah)
#Remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )

length( (Curated_BivData  %>%  filter(  is.na(hand.foci.count) == T ) )$fileName )

table(cleanCurated_BivData$hand.foci.count, exclude = NULL)

mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
                       .(mouse),
                       summarize,

                      Nmice = length(unique(mouse)),
                       Ncells = length(unique(fileName)),
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),
                      nbiv = length(Obj.ID),

                      tot.biv = length( !is.na(hand.foci.count) ),
                      # mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbiv), 3),
                        SC.cV = cv(chromosomeLength),

                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),

                  proportion_0CO = (n0CO / nbiv),
                  proportion_1CO = (n1CO / nbiv),
                 proportion_2CO = (n2CO / nbiv),
                     proportion_3CO = (n3CO / nbiv),

                  #calq mean.sc by chrm class
                 mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                #mean 1CO.pos
                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
                #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
                # IFD things
             mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
             mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),

              #sis.co.ten
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE),
                 #sis.co.ten
                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
)

mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
mouse.level.BivData_table <- add_species(mouse.level.BivData_table)

mouse.level.BivData_table <- add_category(mouse.level.BivData_table)

#unordered for the mixed models (move this to the MM blocks...?)
mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)

mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))

```

```{r CurBiv.Mouse.av.TABLE, echo=FALSE, eval=FALSE}

#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")

#cleanCurated_BivData
#removed the no hand foci count

#there's alread a Mouse.Table_BivData

mouse.BivData_avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(unique( fileName) ),
                        nbiv = length(Obj.ID),

                        mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

              #can't figure out how to get this format to work  
                       n0CO = sum(length( which(hand.foci.count == 0) )  ),
                       n1CO = sum(length( which(hand.foci.count == 1) )  ),
                      n2CO = sum(length( which(hand.foci.count == 2) )  ),
                      n3CO = sum(length( which(hand.foci.count == 3) )  ),
              #     
                  proportion_0CO = (n0CO / nbiv),
                  proportion_1CO = (n1CO / nbiv),
                 proportion_2CO = (n2CO / nbiv),
                     proportion_3CO = (n3CO / nbiv),

                  #calq mean.sc by chrm class
                 mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                #mean 1CO.pos
                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
                #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                # IFD things
            #  mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
            # mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
mouse.avs_4MM <- add_category(mouse.avs_4MM)
mouse.avs_4MM <- add_species(mouse.avs_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)

#REMOVE JUVI mice
juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice.list, ]  
mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice.list, ]

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))

mouse.BivData.Table <- mouse.avs_4MM
```

```{r no.Foci.Pos, echo=FALSE}
#no posittion
#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]

#table(just.SC.length$category)
```

Main files from this Script:

  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures

  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information

  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)

  4. **MELT_ABS**: make melt version of foci positions for abs measuers

  5. CELL level table (full all mice) (use this to make long biv dataset)

  6. Long Biv Dataset

  7. mouse table (full all mice)

  8. **Strain DFs** for Het models (summary strain DFs and tables?)

  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models

  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**

## Cell Level Table BivData

```{r cell.level.table, echo=FALSE}

cell.level.BivData_table <- ddply(.data=Curated_BivData, 
                       .(fileName),
                       summarize, 
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                       ncells = length(unique(fileName)),
                       nbivs = length(unique(Obj.ID)),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number

                       cv.SC = cv(chromosomeLength),

#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),

                #this below might not be correct
                     mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),

                       var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)

#cent_dis
                      mean_cent_dist = mean(dis.cent, na.rm=TRUE),
                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]], #calq threshold for all cells, pull out 'almost.whole.cells' later
                      Qar.3 = quantile(chromosomeLength, na.rm = TRUE)[[3]],
                      Qar.2 = quantile(chromosomeLength, na.rm = TRUE)[[2]],
                      Qar.1 = quantile(chromosomeLength, na.rm = TRUE)[[1]]

                        )

cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
cell.level.BivData_table <- add_sex(cell.level.BivData_table)
cell.level.BivData_table <- add_strain(cell.level.BivData_table)
cell.level.BivData_table <- add_category(cell.level.BivData_table)
cell.level.BivData_table <- add_species(cell.level.BivData_table)

#as.data.frame( head(cell.level.BivData_table) )
#use quartille 2 as a min, on average it grabs 4 
```

```{r show.pass.rate, echo=FALSE}

pass.rate.hist <- ggplot(cell.level.BivData_table, aes(x = nboxes_passed) )+geom_histogram(binwidth = 1, alpha=.6) + facet_wrap(~category)+ggtitle("number of objects ID'd per cell")
pass.rate.hist

```

make histogram of the pass rate per cell.

3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.  
8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev

## Reduced Bivalent Dataset (Long and Short)

I want to make a dataset of 'longest' 25% chromosomes from cells with at least 16 segmented bivalents from a single cell.

```{r Reduced.Biv.DataSet.chrm.dataset, echo=FALSE}

#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

#check for blank obj.id
BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")

#check for dup
#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
#dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
#no duplicates

#remove duplicate obj.ids
#BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
#BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]

#make sure obj.id and fileName are charcter for below
BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)

missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ] 
#all these files stem from

#these are bivalents which don't appear the in cell level table, which can break the loop
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
#final cleaned BivData table, 2428

#2. calq the Q4 threshold
#loop for greping the mouse from file name ect
#each Q is a quartille
for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_" )[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q3 <-cell.level.BivData_table$Qar.3[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q2 <-cell.level.BivData_table$Qar.2[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q1 <-cell.level.BivData_table$Qar.1[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
    BivData_from_15biv_cells$cell.Q3[i] <- the.Q3
  BivData_from_15biv_cells$cell.Q2[i] <- the.Q2
  BivData_from_15biv_cells$cell.Q1[i] <- the.Q1

  the.FN <- as.character(cell.level.BivData_table$fileName[which(grepl(new.egg, cell.level.BivData_table$fileName ) )] )

  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)
  BivData_from_15biv_cells$Q1.FN[i] <- as.character(the.FN)
}

#3.mark which bivalents are above the threshold

BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= as.numeric(BivData_from_15biv_cells$cell.Q4)  ), "yes", "no" )

real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]#713 observations

#isolate the short bivdata
BivData_from_15biv_cells$short.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength <= as.numeric(BivData_from_15biv_cells$cell.Q2)  ), "yes", "no" )

real.short.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$short.Biv == "yes",]# 708 much few than

```

The above chunk makes the *real.short.bivData* and *real.long.BivData* BivData.DFs.

```{r plot.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}

#showing SC lengths per cell, with lines for the quartille values.
ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
  geom_hline(aes(yintercept= cell.Q2), color='red') +
 geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
 geom_hline(aes(yintercept= cell.Q4), color='black')

  #this plot shows that the Q1 is capturing only single values -- I should try Q2
#use Q2 as a 
```

The chunck below makes the mouse level tabeles for the 2 reduced DFs.

```{r Mouse.Long.Biv_table, echo=FALSE}
# make a table for the Mixed Models -- from the long.BivData

Long.biv_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        #SC lengths by class
                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

        #          mean_siscoten = mean(siscoten)
        #
)
#mouse averages for long.biv

Long.biv_mouse_table <- add_sex(Long.biv_mouse_table)
Long.biv_mouse_table <- add_strain(Long.biv_mouse_table)
Long.biv_mouse_table <- add_subsp(Long.biv_mouse_table)
Long.biv_mouse_table <- add_category(Long.biv_mouse_table)

#unordered for the mixed models (move this to the MM blocks...?)
Long.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
Long.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)

#rec group
Long.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Long.biv_mouse_table$category), 1, 
                            ifelse(grepl("MSM", Long.biv_mouse_table$category), 1,
                             ifelse(grepl("SKIVE", Long.biv_mouse_table$category), 1, 0)) )

#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)

#1) Long Biv for

#remove all the non-used strains!!
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
#Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)

#remove na's
#Long.biv_mouse_table_HetC <- #Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | #Long.biv_mouse_table_HetC$mouse==""), ]

#table(droplevels( Long.biv_mouse_table_HetC$category) )  #for double checking

#2) Long Biv for male.poly
#just males, just Musculus
#Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
#Long.biv_mouse_table_male.poly <- #Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
```

```{r Mouse.Short.Biv_table, echo=FALSE}
# make a table for the Mixed Models -- from the long.BivData

Short.biv_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        #SC lengths by class
                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

        #          mean_siscoten = mean(siscoten)
        #
)
#mouse averages for long.biv

Short.biv_mouse_table <- add_sex(Short.biv_mouse_table)
Short.biv_mouse_table <- add_strain(Short.biv_mouse_table)
Short.biv_mouse_table <- add_subsp(Short.biv_mouse_table)
Short.biv_mouse_table <- add_category(Short.biv_mouse_table)

#unordered for the mixed models (move this to the MM blocks...?)
Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)

#rec group
Short.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Short.biv_mouse_table$category), 1, 
                            ifelse(grepl("MSM", Short.biv_mouse_table$category), 1,
                             ifelse(grepl("SKIVE", Short.biv_mouse_table$category), 1, 0)) )

```

The above 2 plots make mouse level tables for the reduced single biv datasets.

## Mouse Level Single BivData Table

Mouse level tables are the main meat of the models. There will be several for use in different models. Format these tables for run the mixed models (unordered factors, ect).

The block below creates clean.BivData, which removes ~600 rows with just SC measures and no position data.

```{r mouse.table.HetC.MM, echo=FALSE}

#this mouse level table -- should not have strains without female obs
#only Mmusculus

Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(  unique(fileName) ),
                      tot.biv = length(Obj.ID),

                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#
Mouse.Table_BivData_4MM <- add_sex(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_strain(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_subsp(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_category(Mouse.Table_BivData_4MM)
Mouse.Table_BivData_4MM <- add_species(Mouse.Table_BivData_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
Mouse.Table_BivData_4MM$subsp <- factor(Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
Mouse.Table_BivData_4MM$sex <- factor(Mouse.Table_BivData_4MM$sex, ordered = FALSE)
Mouse.Table_BivData_4MM$strain <- factor(Mouse.Table_BivData_4MM$strain, ordered = FALSE)

#remove all the non-used strains!!

Mouse.Table_BivData_4MM <-  Mouse.Table_BivData_4MM %>%  filter(strain != "AST")  %>% filter(strain != "TOM")  %>% filter(strain != "CZECH")

#Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
#Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),] #so many NAs still

Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
                              ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)))

```

```{r mouse.table.Male.Poly, echo=FALSE, eval=FALSE}

#this makes a redundant table

#this table will have male BivData
#it will include Musc strains

Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

Male.poly.Mouse.Table_BivData_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

Male.poly.Mouse.Table_BivData_4MM <- add_sex(Male.poly.Mouse.Table_BivData_4MM)
Male.poly.Mouse.Table_BivData_4MM <- add_strain(Male.poly.Mouse.Table_BivData_4MM)
Male.poly.Mouse.Table_BivData_4MM <- add_subsp(Male.poly.Mouse.Table_BivData_4MM)
Male.poly.Mouse.Table_BivData_4MM <- add_category(Male.poly.Mouse.Table_BivData_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
Male.poly.Mouse.Table_BivData_4MM$subsp <- factor(Male.poly.Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
Male.poly.Mouse.Table_BivData_4MM$sex <- factor(Male.poly.Mouse.Table_BivData_4MM$sex, ordered = FALSE)
Male.poly.Mouse.Table_BivData_4MM$strain <- factor(Male.poly.Mouse.Table_BivData_4MM$strain, ordered = FALSE)

#remove females
Male.poly.Mouse.Table_BivData_4MM$sex <- droplevels(Male.poly.Mouse.Table_BivData_4MM$sex)

```

analysis using the mouse level table -- to justify that using mouse averages is a good idea.

## BivData Melt Position Table

The BivData MELT DF is primarily for the nrmF1.pos plots/analysis. It is a Melted DF for raw and normalized foci positions from the Curated_BivData.

```{r BivData.MELT.block1, echo=FALSE}

#not sure what the melt table is needed for
col.names.list <- colnames(Curated_BivData)

#all of the cols to use
MELT_Curated_BivData <-   melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse", "category",      "strain",        "sex",           "subsp",         "species", "IFD3_ABS", "IFD1_ABS","IFD2_ABS","IFD1_PER",
              #remoevd PER.Foci positions
"dis.cent", "dis.cent.PER",  "telo_dist", "telo_dist_PER", "SisCoTen", "Rec.group",
              #leave out the Foci
                "Foci1","Foci2","Foci3", "Foci4")],

            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
       "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count",

       "notes", "mouse", 
      "category","strain","sex","subsp","species", "IFD3_ABS","IFD1_ABS","IFD2_ABS","IFD1_PER",
              #remoevd PER.Foci positions
"dis.cent", "dis.cent.PER",  "telo_dist", "telo_dist_PER", "SisCoTen", "Rec.group"))

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]

#this might not be right
colnames(MELT_Curated_BivData) <- c( col.names.list[c(1:11, 16:26, 31:36)], "Foci_Type", "Pos")

#
MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)

#calq normalized melt pos
MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]

```

The above chunk makes the MELT table (*MELT_Curated_BivData*), which can be used for Foci position comparison -- But I might not really use this for that.

# Total SC

The Total_SC excel file is made from total_SC.rmd, it removes 'X' cells and unclear quality scores.

```{r clean.total.sc, echo=FALSE}
#everything are factors again
#clean
Total.SC_DF$notes <- as.character(Total.SC_DF$notes)
Total.SC_DF$REDO.crop <- as.character(Total.SC_DF$REDO.crop)

redo.crops.rows <- Total.SC_DF[ grepl("yes", Total.SC_DF$REDO.crop) , ]#this catches most of the notes

#remove 
Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% redo.crops.rows$fileName, ]
#284 rows removed

# factor -> chr
Total.SC_DF$n <- as.character(Total.SC_DF$n)
Total.SC_DF$nMLH1.foci<- as.character(Total.SC_DF$nMLH1.foci)
Total.SC_DF$quality<- as.character(Total.SC_DF$quality)

Total.SC_DF$bin_size<- as.character(Total.SC_DF$bin_size)
Total.SC_DF$skel_size<- as.character(Total.SC_DF$skel_size)

#chr -> num
Total.SC_DF$n <- as.numeric(Total.SC_DF$n)
Total.SC_DF$nMLH1.foci<- as.numeric(Total.SC_DF$nMLH1.foci)
Total.SC_DF$quality<- as.numeric(Total.SC_DF$quality)

Total.SC_DF$bin_size<- as.numeric(Total.SC_DF$bin_size)
Total.SC_DF$skel_size<- as.numeric(Total.SC_DF$skel_size)

```

```{r Mouse.total.sc.table, echo=FALSE}

#Total.SC_DF

#mouse level
mouse.total.skel <- ddply(.data=Total.SC_DF,
                          .(mouse),
                          summarize, 
                          ncells = length(unique(Original.Name)),
                          nmice = length(unique(mouse)),
                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                          var.MLH1 = var(nMLH1.foci),

                          mean.skel = mean(skel_size),
                          var.skel = var(skel_size),

                          mean.bin = mean(bin_size)
)

mouse.total.skel <- add_strain(mouse.total.skel)
mouse.total.skel <- add_sex(mouse.total.skel)
mouse.total.skel <- add_category(mouse.total.skel)
mouse.total.skel <- add_subsp(mouse.total.skel)
mouse.total.skel <- add_species(mouse.total.skel)

#add rec group
mouse.total.skel$Rec.group <- ifelse(grepl("PWD male", mouse.total.skel$category), 1, 
                              ifelse(grepl("MSM male", mouse.total.skel$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.total.skel$category), 1, 0)))

```

The above chunk makes mouse level *mouse.total.skel*, it uses the whole DF.

# Saving Data file

Now that the dataframes have been process and formated, they will be saved as RData. the main files are:
  1. MLH1_data and original_DF
  2. AP_mouse_table and AP_mouse_table_w.Ages
  3. meta.data

work on cleaning up all other df with bad names

```{r save.RData, echo=FALSE}

#remove non relevant files -- not sure where they load from
#BD data and Peromyscus data
#rm(c(BD_MLH1_data, BDmouse_level_table, MLH1_data.hq) )

save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")

#  OutPut: big large MLH1_data (AP's) df, big DF of BD with just the mice I want.
#  MLH1_data_table, means and variance of AP and BD MLH1 values. made from seperate tables from AP and BD data.
#  make sure decimal places are consistant

```